### 2020-04-24 ms_8.Rmd
Initial commit (data is too large to commit right now)

---
title: "Maps"
author: "Julia Englebert"
date: "4/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(scipen = 999)

library(haven)
library(tidyverse)
library(gt)
library(devtools)
library(tibble)
library(foreign)
library(scales)

# packages for mapping
library(leaflet)
library(spdplyr)
library(rmapshaper)
library(sf)
```

```{r testModel}
#
# model1 <- lm(data = mwdata, perc_school ~ state + land_in_farms)
```

# DATA WRANGLING

```{r allData}
# 1800s data

# Will comment on 1870s data, then repeat the process for all subsequent years
# Note that the codenames ARE DIFFERENT FOR EACH YEAR
# Even if the variable was measured in the same way
# See original codebook for details

# Read in all 1870 csv files
# Originally, R was returning this message:
# Column `STATE` joining factors with different levels, coercing to character vectorColumn `COUNTY` joining
# factors with different levels, coercing to character vector
# I got rid of it by converting all factors to characters

data1870a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds16_1870_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1870b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds17_1870_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

# Join the two datasets by STATEA and COUNTYA, which are the state and county codes

data1870j <- full_join(data1870a, data1870b, by = c("STATEA", "COUNTYA")) %>%

  # Mutate to get rid of the .x suffix on variables

  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%

  # Mutate to rename varialbes of interest to uniform, comprehensible names
  # Some existing varialbes are combined to create the desired variables
  # See codebook

  mutate(value_farm = AJV001, n_farm = AJS001, cows =  AJ0003+AJ0005, horses = AJ0001, 
         avg_size = (AJT001*2 + AJT002*6 + AJT003*15 + AJT004*35 + AJT005*75 + AJT006*300 + AJT007*750 + 
                       AJT008*1200)/AKP001, value_implements = AKL001, pop = AJ3001, male = AKD001, 
         female = AKD002, enrolled = ALG001, urban_pop = AKE001, pct_urban = AKE001/AJ3001*100, 
         illiterate = AJ7001, acres_farms = AJU001+ AJU002+AJU003,

         # Even though irrigation data doesn't exist for this year, I'm filling the columns in with NA
         # Later years will have irrigation data

         acres_irrigated = NA, pct_irrigated = NA) %>%

  # Mutate again to make new variables out of the ones you just created in the last mutate

  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%

  # Select the desired variables

  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm,
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female,
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# Now the dataset is much more manageable!

# Start on the next year, 1880, repeating the exact same process
# Lots of NA values due to scant data this year

data1880a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds22_1880_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1880b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds23_1880_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1880j <- full_join(data1880a, data1880b, by = c("STATEA", "COUNTYA")) %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = APN001+APO001+APQ001, n_farm = AOQ001, cows =  AOG004+AOG005, horses = AOG001, 
         avg_size = APG001, value_implements = APO001, pop = AOT001, male = AP1001, female = AP1002, 
         enrolled = NA, urban_pop = AO4001, pct_urban = AO4001/AOT001*100, illiterate = NA, 
         acres_farms = AOR001, acres_irrigated = NA, pct_irrigated = NA) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# Start on 1890 data

data1890a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds26_1890_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1890b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds27_1890_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1890j <- full_join(data1890a, data1890b, by = c("STATEA", "COUNTYA")) %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = AUK001+AUK002+AUK003, n_farm = AS9001, cows = ASY002, horses = ASY003, 
         avg_size = ATW001, value_implements = AUK002, pop = AUM001, male = AV0001+AV0003+AV0005+AV0007, 
         female = AV0002+AV0004+AV0006+AV0008, enrolled = AVK005+AVK006+AVK007+AVK008, urban_pop = AUU001, 
         pct_urban = AUU001/AUM001*100, illiterate = NA, acres_farms = AUJ001+AUJ002, 
         acres_irrigated = AT2001, pct_irrigated = AT2001/(AUJ001+AUJ002)*100) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# Before I mutated all of the data, running this code showed that all .x values are the same 
# as all .y values, so it was safe to eliminate the double columns as long as "problems" returned 0
# And it did!

#problems <- bind_rows(data1870j, data1880j, data1890j) %>% 
  #select(STATEA, COUNTYA, GISJOIN.x, GISJOIN.y, YEAR.x, YEAR.y, STATE.x, STATE.y, 
         #COUNTY.x, COUNTY.y, AREANAME.x, AREANAME.y) %>%
  #mutate(same = ifelse(GISJOIN.x == GISJOIN.y & YEAR.x == YEAR.y & STATE.x == STATE.y & 
                         #COUNTY.x == COUNTY.y & AREANAME.x == AREANAME.y, TRUE, FALSE)) %>%
  #filter(same == "FALSE") %>%
  #nrow()

# Read in the 1900s data

# Only do decennial census data
# Intermittent years only have AF12001: estimated population

# 1900

data1900a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds30_1900_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1900b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds31_1900_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1900j <- full_join(data1900a, data1900b, by = c("STATEA", "COUNTYA")) %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = AWV001, n_farm = AW3001, cows = AW7001, horses = AW7002, 
         avg_size = AXE001, value_implements = AWW003, pop = AWS001, male = AZ0001,female = AZ0001, 
         enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = AYS001+AYS002+AYS003+AYS004+AYS005, 
         acres_farms = AWT001, acres_irrigated = AX8001, pct_irrigated = AX8001/AWT001*100) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# 1910

data1910a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds36_1910_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1910b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds37_1910_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1910j <- full_join(data1910a, data1910b, by = c("STATEA", "COUNTYA")) %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = A0F001, n_farm = A1H001, cows = A1F001+A1F002, horses = NA, 
         avg_size = A21001/A1H001, value_implements = A0Q005, pop = A0E001, male = NA, female = NA, 
         enrolled = A4F001, urban_pop = A36001, pct_urban = A36001/A0E001*100, illiterate = A38001, 
         acres_farms = A21001, acres_irrigated = A0G001, pct_irrigated = A0G001/A21001*100) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

#1920

data1920a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds43_1920_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME =  as.character(AREANAME))

data1920b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1920_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1920c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds210_1920_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1920j <- full_join(data1920a, data1920b, by = c("STATEA", "COUNTYA"))
data1920j <- full_join(data1920j, data1920c, by = c("STATEA", "COUNTYA"))
data1920j <- data1920j %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = A8P001+A8P002+A8P003+A8P004, n_farm = AB3P001, cows = NA, horses = NA, 
         avg_size = AB3T001/AB3P001, value_implements = A8P003, pop = A81001, male = A8B001, 
         female = A8B002, enrolled = A7N001+A7N002+A7N003+A7N004, urban_pop = A80001, 
         pct_urban = A80001/A81001*100, illiterate = A7U001, acres_farms = AB3T001, 
         acres_irrigated = AB6L001, pct_irrigated = AB6L002/AB3T001*100) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

#1930
# 1930 has more datasets available than previous years

data1930a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds212_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds52_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930d <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds53_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930e <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds54_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930f <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds55_1930_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1930j <- full_join(data1930a, data1930b, by = c("STATEA", "COUNTYA"))
data1930j <- full_join(data1930j, data1930c, by = c("STATEA", "COUNTYA"))
data1930j <- full_join(data1930j, data1930d, by = c("STATEA", "COUNTYA"))
data1930j <- full_join(data1930j, data1930e, by = c("STATEA", "COUNTYA"))
data1930j <- full_join(data1930j, data1930f, by = c("STATEA", "COUNTYA"))

data1930j <- data1930j %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
  mutate(value_farm = BFD001+BFD002+BFD004+ACEO001, n_farm = AB95001, cows = NA, horses = NA, 
         avg_size = ACAA001, value_implements = BFD004, pop = BCX001, male = BC3001, 
         female = BC3002, enrolled = BDT001+BDT002+BDT003+BDT004, urban_pop = BDX001, 
         pct_urban = BDX001/BDP001*100, illiterate = BDV001, acres_farms = AB98001, 
         acres_irrigated = ACEX002, pct_irrigated = ACEX002/AB98001*100) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# 1940

data1940a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1940_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1940b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds77_1940_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1940c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds78_1940_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1940j <- full_join(data1940a, data1940b, by = c("STATEA", "COUNTYA"))
data1940j <- full_join(data1940j, data1940c, by = c("STATEA", "COUNTYA"))
data1940j <- data1940j %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%

  # value_farm is missing the land & livestock components!!
  # this variable only includings buildings and implements

  mutate(value_farm = BWD001+BWD002, n_farm = BY1001, cows = NA, horses = NA, 
         avg_size = BY0001, value_implements = BWD002, pop = BVU001, male = BVY001, 
         female = BVY002, enrolled = BWU001+BWU002+BWU003+BWU004+BWU005, urban_pop = NA, 
         pct_urban = NA, illiterate = NA, acres_farms = BY3001, 
         acres_irrigated = NA, pct_irrigated = NA) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, 
         pct_illiterate = illiterate/pop) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

# 1950

data1950a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds83_1950_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME =  as.character(AREANAME))

data1950b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds84_1950_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1950j <- full_join(data1950a, data1950b, by = c("STATEA", "COUNTYA")) %>%
  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x, 
         B18001 = as.numeric(B18001)) %>%

  # value_farm is missing the implements & livestock components!!
  # This variable only includings buildings and land

  mutate(value_farm = B3M001, n_farm = B3Z001, cows = NA, horses = NA, 
         avg_size = B3L001, value_implements = NA, pop = B18001, male = B3E001, 
         female = B3E002, enrolled = B2A001+B2A002+B2A003+B2A004+B2A005, urban_pop = B2J001, 
         pct_urban = B2J001/B18001*101, illiterate = NA, acres_farms = B3K001, 
         acres_irrigated = NA, pct_irrigated = NA) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

#1960
# The data are SPARSE yikes

data1960a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds89_1960_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME =  as.character(AREANAME))

data1960b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds90_1960_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1960c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds91_1960_county.csv")) %>%
  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
         AREANAME = as.character(AREANAME))

data1960j <- full_join(data1960a, data1960b, by = c("STATEA", "COUNTYA"))
data1960j <- full_join(data1960j, data1960c, by = c("STATEA", "COUNTYA"))
data1960j <- data1960j %>%
  mutate(value_farm = NA, n_farm = NA, cows = NA, horses = NA, 
         avg_size = NA, value_implements = NA, pop = B47001, male = B5H001, 
         female = B5H002, enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = NA, 
         acres_farms = NA, acres_irrigated = NA, pct_irrigated = NA) %>%
  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)

```

## Join the cleaned sets

```{r boundData}
# Running this code (before I renamed variables) showed that all .x values are the same as all .y values, 
# so it was safe to eliminate the double columns
# as long as problems returned 0 (and it did!)
# this didn't check for .x.x or .y.x. columns, etc.

#problems <- bind_rows(data1900j, data1910j, data1920j, data1930j, data1940j, data1950j, data1960j) %>% 
  #select(STATEA, COUNTYA, GISJOIN.x, GISJOIN.y, YEAR.x, YEAR.y, STATE.x, STATE.y, 
         #COUNTY.x, COUNTY.y) %>%
  #mutate(same = ifelse(GISJOIN.x == GISJOIN.y & YEAR.x == YEAR.y & STATE.x == STATE.y & 
                         #COUNTY.x == COUNTY.y, TRUE, FALSE)) %>%
  #filter(same == "FALSE") %>%
  #nrow()

# Now bind all of the years together!
# Add pct_male and avg_value_farm variables
# Filter for only midwestern states/territories

allData <- as_tibble(bind_rows(data1870j, data1880j, data1890j, 
                       data1900j, data1910j, data1920j, data1930j, data1940j, data1950j, data1960j)) %>%
  mutate(pct_male = male/pop, avg_value_farm = value_farm/n_farm, cows_per_farm = cows_per_person, 
         avg_value_implements = value_implements/n_farm) %>%

  # I mislabeled cows_per_farm as cows_per_person, so I'm doing damage control now

  select(!cows_per_person) %>%
  filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Dakota Territory" | 
           STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | 
           STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | 
           STATE == "Indiana" | STATE == "Ohio")

# This will be used for mapping!
# All relevant columns must be numeric in order to function properly
# Renaming them also helps make the maps more easily interpretable
# But it's not ideal for working with the data in other contexts, hence the separate set

mwData <- allData %>%
  mutate("Value of Farms" = as.numeric(value_farm), "Average Farm Value" = as.numeric(avg_value_farm), 
         "Value of Implements" = as.numeric(value_implements), 
         "Average Implement Value" = as.numeric(avg_value_implements), 
         "Average Farm Size" = as.numeric(avg_size), "Number of Farms" = as.numeric(n_farm), 
         "Acres of Farmland" = as.numeric(acres_farms), "Irrigated Acres" = as.numeric(acres_irrigated), 
         "Percentage of Farmland Irrigated" = as.numeric(pct_irrigated), "Number of Cows" = as.numeric(cows),
         "Number of Cows per Farm" = as.numeric(cows_per_farm), "Number of Horses" = as.numeric(horses), 
         "Total Population" = as.numeric(pop), "Male Population" = as.numeric(male), 
         "Female Population" = as.numeric(female), "Male Percentage of Population" = as.numeric(pct_male),
         "Total School Enrollment" = as.numeric(enrolled), 
         "Percentage School Enrollment" = as.numeric(pct_enrolled), 
         "Urban Population" = as.numeric(urban_pop), 
         "Percent of Population in Urban Areas" = as.numeric(pct_urban), 
         "Illiterate Population" = as.numeric(illiterate), 
         "Illiteracy Rate" = as.numeric(pct_illiterate)) %>%
  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, 
         "Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
                   "Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
                   "Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
                   "Number of Horses", "Total Population", "Male Population", "Female Population", 
                   "Male Percentage of Population", "Total School Enrollment", 
                   "Percentage School Enrollment", "Urban Population", 
                   "Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")

# saving this for later
# will need in shiny app
#sliderOptions <- c("Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
                   #"Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
                   #"Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
                   #"Number of Horses", "Total Population", "Male Population", "Female Population", 
                   #"Male Percentage of Population", "Total School Enrollment", 
                   #"Percentage School Enrollment", "Urban Population", 
                   #"Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")

```

# Shiny 
## Preparation

```{r shapefiles}

# Now save the gis data with short names for sake of the shiny app
# hopefully this will work...

dsn1870 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870"
  layer1870 <- "US_county_1870_conflated"
dsn1880 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1880"
  layer1880 <- "US_county_1880_conflated"
dsn1890 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1890"
  layer1890 <- "US_county_1890_conflated"
dsn1900 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1900"
  layer1900 <- "US_county_1900_conflated"
dsn1910 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1910"
  layer1910 <- "US_county_1910_conflated"
dsn1920 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1920"
  layer1920 <- "US_county_1920_conflated"
dsn1930 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1930"
  layer1930 <- "US_county_1930_conflated"
dsn1940 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1940"
  layer1940 <- "US_county_1940_conflated"
dsn1950 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1950"
  layer1950 <- "US_county_1950_conflated"
dsn1960 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1960"
  layer1960 <- "US_county_1960_conflated"

# Make a tibble of all the shapefiles that you'll need
  shapefiles <- tibble(Year = c(1870, 1880, 1890, 1900, 1910, 1920, 1930, 1940, 1950, 1960), 
                     Dsn = c(dsn1870, dsn1880, dsn1890, dsn1900, dsn1910, dsn1920, dsn1930, 
                             dsn1940, dsn1950, dsn1960),
                     Layer = c(layer1870, layer1880, layer1890, layer1900, layer1910, layer1920, 
                     layer1930, layer1940, layer1950, layer1960))
```

## Output

```{r genericGis}

# df %>% filter(b == !!b)

# gis template for any year/variable desired
# only change these two lines!
year <- 1870
myvar <- "Average Farm Value"

datayear <- mwData %>%
  # use get() to use the stored variable
  # idk why, but it doesn't work without this
  # the variable must be converted to a double in the original mwData dataset
  mutate(myvariable = (get(myvar))) %>%
  filter(YEAR == year, myvariable > 0) %>%
  select(GISJOIN, myvariable)

# Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
# dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
# layer is the filename of the .shp file inside the
# folder dsn points to.

dsnyear <- shapefiles %>% 
  filter(Year == year) %>% 
  pull(Dsn)
layeryear <- shapefiles %>% 
  filter(Year == year) %>% 
  pull(Layer)

countyyear <- sf::st_read(dsn = dsnyear,
                         layer = layeryear)

countyyear <-
  countyyear %>%
  merge(datayear, "GISJOIN")

# Set projection of tracts dataset to `projection` required by leaflet

countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")

# Condense size of data for faster processing

countyyear <- rmapshaper::ms_simplify(countyyear)

# Set palette color

pal <- colorNumeric("viridis", NULL)

#  Plot the data

# use shiny to add a title, which will be be equivalent to myvar

leaflet(countyyear) %>%
  addTiles() %>%
  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
              fillColor = ~pal(myvariable)) %>%
  addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)

```

```{r shiny, eval=FALSE}

# Pasting this here in case my computer explodes

# This is a Shiny web application. You can run the application by clicking
# the 'Run App' button above.
#
# Find out more about building applications with Shiny here:
#
#    http://shiny.rstudio.com/
#

library(shiny)
library(haven)
library(tidyverse)
library(gt)
library(devtools)
library(tibble)
library(foreign)
library(scales)
library(shinythemes)
library(plotly)
library(leaflet)
library(broom)

# packages for mapping
library(leaflet)
library(spdplyr)
library(rmapshaper)
library(sf)

# The real data

data1870a <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds16_1870_county.csv")
data1870b <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds17_1870_county.csv")
data1870j <- full_join(data1870a, data1870b, by = c("GISJOIN", "STATE", "COUNTY"))

data1880a <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds22_1880_county.csv") 
data1880b <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds23_1880_county.csv")
data1880j <- full_join(data1880a, data1880b, by = c("GISJOIN", "STATE", "COUNTY"))

data1890a <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds26_1890_county.csv")
data1890b <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds27_1890_county.csv")
data1890j <- full_join(data1890a, data1890b, by = c("GISJOIN", "STATE", "COUNTY"))

data1800s <- bind_rows(data1870j, data1880j, data1890j)

# I'm not actually using this data anymore
# I just wanted to make sure I could get my plots running smoothly
tbl_1 <- read_por("ICPSR_04254_ag/DS0001/04254-0001-Data.por", col_select = c(NAME, UNFIPS, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_XX_V, IMP_XX_V, AFP_XX_V)) %>%
    rename(name = NAME, fips = UNFIPS, land_in_farms = FML_XX_A, total_area = ARE_XX_A, num_farms = FRM_XX_Q, value_farms = PRP_XX_V, value_implements = IMP_XX_V, value_production_betterments_stocks = AFP_XX_V)

tbl_3 <- read_por("ICPSR_04254_ag/DS0003/04254-0003-Data.por", col_select = c(NAME, UNFIPS, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_XX_V, IMP_XX_V, AFP_XX_V))%>%
    rename(name = NAME, fips = UNFIPS, land_in_farms = FML_XX_A, total_area = ARE_XX_A, num_farms = FRM_XX_Q, value_farms = PRP_XX_V, value_implements = IMP_XX_V, value_production_betterments_stocks = AFP_XX_V)

tbl_5 <- read_por("ICPSR_04254_ag/DS0005/04254-0005-Data.por", col_select = c(NAME, UNFIPS, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_OF_V, IMP_XX_V, CRO_XX_V)) %>%
    rename(name = NAME, fips = UNFIPS, land_in_farms = FML_XX_A, total_area = ARE_XX_A, num_farms = FRM_XX_Q, value_implements = IMP_XX_V, value_crops = CRO_XX_V) %>%
    mutate(value_farms = PRP_OF_V) %>%
    mutate(value_livestock = "NA", value_productsNotFed = "NA") %>%
    select(name, fips, land_in_farms, total_area, num_farms, value_farms, value_implements, value_livestock, value_productsNotFed)

tbl_1_d <- read_por("ICPSR_04296_demog/DS0001/04296-0001-Data.por", col_select = c(NAME, UNFIPS, TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, POP10P, SCH)) %>%
    rename(name = NAME, fips = UNFIPS, total = TTOT, people_over_10 = POP10P, pupils_in_school = SCH) %>%
    mutate(fem_5_24 = FT5_9 + FT10_14 + FT15_19 + FT20_24, mal_5_24 = MT5_9 + MT10_14 + MT15_19 + MT20_24)

tbl_3_d <- read_por("ICPSR_04296_demog/DS0003/04296-0003-Data.por", col_select = c(NAME, UNFIPS, TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH)) %>%
    rename(name = NAME, fips = UNFIPS, total = TTOT, pupils_in_school = SCH) %>%
    mutate(fem_5_24 = FT5_9 + FT10_14 + FT15_19 + FT20_24, mal_5_24 = MT5_9 + MT10_14 + MT15_19 + MT20_24)

tbl_5_d <- read_por("ICPSR_04296_demog/DS0005/04296-0005-Data.por", col_select = c(NAME, UNFIPS, TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH69, SCH1014, SCH1517, SCH1820)) %>% 
    rename(name = NAME, fips = UNFIPS, total = TTOT) %>%
    mutate(fem_5_24 = FT5_9 + FT10_14 + FT15_19 + FT20_24, mal_5_24 = MT5_9 + MT10_14 + MT15_19 + MT20_24,
           # Inconsistent age ranges
           # Not sure how to fix this
           pupils_in_school_6_20 = SCH69 + SCH1014 + SCH1517 + SCH1820)

j_1870 <- full_join(tbl_1, tbl_1_d, by = "fips") %>% 
    mutate(perc_school = pupils_in_school/(fem_5_24 + mal_5_24)*100) %>%
    select(name.x, fips, value_farms, perc_school) %>%
    group_by(fips)

j_1870 <- na.omit(j_1870)

j_1890 <- full_join(tbl_3, tbl_3_d, by = "fips") %>%
    mutate(perc_school = pupils_in_school/(fem_5_24 + mal_5_24)*100) %>%
    select(name.x, fips, value_farms, perc_school) %>%
    group_by(fips)

j_1890 <- na.omit(j_1890)

j_1910 <- full_join(tbl_5, tbl_5_d, by = "fips") %>%
    mutate(perc_school = pupils_in_school_6_20/(fem_5_24 + mal_5_24)*100) %>%
    select(name.x, fips, value_farms, perc_school) %>%
    group_by(fips)

j_1910 <- na.omit(j_1910)

b_1870_1910 <- bind_rows("1870" = j_1870, "1890" = j_1890, "1910" = j_1910, .id = "year") %>%
    group_by(year)

# this should be done by state, not fips!!!
model1 <- lm(data = b_1870_1910, perc_school ~ fips + value_farms) %>% 
    tidy(conf.int = TRUE) %>% 
    gt() %>%
    fmt_number(data = ., columns = c("estimate", "std.error", "statistic", "p.value", "conf.low", "conf.high"), decimals = 2)

# Input

ui <- fluidPage(

    # flatly is a nice theme to make the tabs look put together

    # Application title

    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),

               # remember that tab panel is for what shows up in the tab, and title panel is 
               # for what shows up at the top of the page within the tab

               tabPanel("Tab 1",

                        # Give the page a title
                        titlePanel("Relationship Between Farm Value and Education, by County"),

                        # Generate a row with a sidebar
                        sidebarLayout(

                            # Define the sidebar with one input
                            # insteaad of region, I want year ######
                            sidebarPanel(
                                selectInput("year", "Year:", 
                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
                                            multiple = TRUE)),
                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
                        ),

                        # Create a spot for the plot
                        mainPanel(
                            tabsetPanel(id = "tabsMain",
                                        tabPanel("Plot",

                                                 # this tells the UI what plot to put

                                                 plotOutput("agPlot")  
                                        )))),
               tabPanel("Models",

                        # Give the page a title
                        titlePanel("Relationship Between Farm Value and Education, by County"),
                        p("Make some models! (See notes)"),
                        gt_output("model1")
               ),
               tabPanel("Maps",

                # Give the page a title
                titlePanel("Title Relating to Maps"),
                p("Add a slider for different years/variables & figure out how to get two maps side by side"),
                p("Maps may take up to 30 seconds to load."),
               leafletOutput("map1"),
               p()),

               tabPanel("About",

                        # Give the page a title
                        titlePanel("About this project"),
                        p("My name is Julia Englebert. I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights. Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
                        br(),
                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States. By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century. 
                        I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960. 
                          My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org. IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960. All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
                        br(),
                        p("To see more of my work, please visit my Github account:"),
                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
                        br(),
                        titlePanel("Citations"),
                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles. IPUMS National Historical Geographic Information System: Version 14.0 [Database]. Minneapolis, MN: IPUMS. 2019. http://doi.org/10.18128/D050.V14.0"),
                        p("The College of William and Mary and the Minnesota Population Center. School Attendance Boundary Information System (SABINS): Version 1.0. Minneapolis, MN: University of Minnesota 2011.")
               )
    ))

# use plotly to allow hover text to appear on plots, giving more info about the point

server <- function(input, output) {

    # Fill in the spot we created for a plot
    output$agPlot <- renderPlot({

        # Render a plot
        b_1870_1910 %>% 
            filter(year %in% input$year) %>% 
            group_by(year) %>%
            ggplot(aes(x = value_farms, y = perc_school, color = year)) + 
            geom_point(alpha = 0.5) + 
            ylim(0, 100) + 
            geom_smooth(method = "lm") + 
            labs(title = "Relationship Between Farm Value and Education, by County", 
                 subtitle = "Great Plains Region, United States", 
                 x = "Total Dollar Value of Farms", 
                 y = "Percent Students Ages 5-24 Enrolled in School") +
            scale_x_log10(labels = comma) +
            theme_classic()
    })

    output$map1 <- renderLeaflet({
            data1870 <- data1870j %>%
                filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | STATE == "Indiana" | STATE == "Ohio", AJX001 > 0) %>%
                mutate(valueAllFarmProds1870 = AJX001) %>%
                select(GISJOIN, valueAllFarmProds1870)

            # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
            # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
            # layer is the filename of the .shp file inside the
            # folder dsn points to.

            county1870 <- sf::st_read(dsn = "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870",
                                      layer = "US_county_1870_conflated")

            county1870 <-
                county1870 %>%
                merge(data1870, "GISJOIN")

            # Set projection of tracts dataset to `projection` required by leaflet

            county1870 <- sf::st_transform(county1870, crs="+init=epsg:4326")

            # Condense size of data for faster processing

            county1870 <- rmapshaper::ms_simplify(county1870)

            # Set palette color

            pal <- colorNumeric("viridis", NULL)

            #  Plot the data

            leaflet(county1870) %>%
                addTiles() %>%
                addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
                            fillColor = ~pal(valueAllFarmProds1870)) %>%
                addLegend(pal = pal, values = ~valueAllFarmProds1870, opacity = 1.0, title = "1870 Value of All Farm Productions,\nby Dollar")
    })

    output$model1 <- render_gt({

        # experimental model

        model1

    })
}

# Run the application

shinyApp(ui = ui, server = server)

```

### 2020-04-24 ms_8.Rmd
Commit map() regression

@@
 
+library(shiny)
 library(haven)
@@
 library(scales)
+library(shinythemes)
+library(plotly)
+library(leaflet)
+library(broom)
 
@@
 
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
 
@@
          acres_farms = AOR001, acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
          acres_irrigated = AT2001, pct_irrigated = AT2001/(AUJ001+AUJ002)*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
          acres_farms = AWT001, acres_irrigated = AX8001, pct_irrigated = AX8001/AWT001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
          acres_farms = A21001, acres_irrigated = A0G001, pct_irrigated = A0G001/A21001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
          acres_irrigated = AB6L001, pct_irrigated = AB6L002/AB3T001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
          acres_irrigated = ACEX002, pct_irrigated = ACEX002/AB98001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
          acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, 
-         pct_illiterate = illiterate/pop) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, 
+         pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
          acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
          acres_farms = NA, acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
 ## Output
+### Final
+#### Change
 
-```{r genericGis}
-
-# df %>% filter(b == !!b)
+```{r gisTemplate}
+# might also want to add a filter for state??
+# if time permits...
 
@@
 
+```{r experimentalModels}
+
+modelData <- allData %>%
+  mutate(YEAR = as.factor(YEAR)) %>%
+  filter(! is.na(pct_enrolled) & pct_enrolled > 0) %>%
+  group_by(YEAR)
+
+ggplot(modelData, aes(x = avg_value_implements, y = pct_enrolled, color = YEAR)) + 
+  geom_point(alpha = 0.1) +
+  geom_smooth(method = "lm", se = FALSE) +
+  ylim(0, 100) + 
+  labs(title = "Relationship Between Farm Value and Education, by County",
+       subtitle = "Midwestern United States", 
+       x = "Avergae Value of Farms", 
+       y = "Percent Population Enrolled in School") +
+  scale_x_log10(labels = comma) +
+  theme_classic()
+
+model1 <- modelData %>% 
+  filter(avg_value_farm > 0, avg_value_implements > 0, n_farm > 0) %>%
+  lm(data = ., pct_enrolled ~ avg_value_farm + pct_urban + n_farm) %>% 
+  tidy(conf.int = TRUE) %>% 
+  gt() %>% 
+  fmt_number(data = ., columns = c("estimate", "std.error", "statistic", "p.value", "conf.low", "conf.high"), decimals = 2)
+
+# avg_value_implements and cows_per_farm have an effect on enrollment
+# in general, for 1879 and 1890, more cows per farm = less enrollment
+# aka bigger farms = less enrollment?
+model2a <- modelData %>% 
+  filter(YEAR == "1870") %>%
+  lm(data = ., pct_enrolled ~ cows_per_farm) %>% 
+  tidy(conf.int = TRUE) %>% 
+  select(term, estimate, p.value, conf.low, conf.high) %>%
+  gt() %>% 
+  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
+             decimals = 2)
+
+# modelData$YEAR <- fct_explicit_na(modelData$YEAR)
+
+model2b <- modelData %>% 
+  filter(! is.na(cows_per_farm)) %>%
+  lm(data = ., pct_enrolled ~ cows_per_farm) %>% 
+  tidy(conf.int = TRUE) %>% 
+  select(term, estimate, p.value, conf.low, conf.high) %>%
+  gt() %>% 
+  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
+             decimals = 2)
+
+model2c <- modelData %>% 
+  filter(YEAR == "1910") %>%
+  lm(data = ., pct_enrolled ~ cows_per_farm) %>% 
+  tidy(conf.int = TRUE) %>% 
+  select(term, estimate, p.value, conf.low, conf.high) %>%
+  gt() %>% 
+  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
+             decimals = 2)
+
+model2
+# cow plot
+# incorporate
+cowplot <- modelData %>% 
+  filter(YEAR == "1870" | YEAR == "1890" | YEAR == "1910") %>%
+  ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
+  geom_point(alpha = 0.5) +
+  geom_smooth(method = "lm", se = FALSE) +
+  scale_x_log10(labels = comma) +
+  labs(title = "Cows and Education",
+       subtitle = "Relationship between the number of cows per farm\nand school enrollment, by county",
+       x = "Average Cows per Farm",
+       y = "Percent of Population Enrolled in School") +
+  theme_classic()
+
+modelData %>% 
+  filter(YEAR == "1870" | YEAR == "1890" | YEAR == "1910") %>%
+  ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
+  geom_point() +
+  geom_smooth(se = FALSE)
+
+# Enrollment/Illiteracy model (just for testing)
+
+# This is saying that for every 1% increase in illiterate people, there is a 0.24% 
+# decrease in school enrollment, which makes sense
+model3 <- modelData %>%
+  lm(data = ., pct_enrolled ~ pct_illiterate) %>%
+  tidy(conf.int = TRUE) %>%
+  select(term, estimate, p.value, conf.low, conf.high) %>%
+  gt() %>%
+  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
+             decimals = 2)
+
+mod3plot <- modelData %>%
+  filter(! is.na(pct_enrolled) & ! is.na(pct_illiterate)) %>%
+  ggplot(aes(pct_enrolled, pct_illiterate)) + 
+  geom_point() + 
+  geom_smooth(se = FALSE)
+```
+
+```{r moreModels}
+
+# can only use percentages until adjusting for inflation
+# OR deal with each year independently
+
+# Set the NA values to 0
+# Not ideal, but necessary for running regressions
+allData[is.na(allData)] <- 0
+
+reg_years <- allData %>% 
+  # Get rid of the infinite cows_per_farm values
+  filter(! is.infinite(cows_per_farm)) %>%
+  group_by(YEAR) %>%
+  arrange(desc(YEAR)) %>%
+
+  # Nest to create a tibble of tibbles
+
+  nest() %>%
+
+  # Call map inside of mutate to add new columns to the tibble
+  # reg_results are the results of the regression we ran in question 2
+  # coef is the coefficient of log_migrants for each continent
+  # upper and lower are the upper and lower bounds of the confidence interval created by tidy()
+
+  # options
+  #STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         # acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_farm, horses, pop, male, female, 
+        # enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate
+
+mutate(mod = map(data, ~ lm(pct_enrolled ~ avg_size + avg_value_implements + n_farm + pct_irrigated + cows_per_farm + pct_urban, data = .)),
+         reg_results = map(mod, ~ tidy(., conf.int = TRUE)),
+         coef = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% pull(estimate)),
+         upper = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% pull(conf.high)),
+         lower = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% pull(conf.low)),
+         coef = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% pull(estimate)),
+         upper = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% pull(conf.high)),
+         lower = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% pull(conf.low)),
+       coef = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(estimate)),
+         upper = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(conf.high)),
+         lower = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(conf.low)),
+       coef = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(estimate)),
+         upper = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(conf.high)),
+         lower = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(conf.low)),
+       coef = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(estimate)),
+         upper = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(conf.high)),
+         lower = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(conf.low)),
+       coef = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(estimate)),
+         upper = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(conf.high)),
+         lower = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(conf.low))
+         )
+
+# Now use the resulting tibble to create the gt table
+
+q4_gt <- q4 %>%
+
+  # The original year column was giving me problems because it's a grouping variable
+  # Ungrouping solved the problem
+
+  ungroup() %>%
+
+  # Select the columns needed, as per the instructions
+
+  select(year, coef, lower, upper) %>%
+
+  gt() %>%
+
+  # Add the desired titles, subtitles, and column labels
+
+  tab_header(title = "Effect of Number of Migrants on Amount of Remittances by Year",
+             subtitle = "With 95% Confidence Interval") %>%
+  cols_label(year = "Year", coef = "Migrant Coefficient", 
+             lower = "Lower Bound", upper = "Upper Bound") %>%
+
+  # Round all values to two decimal places
+
+  fmt_number(columns = vars(coef, lower, upper), decimals = 2)
+
+```
+
+### SHINY FINAL
+```{r shiny, eval=FALSE}
+
+# PASTE DATA and LIBRARIES here after
+
+# Input
+
+ui <- fluidPage(
+
+    # flatly is a nice theme to make the tabs look put together
+
+    # Application title
+
+    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),
+
+               # remember that tab panel is for what shows up in the tab, and title panel is 
+               # for what shows up at the top of the page within the tab
+
+               tabPanel("Tab 1",
+
+                        # Give the page a title
+                        titlePanel("Relationship Between Farm Value and Education, by County"),
+
+                        # Generate a row with a sidebar
+                        sidebarLayout(
+
+                            # Define the sidebar with one input
+                            # insteaad of region, I want year ######
+                            sidebarPanel(
+                                selectInput("year", "Year:", 
+                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
+                                            multiple = TRUE)),
+                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
+                        ),
+
+                        # Create a spot for the plot
+                        mainPanel(
+                            tabsetPanel(id = "tabsMain",
+                                        tabPanel("Plot",
+
+                                                 # this tells the UI what plot to put
+
+                                                 plotOutput("agPlot")  
+                                        )))),
+               tabPanel("Models",
+
+                        # Give the page a title
+                        titlePanel("Relationship Between Farm Value and Education, by County"),
+                        p("Make some models! (See notes)"),
+                        gt_output("model1")
+               ),
+               tabPanel("Maps",
+
+                # Give the page a title
+                titlePanel("Title Relating to Maps"),
+                p("Add a slider for different years/variables & figure out how to get two maps side by side"),
+                p("Maps may take up to 30 seconds to load."),
+               leafletOutput("map1"),
+               p()),
+
+               tabPanel("About",
+
+                        # Give the page a title
+                        titlePanel("About this project"),
+                        p("My name is Julia Englebert.
+I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights.
+Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
+                        br(),
+                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States.
+By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century.
+                        I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960.
+                          My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org.
+IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960.
+All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
+                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
+                        br(),
+                        p("To see more of my work, please visit my Github account:"),
+                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
+                        br(),
+                        titlePanel("Citations"),
+                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles.
+IPUMS National Historical Geographic Information System: Version 14.0 [Database].
+Minneapolis, MN: IPUMS.
+2019.
+http://doi.org/10.18128/D050.V14.0"),
+                        p("The College of William and Mary and the Minnesota Population Center.
+School Attendance Boundary Information System (SABINS): Version 1.0.
+Minneapolis, MN: University of Minnesota 2011.")
+               )
+    ))
+
+# use plotly to allow hover text to appear on plots, giving more info about the point
+
+server <- function(input, output) {
+
+    # Fill in the spot we created for a plot
+    output$agPlot <- renderPlot({
+
+        # Render a plot
+        b_1870_1910 %>% 
+            filter(year %in% input$year) %>% 
+            group_by(year) %>%
+            ggplot(aes(x = value_farms, y = perc_school, color = year)) + 
+            geom_point(alpha = 0.5) + 
+            ylim(0, 100) + 
+            geom_smooth(method = "lm") + 
+            labs(title = "Relationship Between Farm Value and Education, by County", 
+                 subtitle = "Great Plains Region, United States", 
+                 x = "Total Dollar Value of Farms", 
+                 y = "Percent Students Ages 5-24 Enrolled in School") +
+            scale_x_log10(labels = comma) +
+            theme_classic()
+    })
+
+    output$map1 <- renderLeaflet({
+            data1870 <- data1870j %>%
+                filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | STATE == "Indiana" | STATE == "Ohio", AJX001 > 0) %>%
+                mutate(valueAllFarmProds1870 = AJX001) %>%
+                select(GISJOIN, valueAllFarmProds1870)
+
+            # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
+            # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
+            # layer is the filename of the .shp file inside the
+            # folder dsn points to.
+
+            county1870 <- sf::st_read(dsn = "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870",
+                                      layer = "US_county_1870_conflated")
+
+            county1870 <-
+                county1870 %>%
+                merge(data1870, "GISJOIN")
+
+            # Set projection of tracts dataset to `projection` required by leaflet
+
+            county1870 <- sf::st_transform(county1870, crs="+init=epsg:4326")
+
+            # Condense size of data for faster processing
+
+            county1870 <- rmapshaper::ms_simplify(county1870)
+
+            # Set palette color
+
+            pal <- colorNumeric("viridis", NULL)
+
+            #  Plot the data
+
+            leaflet(county1870) %>%
+                addTiles() %>%
+                addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
+                            fillColor = ~pal(valueAllFarmProds1870)) %>%
+                addLegend(pal = pal, values = ~valueAllFarmProds1870, opacity = 1.0, title = "1870 Value of All Farm Productions,\nby Dollar")
+    })
+
+    output$model1 <- render_gt({
+
+        # experimental model
+
+        model1
+
+    })
+}
+
+# Run the application
+
+shinyApp(ui = ui, server = server)
+```
+
+# DO NOT MAKE CHANGES BELOW THIS LINE
+
+### Map Code
+
+```{r genericGis}
+
+# DO NOT CHANGE THIS
+
+# might also want to add a filter for state??
+# if time permits...
+
+# gis template for any year/variable desired
+# only change these two lines!
+year <- 1870
+myvar <- "Average Farm Value"
+
+datayear <- mwData %>%
+  # use get() to use the stored variable
+  # idk why, but it doesn't work without this
+  # the variable must be converted to a double in the original mwData dataset
+  mutate(myvariable = (get(myvar))) %>%
+  filter(YEAR == year, myvariable > 0) %>%
+  select(GISJOIN, myvariable)
+
+# Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
+# dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
+# layer is the filename of the .shp file inside the
+# folder dsn points to.
+
+dsnyear <- shapefiles %>% 
+  filter(Year == year) %>% 
+  pull(Dsn)
+layeryear <- shapefiles %>% 
+  filter(Year == year) %>% 
+  pull(Layer)
+
+countyyear <- sf::st_read(dsn = dsnyear,
+                         layer = layeryear)
+
+countyyear <-
+  countyyear %>%
+  merge(datayear, "GISJOIN")
+
+# Set projection of tracts dataset to `projection` required by leaflet
+
+countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
+
+# Condense size of data for faster processing
+
+countyyear <- rmapshaper::ms_simplify(countyyear)
+
+# Set palette color
+
+pal <- colorNumeric("viridis", NULL)
+
+#  Plot the data
+
+# use shiny to add a title, which will be be equivalent to myvar
+
+leaflet(countyyear) %>%
+  addTiles() %>%
+  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
+              fillColor = ~pal(myvariable)) %>%
+  addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
+
+```
+
+# Save This
 ```{r shiny, eval=FALSE}

### 2020-04-25 ms_8.Rmd
Commit some models (app probably not running now)

@@
 library(broom)
+library(readxl)
 
@@
 
-```{r allData}
+```{r allData, cache=TRUE}
 # 1800s data
@@
   mutate(value_farm = B3M001, n_farm = B3Z001, cows = NA, horses = NA, 
-         avg_size = B3L001, value_implements = NA, pop = B18001, male = B3E001, 
+         avg_size = B3L001, value_implements = NA, pop = B13001, male = B3E001, 
          female = B3E002, enrolled = B2A001+B2A002+B2A003+B2A004+B2A005, urban_pop = B2J001, 
@@
 
-```{r boundData}
+```{r boundData, cache=TRUE}
 # Running this code (before I renamed variables) showed that all .x values are the same as all .y values, 
@@
 
-allData <- as_tibble(bind_rows(data1870j, data1880j, data1890j, 
+allData_inflated <- as_tibble(bind_rows(data1870j, data1880j, data1890j, 
                        data1900j, data1910j, data1920j, data1930j, data1940j, data1950j, data1960j)) %>%
-  mutate(pct_male = male/pop, avg_value_farm = value_farm/n_farm, cows_per_farm = cows_per_person, 
+  mutate(pct_male = male/pop*100, avg_value_farm = value_farm/n_farm, cows_per_farm = cows_per_person, 
          avg_value_implements = value_implements/n_farm) %>%
@@
            STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | 
-           STATE == "Indiana" | STATE == "Ohio")
-
-# This will be used for mapping!
+           STATE == "Indiana" | STATE == "Ohio",
+
+         # Get rid of the weird infinite values in this column
+
+         ! is.infinite(cows_per_farm))
+
+# Now adjust everything for inflation
+# I made a spreadsheet based on this website: 
+# https://www.officialdata.org/us/inflation/1960?amount=1
+
+inflation <- read_excel("inflation.xlsx")
+
+clean_data <- left_join(allData_inflated, inflation, by = c("YEAR" = "year")) %>%
+  # Use the multiplier column to adjust for inflation
+  mutate(value_farm = value_farm*multiplier, value_implements = value_implements*multiplier) %>%
+  mutate(avg_value_farm = value_farm/n_farm, avg_value_implements = value_implements/n_farm)
+
+# Now save as an RDS file to save on space
+
+saveRDS(clean_data, file = "final/my_data.rds")
+saveRDS(clean_data, file = "my_data.rds")
+
+```
+
+```{r mapDataset}
+
+# Now just read in the RDS file
+# This is all you need to copy into the shiny app
+
+allData <- readRDS(file = "my_data.rds")
+
+# this is workable data, where NA columns have been removed
+# There was too much variable variation & inconsistencies to draw conclusions from all years
+
+subData <- allData %>% 
+  filter(YEAR == "1870" | YEAR == "1890" | YEAR == "1910") %>%
+  select(-acres_irrigated, -pct_irrigated, -illiterate, -pct_illiterate, -male, -female, -pct_male)
+
+# This next bitwill be used for mapping only!
 # All relevant columns must be numeric in order to function properly
@@
 
+#choices for myvar will be determined by: 
+appPlan <- allData %>%
+  filter(YEAR == "1870") %>%
+  select_if(~ !all(is.na(.))) %>%
+  select(-STATEA:-COUNTY) %>%
+  select(-multiplier)
+
 datayear <- mwData %>%
@@
 ```{r experimentalModels}
-
-modelData <- allData %>%
+# USE subData FOR ALL MODELS
+# EVERYTHING ELSE IS DUBIOUS
+
+# Incorporate
+# Value of implements functions as a stand in for mechanization
+impedu_plot <- allData %>%
   mutate(YEAR = as.factor(YEAR)) %>%
-  filter(! is.na(pct_enrolled) & pct_enrolled > 0) %>%
-  group_by(YEAR)
-
-ggplot(modelData, aes(x = avg_value_implements, y = pct_enrolled, color = YEAR)) + 
+  group_by(YEAR) %>%
+  select(YEAR, avg_value_implements, pct_enrolled) %>%
+  na.omit() %>%
+  ggplot(aes(x = avg_value_implements, y = pct_enrolled, color = YEAR)) + 
   geom_point(alpha = 0.1) +
   geom_smooth(method = "lm", se = FALSE) +
-  ylim(0, 100) + 
+  ylim(0, 45) + 
   labs(title = "Relationship Between Farm Value and Education, by County",
        subtitle = "Midwestern United States", 
-       x = "Avergae Value of Farms", 
+       x = "Average Value of Implements", 
        y = "Percent Population Enrolled in School") +
@@
 
-model1 <- modelData %>% 
-  filter(avg_value_farm > 0, avg_value_implements > 0, n_farm > 0) %>%
-  lm(data = ., pct_enrolled ~ avg_value_farm + pct_urban + n_farm) %>% 
-  tidy(conf.int = TRUE) %>% 
+# Incorporate
+# States are gaining more implements each year
+imp_plot <- allData %>%
+  mutate(YEAR = as.factor(YEAR)) %>%
+  select(YEAR, STATE, value_implements) %>%
+  na.omit() %>%
+  group_by(YEAR, STATE) %>%
+  summarize(mean_imp = mean(value_implements)) %>% 
+  ungroup() %>%
+  group_by(STATE) %>%
+  ggplot(aes(x = YEAR, y = mean_imp, fill = STATE)) + 
+  geom_col() +
+  labs(title = "Implement Value by Year, by State",
+       subtitle = "Dollar Values Adjusted for Inflation",
+       x = "Year",
+       y = "Average Value of All Implements",
+       fill = "State") +
+  theme_classic()
+
+# INCORPORATE
+# having more farms and being male leads to lower school enrollment
+# supports my hypothesis of farm work affecting enrollment, since males might be expected to due more manual labor
+maleedu_model <- allData %>%
+  select(YEAR, pct_enrolled, n_farm, pct_male) %>%
+  na.omit() %>%
+  group_by(YEAR) %>%
+  nest() %>%
+
+  # Call map inside of mutate to add new columns to the tibble
+  # upper and lower are the upper and lower bounds of the confidence interval created by tidy()
+
+  # options
+  #STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         # acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_farm, horses, pop, male, female, 
+        # enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate
+
+  mutate(mod = map(data, ~ lm(pct_enrolled ~ n_farm*pct_male, data = .)),
+         reg_results = map(mod, ~ tidy(., conf.int = TRUE)),
+         n_farm_coef = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% 
+                                   pull(estimate)),
+         n_farm_upper = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% 
+                                    pull(conf.high)),
+         n_farm_lower = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% 
+                                    pull(conf.low)),
+         pct_male_coef = map_dbl(reg_results, ~ filter(., term == "pct_male") %>% 
+                                   pull(estimate)),
+         pct_male_upper = map_dbl(reg_results, ~ filter(., term == "pct_male") %>% 
+                                    pull(conf.high)),
+         pct_male_lower = map_dbl(reg_results, ~ filter(., term == "pct_male") %>% 
+                                    pull(conf.low)),
+         interaction_coef = map_dbl(reg_results, ~ filter(., term == "n_farm:pct_male") %>% 
+                                   pull(estimate)),
+         interaction_upper = map_dbl(reg_results, ~ filter(., term == "n_farm:pct_male") %>% 
+                                    pull(conf.high)),
+         interaction_lower = map_dbl(reg_results, ~ filter(., term == "n_farm:pct_male") %>% 
+                                    pull(conf.low))
+         )
+# Incorporate
+maleedu_gt <- maleedu_model %>%
+  ungroup() %>%
+
+  # Select the columns needed, as per the instructions
+
+  select(-data:-reg_results) %>%
+
   gt() %>%
-  fmt_number(data = ., columns = c("estimate", "std.error", "statistic", "p.value", "conf.low", "conf.high"), decimals = 2)
+
+  # Add the desired titles, subtitles, and column labels
+
+  tab_header(title = "Effect of Gender and Number of Farms on Enrollment",
+             subtitle = "With 95% Confidence Interval") %>%
+  tab_spanner(label = "Farm Coefficients", columns = vars(n_farm_lower, n_farm_coef, n_farm_upper)) %>%
+  tab_spanner(label = "Percent Male Coefficients", columns = vars(pct_male_lower, pct_male_coef,pct_male_upper)) %>%
+  tab_spanner(label = "Interaction Coefficients", columns = vars(interaction_lower, interaction_coef, interaction_upper)) %>%
+  cols_label(YEAR = "Year", n_farm_coef	= "Estimate", n_farm_upper = "Upper", n_farm_lower = "Lower", pct_male_coef = "Estimate", pct_male_upper = "Upper", pct_male_lower = "Lower",	interaction_coef = "Estimate", interaction_upper = "Upper", interaction_lower = "Lower") %>%
+  # Round all values to two decimal places
+
+  fmt_number(columns = vars(n_farm_lower, n_farm_coef, n_farm_upper, pct_male_lower, pct_male_coef,pct_male_upper, interaction_lower, interaction_coef, interaction_upper), decimals = 4)
+
+# INCORPORATE
+# Effect of gender on enrollment dwindles over time
+maleedu_plot <- maleedu_model %>%
+  ungroup() %>%
+  select(YEAR, pct_male_coef, pct_male_upper, pct_male_lower) %>%
+  mutate(YEAR = as.factor(YEAR)) %>%
+  group_by(YEAR) %>%
+  ggplot(aes(x = YEAR, y = pct_male_coef)) + 
+  geom_errorbar(ymin = maleedu_model$pct_male_lower, ymax = maleedu_model$pct_male_upper) + 
+  geom_point() +
+  theme_classic() +
+  ylim(-1.5, 0.5) + 
+  labs(x = "Year",
+       y = "Coefficient",
+       title = "Percent Male Coefficients Over Time",
+       subtitle = "Measuring the effect of being male on school enrollment")
 
@@
 
-# modelData$YEAR <- fct_explicit_na(modelData$YEAR)
-
-model2b <- modelData %>% 
-  filter(! is.na(cows_per_farm)) %>%
+# there was an infinite value left in the cows_per_farm column for this year
+# not sure why, but I got rid of it
+model2a <- modelData %>% 
+  filter(YEAR == "1890", !is.infinite(cows_per_farm)) %>%
   lm(data = ., pct_enrolled ~ cows_per_farm) %>% 
@@
 
-# Set the NA values to 0
+# Set the NA values to almost zero
 # Not ideal, but necessary for running regressions
-allData[is.na(allData)] <- 0
+# setting to zero returns NaNs
 
 reg_years <- allData %>% 
+  # sample the data to run faster
+  # change this later
+  sample_n(size = 5000) %>%
   # Get rid of the infinite cows_per_farm values
-  filter(! is.infinite(cows_per_farm)) %>%
+  filter(! (is.na(pct_enrolled) & is.na(avg_size))) %>%
   group_by(YEAR) %>%
   arrange(desc(YEAR)) %>% 
-
-  # Nest to create a tibble of tibbles
-
-  nest() %>%
+  nest()
 
@@
         # enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate
-
+reg_years1 <- reg_years %>%
   mutate(mod = map(data, ~ lm(pct_enrolled ~ avg_size + avg_value_implements + n_farm + pct_irrigated + cows_per_farm + pct_urban, data = .)),
          reg_results = map(mod, ~ tidy(., conf.int = TRUE)),
-         coef = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% pull(estimate)),
-         upper = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% pull(conf.high)),
-         lower = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% pull(conf.low)),
-         coef = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% pull(estimate)),
-         upper = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% pull(conf.high)),
-         lower = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% pull(conf.low)),
-       coef = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(estimate)),
-         upper = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(conf.high)),
-         lower = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(conf.low)),
-       coef = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(estimate)),
-         upper = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(conf.high)),
-         lower = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(conf.low)),
-       coef = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(estimate)),
-         upper = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(conf.high)),
-         lower = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(conf.low)),
-       coef = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(estimate)),
-         upper = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(conf.high)),
-         lower = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(conf.low))
+         avg_size_coef = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% 
+                                   pull(estimate)),
+         avg_size_upper = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% 
+                                    pull(conf.high)),
+         avg_size_lower = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% 
+                                    pull(conf.low)),
+         avg_value_implements_coef = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% 
+                                               pull(estimate)),
+         avg_value_implements_upper = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% 
+                                                pull(conf.high)),
+         avg_value_implements_lower = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% 
+                                                pull(conf.low)),
+         n_farm_coef = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(estimate)),
+         n_farm_upper = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(conf.high)),
+         n_farm_lower = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(conf.low)),
+         pct_irrigated_coef = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(estimate)),
+         pct_irrigated_upper = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(conf.high)),
+         pct_irrigated_lower = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(conf.low)),
+         cows_per_farm_coef = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(estimate)),
+         cows_per_farm_upper = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(conf.high)),
+         cows_per_farm_lower = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(conf.low)),
+         pct_urban_coef = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(estimate)),
+         pct_urban_upper = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(conf.high)),
+         pct_urban_lower = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(conf.low))
   )
@@
 
-  select(year, coef, lower, upper) %>%
+  select(YEAR, coef, lower, upper) %>%
 
@@
 
-  tab_header(title = "Effect of Number of Migrants on Amount of Remittances by Year",
+  tab_header(title = "Title",
              subtitle = "With 95% Confidence Interval") %>%
-  cols_label(year = "Year", coef = "Migrant Coefficient", 
+  cols_label(YEAR = "Year", coef = "Migrant Coefficient", 
              lower = "Lower Bound", upper = "Upper Bound") %>%

### 2020-04-25 scratch-save.Rmd
Initial commit

--- 2020-04-24 ms_8.Rmd
@@
 ---
-title: "Maps"
+title: "scratch-save"
 author: "Julia Englebert"
-date: "4/22/2020"
+date: "4/25/2020"
 output: html_document
@@
 knitr::opts_chunk$set(echo = TRUE)
-
-options(scipen = 999)
-
-library(haven)
-library(tidyverse)
-library(gt)
-library(devtools)
-library(tibble)
-library(foreign)
-library(scales)
-
-# packages for mapping
-library(leaflet)
-library(spdplyr)
-library(rmapshaper)
-library(sf)
 ```
 
-```{r testModel}
-#
-# model1 <- lm(data = mwdata, perc_school ~ state + land_in_farms)
-```
+# DO NOT MAKE CHANGES BELOW THIS LINE
 
-# DATA WRANGLING
-
-```{r allData}
-# 1800s data
-
-# Will comment on 1870s data, then repeat the process for all subsequent years
-# Note that the codenames ARE DIFFERENT FOR EACH YEAR
-# Even if the variable was measured in the same way
-# See original codebook for details
-
-# Read in all 1870 csv files
-# Originally, R was returning this message:
-# Column `STATE` joining factors with different levels, coercing to character vectorColumn `COUNTY` joining
-# factors with different levels, coercing to character vector
-# I got rid of it by converting all factors to characters
-
-data1870a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds16_1870_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1870b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds17_1870_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-# Join the two datasets by STATEA and COUNTYA, which are the state and county codes
-
-data1870j <- full_join(data1870a, data1870b, by = c("STATEA", "COUNTYA")) %>%
-
-  # Mutate to get rid of the .x suffix on variables
-
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-
-  # Mutate to rename varialbes of interest to uniform, comprehensible names
-  # Some existing varialbes are combined to create the desired variables
-  # See codebook
-
-  mutate(value_farm = AJV001, n_farm = AJS001, cows =  AJ0003+AJ0005, horses = AJ0001, 
-         avg_size = (AJT001*2 + AJT002*6 + AJT003*15 + AJT004*35 + AJT005*75 + AJT006*300 + AJT007*750 + 
-                       AJT008*1200)/AKP001, value_implements = AKL001, pop = AJ3001, male = AKD001, 
-         female = AKD002, enrolled = ALG001, urban_pop = AKE001, pct_urban = AKE001/AJ3001*100, 
-         illiterate = AJ7001, acres_farms = AJU001+ AJU002+AJU003,
-
-         # Even though irrigation data doesn't exist for this year, I'm filling the columns in with NA
-         # Later years will have irrigation data
-
-         acres_irrigated = NA, pct_irrigated = NA) %>%
-
-  # Mutate again to make new variables out of the ones you just created in the last mutate
-
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
-
-  # Select the desired variables
-
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm,
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female,
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# Now the dataset is much more manageable!
-
-# Start on the next year, 1880, repeating the exact same process
-# Lots of NA values due to scant data this year
-
-data1880a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds22_1880_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1880b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds23_1880_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1880j <- full_join(data1880a, data1880b, by = c("STATEA", "COUNTYA")) %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = APN001+APO001+APQ001, n_farm = AOQ001, cows =  AOG004+AOG005, horses = AOG001, 
-         avg_size = APG001, value_implements = APO001, pop = AOT001, male = AP1001, female = AP1002, 
-         enrolled = NA, urban_pop = AO4001, pct_urban = AO4001/AOT001*100, illiterate = NA, 
-         acres_farms = AOR001, acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# Start on 1890 data
-
-data1890a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds26_1890_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1890b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds27_1890_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1890j <- full_join(data1890a, data1890b, by = c("STATEA", "COUNTYA")) %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = AUK001+AUK002+AUK003, n_farm = AS9001, cows = ASY002, horses = ASY003, 
-         avg_size = ATW001, value_implements = AUK002, pop = AUM001, male = AV0001+AV0003+AV0005+AV0007, 
-         female = AV0002+AV0004+AV0006+AV0008, enrolled = AVK005+AVK006+AVK007+AVK008, urban_pop = AUU001, 
-         pct_urban = AUU001/AUM001*100, illiterate = NA, acres_farms = AUJ001+AUJ002, 
-         acres_irrigated = AT2001, pct_irrigated = AT2001/(AUJ001+AUJ002)*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# Before I mutated all of the data, running this code showed that all .x values are the same 
-# as all .y values, so it was safe to eliminate the double columns as long as "problems" returned 0
-# And it did!
-
-#problems <- bind_rows(data1870j, data1880j, data1890j) %>% 
-  #select(STATEA, COUNTYA, GISJOIN.x, GISJOIN.y, YEAR.x, YEAR.y, STATE.x, STATE.y, 
-         #COUNTY.x, COUNTY.y, AREANAME.x, AREANAME.y) %>%
-  #mutate(same = ifelse(GISJOIN.x == GISJOIN.y & YEAR.x == YEAR.y & STATE.x == STATE.y & 
-                         #COUNTY.x == COUNTY.y & AREANAME.x == AREANAME.y, TRUE, FALSE)) %>%
-  #filter(same == "FALSE") %>%
-  #nrow()
-
-# Read in the 1900s data
-
-# Only do decennial census data
-# Intermittent years only have AF12001: estimated population
-
-# 1900
-
-data1900a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds30_1900_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1900b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds31_1900_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1900j <- full_join(data1900a, data1900b, by = c("STATEA", "COUNTYA")) %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = AWV001, n_farm = AW3001, cows = AW7001, horses = AW7002, 
-         avg_size = AXE001, value_implements = AWW003, pop = AWS001, male = AZ0001,female = AZ0001, 
-         enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = AYS001+AYS002+AYS003+AYS004+AYS005, 
-         acres_farms = AWT001, acres_irrigated = AX8001, pct_irrigated = AX8001/AWT001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# 1910
-
-data1910a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds36_1910_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1910b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds37_1910_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1910j <- full_join(data1910a, data1910b, by = c("STATEA", "COUNTYA")) %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = A0F001, n_farm = A1H001, cows = A1F001+A1F002, horses = NA, 
-         avg_size = A21001/A1H001, value_implements = A0Q005, pop = A0E001, male = NA, female = NA, 
-         enrolled = A4F001, urban_pop = A36001, pct_urban = A36001/A0E001*100, illiterate = A38001, 
-         acres_farms = A21001, acres_irrigated = A0G001, pct_irrigated = A0G001/A21001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-#1920
-
-data1920a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds43_1920_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME =  as.character(AREANAME))
-
-data1920b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1920_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1920c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds210_1920_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1920j <- full_join(data1920a, data1920b, by = c("STATEA", "COUNTYA"))
-data1920j <- full_join(data1920j, data1920c, by = c("STATEA", "COUNTYA"))
-data1920j <- data1920j %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = A8P001+A8P002+A8P003+A8P004, n_farm = AB3P001, cows = NA, horses = NA, 
-         avg_size = AB3T001/AB3P001, value_implements = A8P003, pop = A81001, male = A8B001, 
-         female = A8B002, enrolled = A7N001+A7N002+A7N003+A7N004, urban_pop = A80001, 
-         pct_urban = A80001/A81001*100, illiterate = A7U001, acres_farms = AB3T001, 
-         acres_irrigated = AB6L001, pct_irrigated = AB6L002/AB3T001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-#1930
-# 1930 has more datasets available than previous years
-
-data1930a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds212_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds52_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930d <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds53_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930e <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds54_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930f <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds55_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930j <- full_join(data1930a, data1930b, by = c("STATEA", "COUNTYA"))
-data1930j <- full_join(data1930j, data1930c, by = c("STATEA", "COUNTYA"))
-data1930j <- full_join(data1930j, data1930d, by = c("STATEA", "COUNTYA"))
-data1930j <- full_join(data1930j, data1930e, by = c("STATEA", "COUNTYA"))
-data1930j <- full_join(data1930j, data1930f, by = c("STATEA", "COUNTYA"))
-
-data1930j <- data1930j %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = BFD001+BFD002+BFD004+ACEO001, n_farm = AB95001, cows = NA, horses = NA, 
-         avg_size = ACAA001, value_implements = BFD004, pop = BCX001, male = BC3001, 
-         female = BC3002, enrolled = BDT001+BDT002+BDT003+BDT004, urban_pop = BDX001, 
-         pct_urban = BDX001/BDP001*100, illiterate = BDV001, acres_farms = AB98001, 
-         acres_irrigated = ACEX002, pct_irrigated = ACEX002/AB98001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# 1940
-
-data1940a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1940_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1940b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds77_1940_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1940c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds78_1940_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1940j <- full_join(data1940a, data1940b, by = c("STATEA", "COUNTYA"))
-data1940j <- full_join(data1940j, data1940c, by = c("STATEA", "COUNTYA"))
-data1940j <- data1940j %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-
-  # value_farm is missing the land & livestock components!!
-  # this variable only includings buildings and implements
-
-  mutate(value_farm = BWD001+BWD002, n_farm = BY1001, cows = NA, horses = NA, 
-         avg_size = BY0001, value_implements = BWD002, pop = BVU001, male = BVY001, 
-         female = BVY002, enrolled = BWU001+BWU002+BWU003+BWU004+BWU005, urban_pop = NA, 
-         pct_urban = NA, illiterate = NA, acres_farms = BY3001, 
-         acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, 
-         pct_illiterate = illiterate/pop) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# 1950
-
-data1950a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds83_1950_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME =  as.character(AREANAME))
-
-data1950b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds84_1950_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1950j <- full_join(data1950a, data1950b, by = c("STATEA", "COUNTYA")) %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x, 
-         B18001 = as.numeric(B18001)) %>%
-
-  # value_farm is missing the implements & livestock components!!
-  # This variable only includings buildings and land
-
-  mutate(value_farm = B3M001, n_farm = B3Z001, cows = NA, horses = NA, 
-         avg_size = B3L001, value_implements = NA, pop = B18001, male = B3E001, 
-         female = B3E002, enrolled = B2A001+B2A002+B2A003+B2A004+B2A005, urban_pop = B2J001, 
-         pct_urban = B2J001/B18001*101, illiterate = NA, acres_farms = B3K001, 
-         acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-#1960
-# The data are SPARSE yikes
-
-data1960a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds89_1960_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME =  as.character(AREANAME))
-
-data1960b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds90_1960_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1960c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds91_1960_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1960j <- full_join(data1960a, data1960b, by = c("STATEA", "COUNTYA"))
-data1960j <- full_join(data1960j, data1960c, by = c("STATEA", "COUNTYA"))
-data1960j <- data1960j %>%
-  mutate(value_farm = NA, n_farm = NA, cows = NA, horses = NA, 
-         avg_size = NA, value_implements = NA, pop = B47001, male = B5H001, 
-         female = B5H002, enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = NA, 
-         acres_farms = NA, acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop, pct_illiterate = illiterate/pop) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-```
-
-## Join the cleaned sets
-
-```{r boundData}
-# Running this code (before I renamed variables) showed that all .x values are the same as all .y values, 
-# so it was safe to eliminate the double columns
-# as long as problems returned 0 (and it did!)
-# this didn't check for .x.x or .y.x. columns, etc.
-
-#problems <- bind_rows(data1900j, data1910j, data1920j, data1930j, data1940j, data1950j, data1960j) %>% 
-  #select(STATEA, COUNTYA, GISJOIN.x, GISJOIN.y, YEAR.x, YEAR.y, STATE.x, STATE.y, 
-         #COUNTY.x, COUNTY.y) %>%
-  #mutate(same = ifelse(GISJOIN.x == GISJOIN.y & YEAR.x == YEAR.y & STATE.x == STATE.y & 
-                         #COUNTY.x == COUNTY.y, TRUE, FALSE)) %>%
-  #filter(same == "FALSE") %>%
-  #nrow()
-
-# Now bind all of the years together!
-# Add pct_male and avg_value_farm variables
-# Filter for only midwestern states/territories
-
-allData <- as_tibble(bind_rows(data1870j, data1880j, data1890j, 
-                       data1900j, data1910j, data1920j, data1930j, data1940j, data1950j, data1960j)) %>%
-  mutate(pct_male = male/pop, avg_value_farm = value_farm/n_farm, cows_per_farm = cows_per_person, 
-         avg_value_implements = value_implements/n_farm) %>%
-
-  # I mislabeled cows_per_farm as cows_per_person, so I'm doing damage control now
-
-  select(!cows_per_person) %>%
-  filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Dakota Territory" | 
-           STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | 
-           STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | 
-           STATE == "Indiana" | STATE == "Ohio")
-
-# This will be used for mapping!
-# All relevant columns must be numeric in order to function properly
-# Renaming them also helps make the maps more easily interpretable
-# But it's not ideal for working with the data in other contexts, hence the separate set
-
-mwData <- allData %>%
-  mutate("Value of Farms" = as.numeric(value_farm), "Average Farm Value" = as.numeric(avg_value_farm), 
-         "Value of Implements" = as.numeric(value_implements), 
-         "Average Implement Value" = as.numeric(avg_value_implements), 
-         "Average Farm Size" = as.numeric(avg_size), "Number of Farms" = as.numeric(n_farm), 
-         "Acres of Farmland" = as.numeric(acres_farms), "Irrigated Acres" = as.numeric(acres_irrigated), 
-         "Percentage of Farmland Irrigated" = as.numeric(pct_irrigated), "Number of Cows" = as.numeric(cows),
-         "Number of Cows per Farm" = as.numeric(cows_per_farm), "Number of Horses" = as.numeric(horses), 
-         "Total Population" = as.numeric(pop), "Male Population" = as.numeric(male), 
-         "Female Population" = as.numeric(female), "Male Percentage of Population" = as.numeric(pct_male),
-         "Total School Enrollment" = as.numeric(enrolled), 
-         "Percentage School Enrollment" = as.numeric(pct_enrolled), 
-         "Urban Population" = as.numeric(urban_pop), 
-         "Percent of Population in Urban Areas" = as.numeric(pct_urban), 
-         "Illiterate Population" = as.numeric(illiterate), 
-         "Illiteracy Rate" = as.numeric(pct_illiterate)) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, 
-         "Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
-                   "Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
-                   "Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
-                   "Number of Horses", "Total Population", "Male Population", "Female Population", 
-                   "Male Percentage of Population", "Total School Enrollment", 
-                   "Percentage School Enrollment", "Urban Population", 
-                   "Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")
-
-# saving this for later
-# will need in shiny app
-#sliderOptions <- c("Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
-                   #"Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
-                   #"Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
-                   #"Number of Horses", "Total Population", "Male Population", "Female Population", 
-                   #"Male Percentage of Population", "Total School Enrollment", 
-                   #"Percentage School Enrollment", "Urban Population", 
-                   #"Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")
-
-```
-
-# Shiny 
-## Preparation
-
-```{r shapefiles}
-
-# Now save the gis data with short names for sake of the shiny app
-# hopefully this will work...
-
-dsn1870 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870"
-  layer1870 <- "US_county_1870_conflated"
-dsn1880 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1880"
-  layer1880 <- "US_county_1880_conflated"
-dsn1890 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1890"
-  layer1890 <- "US_county_1890_conflated"
-dsn1900 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1900"
-  layer1900 <- "US_county_1900_conflated"
-dsn1910 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1910"
-  layer1910 <- "US_county_1910_conflated"
-dsn1920 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1920"
-  layer1920 <- "US_county_1920_conflated"
-dsn1930 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1930"
-  layer1930 <- "US_county_1930_conflated"
-dsn1940 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1940"
-  layer1940 <- "US_county_1940_conflated"
-dsn1950 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1950"
-  layer1950 <- "US_county_1950_conflated"
-dsn1960 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1960"
-  layer1960 <- "US_county_1960_conflated"
-
-# Make a tibble of all the shapefiles that you'll need
-  shapefiles <- tibble(Year = c(1870, 1880, 1890, 1900, 1910, 1920, 1930, 1940, 1950, 1960), 
-                     Dsn = c(dsn1870, dsn1880, dsn1890, dsn1900, dsn1910, dsn1920, dsn1930, 
-                             dsn1940, dsn1950, dsn1960),
-                     Layer = c(layer1870, layer1880, layer1890, layer1900, layer1910, layer1920, 
-                     layer1930, layer1940, layer1950, layer1960))
-```
-
-## Output
+### Map Code
+Libraries and data are not loaded, so it won't actually run in this environment
 
@@
 
-# df %>% filter(b == !!b)
+# DO NOT CHANGE THIS
+
+# might also want to add a filter for state??
+# if time permits...
 
@@
 
+# Save This
 ```{r shiny, eval=FALSE}
+
+# DO NOT CHANGE
 

### 2020-04-25 ms_8.Rmd
Commit all shiny components

@@
 library(readxl)
+library(gganimate)
 
@@
   mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = APN001+APO001+APQ001, n_farm = AOQ001, cows =  AOG004+AOG005, horses = AOG001, 
-         avg_size = APG001, value_implements = APO001, pop = AOT001, male = AP1001, female = AP1002, 
+  mutate(value_farm = APN001+APO001+APQ001, n_farm = AOQ001, cows =  AOG004+AOG005, horses = AOG001, value_implements = APO001, pop = AOT001, male = AP1001, female = AP1002, 
          enrolled = NA, urban_pop = AO4001, pct_urban = AO4001/AOT001*100, illiterate = NA, 
          acres_farms = AOR001, acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
   mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = AUK001+AUK002+AUK003, n_farm = AS9001, cows = ASY002, horses = ASY003, 
-         avg_size = ATW001, value_implements = AUK002, pop = AUM001, male = AV0001+AV0003+AV0005+AV0007, 
+  mutate(value_farm = AUK001+AUK002+AUK003, n_farm = AS9001, cows = ASY002, horses = ASY003, value_implements = AUK002, pop = AUM001, male = AV0001+AV0003+AV0005+AV0007, 
          female = AV0002+AV0004+AV0006+AV0008, enrolled = AVK005+AVK006+AVK007+AVK008, urban_pop = AUU001, 
@@
          acres_irrigated = AT2001, pct_irrigated = AT2001/(AUJ001+AUJ002)*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
   mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = AWV001, n_farm = AW3001, cows = AW7001, horses = AW7002, 
-         avg_size = AXE001, value_implements = AWW003, pop = AWS001, male = AZ0001,female = AZ0001, 
+  mutate(value_farm = AWV001, n_farm = AW3001, cows = AW7001, horses = AW7002, value_implements = AWW003, pop = AWS001, male = AZ0001,female = AZ0001, 
          enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = AYS001+AYS002+AYS003+AYS004+AYS005, 
          acres_farms = AWT001, acres_irrigated = AX8001, pct_irrigated = AX8001/AWT001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
   mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = BFD001+BFD002+BFD004+ACEO001, n_farm = AB95001, cows = NA, horses = NA, 
-         avg_size = ACAA001, value_implements = BFD004, pop = BCX001, male = BC3001, 
+  mutate(value_farm = BFD001+BFD002+BFD004+ACEO001, n_farm = AB95001, cows = NA, horses = NA, value_implements = BFD004, pop = BCX001, male = BC3001, 
          female = BC3002, enrolled = BDT001+BDT002+BDT003+BDT004, urban_pop = BDX001, 
@@
          acres_irrigated = ACEX002, pct_irrigated = ACEX002/AB98001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
 
-  mutate(value_farm = BWD001+BWD002, n_farm = BY1001, cows = NA, horses = NA, 
-         avg_size = BY0001, value_implements = BWD002, pop = BVU001, male = BVY001, 
+  mutate(value_farm = BWD001+BWD002, n_farm = BY1001, cows = NA, horses = NA, value_implements = BWD002, pop = BVU001, male = BVY001, 
          female = BVY002, enrolled = BWU001+BWU002+BWU003+BWU004+BWU005, urban_pop = NA, 
@@
          acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, 
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, 
          pct_illiterate = illiterate/pop*100) %>%
@@
 
-  mutate(value_farm = B3M001, n_farm = B3Z001, cows = NA, horses = NA, 
-         avg_size = B3L001, value_implements = NA, pop = B13001, male = B3E001, 
+  mutate(value_farm = B3M001, n_farm = B3Z001, cows = NA, horses = NA, value_implements = NA, pop = B13001, male = B3E001, 
          female = B3E002, enrolled = B2A001+B2A002+B2A003+B2A004+B2A005, urban_pop = B2J001, 
@@
          acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
   select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
@@
 ```{r experimentalModels}
-# USE subData FOR ALL MODELS
-# EVERYTHING ELSE IS DUBIOUS
+
+# INCORPORATE ALL OF THESE IN SHINY
 
@@
   geom_smooth(method = "lm", se = FALSE) +
+  scale_fill_viridis_d(aesthetics = "color") +
   ylim(0, 45) + 
-  labs(title = "Relationship Between Farm Value and Education, by County",
+  labs(title = "Relationship Between Farm Implements and Education, by County",
        subtitle = "Midwestern United States", 
@@
 
-# Incorporate
-# States are gaining more implements each year
-imp_plot <- allData %>%
-  mutate(YEAR = as.factor(YEAR)) %>%
-  select(YEAR, STATE, value_implements) %>%
-  na.omit() %>%
-  group_by(YEAR, STATE) %>%
-  summarize(mean_imp = mean(value_implements)) %>% 
-  ungroup() %>%
-  group_by(STATE) %>%
-  ggplot(aes(x = YEAR, y = mean_imp, fill = STATE)) + 
-  geom_col() +
-  labs(title = "Implement Value by Year, by State",
-       subtitle = "Dollar Values Adjusted for Inflation",
-       x = "Year",
-       y = "Average Value of All Implements",
-       fill = "State") +
-  theme_classic()
+ggsave("final/impedu_plot.png", plot = impedu_plot)
+
+# See implement plot (plot2) saved as gif
 
@@
 
+# couldn't get gtsave to save to the "final" subdirectory
+# will need to copy to "final" manually??
+gtsave(data = maleedu_gt, filename = "maleedu_gt.html")
+
 # INCORPORATE
@@
   ggplot(aes(x = YEAR, y = pct_male_coef)) + 
-  geom_errorbar(ymin = maleedu_model$pct_male_lower, ymax = maleedu_model$pct_male_upper) + 
-  geom_point() +
+  geom_errorbar(ymin = maleedu_model$pct_male_lower, ymax = maleedu_model$pct_male_upper, color = "#453781FF") + 
+  geom_point(color = "#440154FF") +
   theme_classic() +
@@
 
-# avg_value_implements and cows_per_farm have an effect on enrollment
-# in general, for 1879 and 1890, more cows per farm = less enrollment
-# aka bigger farms = less enrollment?
-model2a <- modelData %>% 
-  filter(YEAR == "1870") %>%
-  lm(data = ., pct_enrolled ~ cows_per_farm) %>% 
-  tidy(conf.int = TRUE) %>% 
-  select(term, estimate, p.value, conf.low, conf.high) %>%
+ggsave("final/maleedu_plot.png", plot = maleedu_plot)
+
+# INCORPORATE
+# more cows = more enrollment
+cowedu_model <- allData %>%
+  select(YEAR, pct_enrolled, cows_per_farm, avg_value_implements) %>%
+  na.omit() %>%
+  lm(pct_enrolled ~ cows_per_farm*avg_value_implements, data = .) %>%
+  tidy(conf.int = TRUE)
+
+cowedu_gt <- cowedu_model %>%
+  select(term, conf.low, estimate, conf.high) %>%
   gt() %>%
-  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
-             decimals = 2)
-
-# there was an infinite value left in the cows_per_farm column for this year
-# not sure why, but I got rid of it
-model2a <- modelData %>% 
-  filter(YEAR == "1890", !is.infinite(cows_per_farm)) %>%
-  lm(data = ., pct_enrolled ~ cows_per_farm) %>% 
-  tidy(conf.int = TRUE) %>% 
-  select(term, estimate, p.value, conf.low, conf.high) %>%
-  gt() %>% 
-  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
-             decimals = 2)
-
-model2c <- modelData %>% 
-  filter(YEAR == "1910") %>%
-  lm(data = ., pct_enrolled ~ cows_per_farm) %>% 
-  tidy(conf.int = TRUE) %>% 
-  select(term, estimate, p.value, conf.low, conf.high) %>%
-  gt() %>% 
-  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
-             decimals = 2)
-
-model2
+  tab_header(title = "Effect Cows and Implements on Enrollment",
+             subtitle = "With 95% Confidence Interval") %>%
+  cols_label(term = "Coefficient Name", conf.low = "Lower Bound", estimate = "Estimated Value", conf.high = "Upper Bound") %>%
+  fmt_number(columns = vars(conf.low, estimate, conf.high), decimals = 4)
+
+# copy to "final"
+gtsave(data = cowedu_gt, filename = "cowedu_gt.html")
+
 # cow plot
 # incorporate
-cowplot <- modelData %>% 
-  filter(YEAR == "1870" | YEAR == "1890" | YEAR == "1910") %>%
+cowedu_plot <- allData %>%
+  select(YEAR, pct_enrolled, cows_per_farm) %>%
+  na.omit() %>%
+  mutate(YEAR = as.factor(YEAR)) %>%
+  group_by(YEAR) %>%
   ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
-  geom_point(alpha = 0.5) +
+  geom_point(alpha = 0.1) +
   geom_smooth(method = "lm", se = FALSE) +
+  scale_color_viridis_d() +
+  ylim(0, 45) +
   scale_x_log10(labels = comma) +
@@
 
-modelData %>% 
-  filter(YEAR == "1870" | YEAR == "1890" | YEAR == "1910") %>%
-  ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
-  geom_point() +
-  geom_smooth(se = FALSE)
-
-# Enrollment/Illiteracy model (just for testing)
-
-# This is saying that for every 1% increase in illiterate people, there is a 0.24% 
-# decrease in school enrollment, which makes sense
-model3 <- modelData %>%
-  lm(data = ., pct_enrolled ~ pct_illiterate) %>%
-  tidy(conf.int = TRUE) %>%
-  select(term, estimate, p.value, conf.low, conf.high) %>%
-  gt() %>%
-  fmt_number(data = ., columns = c("estimate", "p.value", "conf.low", "conf.high"), 
-             decimals = 2)
-
-mod3plot <- modelData %>%
-  filter(! is.na(pct_enrolled) & ! is.na(pct_illiterate)) %>%
-  ggplot(aes(pct_enrolled, pct_illiterate)) + 
-  geom_point() + 
-  geom_smooth(se = FALSE)
-```
-
-```{r moreModels}
-
-# can only use percentages until adjusting for inflation
-# OR deal with each year independently
-
-# Set the NA values to almost zero
-# Not ideal, but necessary for running regressions
-# setting to zero returns NaNs
-
-reg_years <- allData %>% 
-  # sample the data to run faster
-  # change this later
-  sample_n(size = 5000) %>%
-  # Get rid of the infinite cows_per_farm values
-  filter(! (is.na(pct_enrolled) & is.na(avg_size))) %>%
-  group_by(YEAR) %>%
-  arrange(desc(YEAR)) %>% 
-  nest()
-
-  # Call map inside of mutate to add new columns to the tibble
-  # reg_results are the results of the regression we ran in question 2
-  # coef is the coefficient of log_migrants for each continent
-  # upper and lower are the upper and lower bounds of the confidence interval created by tidy()
-
-  # options
-  #STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         # acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_farm, horses, pop, male, female, 
-        # enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate
-reg_years1 <- reg_years %>%
-  mutate(mod = map(data, ~ lm(pct_enrolled ~ avg_size + avg_value_implements + n_farm + pct_irrigated + cows_per_farm + pct_urban, data = .)),
-         reg_results = map(mod, ~ tidy(., conf.int = TRUE)),
-         avg_size_coef = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% 
-                                   pull(estimate)),
-         avg_size_upper = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% 
-                                    pull(conf.high)),
-         avg_size_lower = map_dbl(reg_results, ~ filter(., term == "avg_size") %>% 
-                                    pull(conf.low)),
-         avg_value_implements_coef = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% 
-                                               pull(estimate)),
-         avg_value_implements_upper = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% 
-                                                pull(conf.high)),
-         avg_value_implements_lower = map_dbl(reg_results, ~ filter(., term == "avg_value_implements") %>% 
-                                                pull(conf.low)),
-         n_farm_coef = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(estimate)),
-         n_farm_upper = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(conf.high)),
-         n_farm_lower = map_dbl(reg_results, ~ filter(., term == "n_farm") %>% pull(conf.low)),
-         pct_irrigated_coef = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(estimate)),
-         pct_irrigated_upper = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(conf.high)),
-         pct_irrigated_lower = map_dbl(reg_results, ~ filter(., term == "pct_irrigated") %>% pull(conf.low)),
-         cows_per_farm_coef = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(estimate)),
-         cows_per_farm_upper = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(conf.high)),
-         cows_per_farm_lower = map_dbl(reg_results, ~ filter(., term == "cows_per_farm") %>% pull(conf.low)),
-         pct_urban_coef = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(estimate)),
-         pct_urban_upper = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(conf.high)),
-         pct_urban_lower = map_dbl(reg_results, ~ filter(., term == "pct_urban") %>% pull(conf.low))
-  )
-
-# Now use the resulting tibble to create the gt table
-
-q4_gt <- q4 %>%
-
-  # The original year column was giving me problems because it's a grouping variable
-  # Ungrouping solved the problem
-
+ggsave("final/cowedu_plot.png", plot = cowedu_plot)
+
+```
+
+```{r farmTrends}
+# this takes foever--cache it
+
+# INCORPORATE 
+# just the gif files, not the whole code
+# takes too long
+
+# Trends in Midwestern Farming over Time
+
+plot1a <- allData %>%
+  select(YEAR, STATE, acres_farms, n_farm) %>%
+  na.omit() %>%
+  group_by(STATE, YEAR) %>%
+  summarize(mean_acres = mean(acres_farms), mean_farm = mean(n_farm)) %>%
+  ggplot(aes(x = YEAR, y = mean_acres, fill = STATE)) +
+  geom_col() + 
+  scale_fill_viridis_d() +
+  labs(x = "Year",
+       y = "Mean Acreage (Cumulative)",
+       title = "Average Farmland Acreage over Time",
+       fill = "State") +
+  theme_classic() + 
+  transition_states(YEAR, wrap = FALSE) +
+  shadow_mark()
+
+anim_save("final/plot1a.gif", plot1a)
+
+plot1b_prep <- allData %>%
+  select(YEAR, STATE, n_farm) %>%
+  na.omit() %>%
+  group_by(STATE, YEAR)
+
+plot1b <- aggregate(plot1b_prep$n_farm, by = list(plot1b_prep$STATE, 
+                                                  plot1b_prep$YEAR), FUN=sum) %>%
+    mutate(STATE = Group.1, YEAR = Group.2, n = x) %>%
+  ggplot(aes(x = YEAR, y = n, color = STATE)) +
+  geom_line() + 
+  scale_fill_viridis_d(aesthetics = "color") +
+  labs(x = "Year",
+       y = "Number of Farms",
+       title = "Number of Farms over Time, by State",
+       subtitle = "The number of farms increases from 1870 to the early 1900s, but is in decline by 1930",
+       color = "State") +
+    theme_classic() + 
+    # animate
+    transition_reveal(YEAR)
+
+anim_save("final/plot1b.gif", plot1b)
+
+# States are gaining more implements each year
+plot2 <- allData %>%
+  mutate(YEAR = as.factor(YEAR)) %>%
+  select(YEAR, STATE, value_implements) %>%
+  na.omit() %>%
+  group_by(YEAR, STATE) %>%
+  summarize(mean_imp = mean(value_implements)) %>% 
   ungroup() %>%
-
-  # Select the columns needed, as per the instructions
-
-  select(YEAR, coef, lower, upper) %>%
-
-  gt() %>%
-
-  # Add the desired titles, subtitles, and column labels
-
-  tab_header(title = "Title",
-             subtitle = "With 95% Confidence Interval") %>%
-  cols_label(YEAR = "Year", coef = "Migrant Coefficient", 
-             lower = "Lower Bound", upper = "Upper Bound") %>%
-
-  # Round all values to two decimal places
-
-  fmt_number(columns = vars(coef, lower, upper), decimals = 2)
-
+  group_by(STATE) %>%
+  ggplot(aes(x = YEAR, y = mean_imp, fill = STATE)) + 
+  geom_col() +
+  scale_fill_viridis_d() +
+  labs(title = "Implement Value by Year, by State",
+       subtitle = "Dollar Values Adjusted for Inflation",
+       x = "Year",
+       y = "Average Value of All Implements",
+       fill = "State") +
+  theme_classic() + 
+  transition_states(YEAR, wrap = FALSE) +
+  shadow_mark()
+
+anim_save("final/plot2.gif", plot2)
+
+# Incorporate
+# population plot
+
+plot3_prep <- allData %>%
+  select(YEAR, STATE, pop) %>%
+  na.omit() %>%
+  group_by(STATE, YEAR)
+
+plot3 <- aggregate(plot3_prep$pop, by = list(plot3_prep$STATE, 
+                                                  plot3_prep$YEAR), FUN=sum) %>%
+    mutate(STATE = Group.1, YEAR = Group.2, n = x) %>%
+  ggplot(aes(x = YEAR, y = n, color = STATE)) +
+  geom_line() + 
+  scale_fill_viridis_d(aesthetics = "color") +
+  labs(x = "Year",
+       y = "Population",
+       title = "Midwest Population over Time, by State",
+       color = "State") +
+    theme_classic() + 
+    # animate
+    transition_reveal(YEAR)
+
+anim_save("final/plot3.gif", plot3)
+
+# Urbanization plot
+# Urban = 2,500+
+
+plot4_prep <- allData %>%
+  select(YEAR, pop, urban_pop) %>%
+  na.omit() %>%
+  mutate(pop = pop-urban_pop)
+
+plot4_prepb <- aggregate(plot4_prep$pop, by = list(plot4_prep$YEAR), 
+                   FUN=sum) %>%
+  mutate(YEAR = Group.1, n_pop = x)
+
+plot4_prepc <- aggregate(plot4_prep$urban_pop, by = list(plot4_prep$YEAR), 
+                   FUN=sum) %>% 
+  mutate(YEAR = Group.1, n_urban_pop = x)
+
+plot4_prepd <- full_join(plot4_prepb, plot4_prepc, by = c("YEAR")) %>%
+  select(YEAR, n_pop, n_urban_pop) %>%
+  mutate(YEAR = as.factor(YEAR)) %>%
+  # convert to long format
+  melt(id="YEAR")
+
+limits4 <- c("1", "2", "3", "4", "5", "6", "7")
+labels4 <- c("1870", "1880", "1890", "1910", "1920", "1930", "1950")
+
+plot4 <- as_tibble(plot4_prepd) %>%
+  # as.numeric is changing years to numbers 1:7
+  mutate(YEAR = as.numeric(YEAR), variable = as.character(variable), 
+         value = value/1000) %>%
+  ggplot(aes(YEAR, value, color = variable)) +
+  geom_line() +
+  scale_x_discrete(limits = limits4, labels = labels4) +
+  scale_fill_viridis_d(aesthetics = "color", breaks=c("n_urban_pop", "n_pop"), labels = c("Urban", "Rural")) +
+  labs(x = "Year",
+       y = "Population (in thousands)",
+       title = "Midwest Urbanization",
+       subtitle = "Rural and urban populations over Time",
+       color = "") +
+  theme_classic() +
+  geom_point(aes(group = seq_along(YEAR))) +
+  transition_reveal(YEAR)
+
+anim_save("final/plot4.gif", plot4)
 ```
@@
 
-# PASTE DATA and LIBRARIES here after
-
-# Input
-
-ui <- fluidPage(
-
-    # flatly is a nice theme to make the tabs look put together
-
-    # Application title
-
-    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),
-
-               # remember that tab panel is for what shows up in the tab, and title panel is 
-               # for what shows up at the top of the page within the tab
-
-               tabPanel("Tab 1",
-
-                        # Give the page a title
-                        titlePanel("Relationship Between Farm Value and Education, by County"),
-
-                        # Generate a row with a sidebar
-                        sidebarLayout(
-
-                            # Define the sidebar with one input
-                            # insteaad of region, I want year ######
-                            sidebarPanel(
-                                selectInput("year", "Year:", 
-                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
-                                            multiple = TRUE)),
-                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
-                        ),
-
-                        # Create a spot for the plot
-                        mainPanel(
-                            tabsetPanel(id = "tabsMain",
-                                        tabPanel("Plot",
-
-                                                 # this tells the UI what plot to put
-
-                                                 plotOutput("agPlot")  
-                                        )))),
-               tabPanel("Models",
-
-                        # Give the page a title
-                        titlePanel("Relationship Between Farm Value and Education, by County"),
-                        p("Make some models! (See notes)"),
-                        gt_output("model1")
-               ),
-               tabPanel("Maps",
-
-                # Give the page a title
-                titlePanel("Title Relating to Maps"),
-                p("Add a slider for different years/variables & figure out how to get two maps side by side"),
-                p("Maps may take up to 30 seconds to load."),
-               leafletOutput("map1"),
-               p()),
-
-               tabPanel("About",
-
-                        # Give the page a title
-                        titlePanel("About this project"),
-                        p("My name is Julia Englebert.
-I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights.
-Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
-                        br(),
-                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States.
-By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century.
-                        I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960.
-                          My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org.
-IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960.
-All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
-                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
-                        br(),
-                        p("To see more of my work, please visit my Github account:"),
-                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
-                        br(),
-                        titlePanel("Citations"),
-                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles.
-IPUMS National Historical Geographic Information System: Version 14.0 [Database].
-Minneapolis, MN: IPUMS.
-2019.
-http://doi.org/10.18128/D050.V14.0"),
-                        p("The College of William and Mary and the Minnesota Population Center.
-School Attendance Boundary Information System (SABINS): Version 1.0.
-Minneapolis, MN: University of Minnesota 2011.")
-               )
-    ))
-
-# use plotly to allow hover text to appear on plots, giving more info about the point
-
-server <- function(input, output) {
-
-    # Fill in the spot we created for a plot
-    output$agPlot <- renderPlot({
-
-        # Render a plot
-        b_1870_1910 %>% 
-            filter(year %in% input$year) %>% 
-            group_by(year) %>%
-            ggplot(aes(x = value_farms, y = perc_school, color = year)) + 
-            geom_point(alpha = 0.5) + 
-            ylim(0, 100) + 
-            geom_smooth(method = "lm") + 
-            labs(title = "Relationship Between Farm Value and Education, by County", 
-                 subtitle = "Great Plains Region, United States", 
-                 x = "Total Dollar Value of Farms", 
-                 y = "Percent Students Ages 5-24 Enrolled in School") +
-            scale_x_log10(labels = comma) +
-            theme_classic()
-    })
-
-    output$map1 <- renderLeaflet({
-            data1870 <- data1870j %>%
-                filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | STATE == "Indiana" | STATE == "Ohio", AJX001 > 0) %>%
-                mutate(valueAllFarmProds1870 = AJX001) %>%
-                select(GISJOIN, valueAllFarmProds1870)
-
-            # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
-            # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
-            # layer is the filename of the .shp file inside the
-            # folder dsn points to.
-
-            county1870 <- sf::st_read(dsn = "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870",
-                                      layer = "US_county_1870_conflated")
-
-            county1870 <-
-                county1870 %>%
-                merge(data1870, "GISJOIN")
-
-            # Set projection of tracts dataset to `projection` required by leaflet
-
-            county1870 <- sf::st_transform(county1870, crs="+init=epsg:4326")
-
-            # Condense size of data for faster processing
-
-            county1870 <- rmapshaper::ms_simplify(county1870)
-
-            # Set palette color
-
-            pal <- colorNumeric("viridis", NULL)
-
-            #  Plot the data
-
-            leaflet(county1870) %>%
-                addTiles() %>%
-                addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
-                            fillColor = ~pal(valueAllFarmProds1870)) %>%
-                addLegend(pal = pal, values = ~valueAllFarmProds1870, opacity = 1.0, title = "1870 Value of All Farm Productions,\nby Dollar")
-    })
-
-    output$model1 <- render_gt({
-
-        # experimental model
-
-        model1
-
-    })
-}
-
-# Run the application
-
-shinyApp(ui = ui, server = server)
-```
-
-# DO NOT MAKE CHANGES BELOW THIS LINE
-
-### Map Code
-
-```{r genericGis}
-
-# DO NOT CHANGE THIS
-
-# might also want to add a filter for state??
-# if time permits...
-
-# gis template for any year/variable desired
-# only change these two lines!
-year <- 1870
-myvar <- "Average Farm Value"
-
-datayear <- mwData %>%
-  # use get() to use the stored variable
-  # idk why, but it doesn't work without this
-  # the variable must be converted to a double in the original mwData dataset
-  mutate(myvariable = (get(myvar))) %>%
-  filter(YEAR == year, myvariable > 0) %>%
-  select(GISJOIN, myvariable)
-
-# Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
-# dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
-# layer is the filename of the .shp file inside the
-# folder dsn points to.
-
-dsnyear <- shapefiles %>% 
-  filter(Year == year) %>% 
-  pull(Dsn)
-layeryear <- shapefiles %>% 
-  filter(Year == year) %>% 
-  pull(Layer)
-
-countyyear <- sf::st_read(dsn = dsnyear,
-                         layer = layeryear)
-
-countyyear <-
-  countyyear %>%
-  merge(datayear, "GISJOIN")
-
-# Set projection of tracts dataset to `projection` required by leaflet
-
-countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
-
-# Condense size of data for faster processing
-
-countyyear <- rmapshaper::ms_simplify(countyyear)
-
-# Set palette color
-
-pal <- colorNumeric("viridis", NULL)
-
-#  Plot the data
-
-# use shiny to add a title, which will be be equivalent to myvar
-
-leaflet(countyyear) %>%
-  addTiles() %>%
-  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
-              fillColor = ~pal(myvariable)) %>%
-  addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
-
-```
-
-# Save This
-```{r shiny, eval=FALSE}
-
-# Pasting this here in case my computer explodes
-
-# This is a Shiny web application. You can run the application by clicking
-# the 'Run App' button above.
-#
-# Find out more about building applications with Shiny here:
-#
-#    http://shiny.rstudio.com/
-#
-
-library(shiny)
-library(haven)
-library(tidyverse)
-library(gt)
-library(devtools)
-library(tibble)
-library(foreign)
-library(scales)
-library(shinythemes)
-library(plotly)
-library(leaflet)
-library(broom)
-
-# packages for mapping
-library(leaflet)
-library(spdplyr)
-library(rmapshaper)
-library(sf)
-
-# The real data
-
-data1870a <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds16_1870_county.csv")
-data1870b <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds17_1870_county.csv")
-data1870j <- full_join(data1870a, data1870b, by = c("GISJOIN", "STATE", "COUNTY"))
-
-data1880a <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds22_1880_county.csv") 
-data1880b <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds23_1880_county.csv")
-data1880j <- full_join(data1880a, data1880b, by = c("GISJOIN", "STATE", "COUNTY"))
-
-data1890a <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds26_1890_county.csv")
-data1890b <- read.csv("raw-data/nhgis0002_csv/nhgis0002_ds27_1890_county.csv")
-data1890j <- full_join(data1890a, data1890b, by = c("GISJOIN", "STATE", "COUNTY"))
-
-data1800s <- bind_rows(data1870j, data1880j, data1890j)
-
-# I'm not actually using this data anymore
-# I just wanted to make sure I could get my plots running smoothly
-tbl_1 <- read_por("ICPSR_04254_ag/DS0001/04254-0001-Data.por", col_select = c(NAME, UNFIPS, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_XX_V, IMP_XX_V, AFP_XX_V)) %>%
-    rename(name = NAME, fips = UNFIPS, land_in_farms = FML_XX_A, total_area = ARE_XX_A, num_farms = FRM_XX_Q, value_farms = PRP_XX_V, value_implements = IMP_XX_V, value_production_betterments_stocks = AFP_XX_V)
-
-tbl_3 <- read_por("ICPSR_04254_ag/DS0003/04254-0003-Data.por", col_select = c(NAME, UNFIPS, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_XX_V, IMP_XX_V, AFP_XX_V))%>%
-    rename(name = NAME, fips = UNFIPS, land_in_farms = FML_XX_A, total_area = ARE_XX_A, num_farms = FRM_XX_Q, value_farms = PRP_XX_V, value_implements = IMP_XX_V, value_production_betterments_stocks = AFP_XX_V)
-
-tbl_5 <- read_por("ICPSR_04254_ag/DS0005/04254-0005-Data.por", col_select = c(NAME, UNFIPS, FML_XX_A, ARE_XX_A, FRM_XX_Q, PRP_OF_V, IMP_XX_V, CRO_XX_V)) %>%
-    rename(name = NAME, fips = UNFIPS, land_in_farms = FML_XX_A, total_area = ARE_XX_A, num_farms = FRM_XX_Q, value_implements = IMP_XX_V, value_crops = CRO_XX_V) %>%
-    mutate(value_farms = PRP_OF_V) %>%
-    mutate(value_livestock = "NA", value_productsNotFed = "NA") %>%
-    select(name, fips, land_in_farms, total_area, num_farms, value_farms, value_implements, value_livestock, value_productsNotFed)
-
-tbl_1_d <- read_por("ICPSR_04296_demog/DS0001/04296-0001-Data.por", col_select = c(NAME, UNFIPS, TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, POP10P, SCH)) %>%
-    rename(name = NAME, fips = UNFIPS, total = TTOT, people_over_10 = POP10P, pupils_in_school = SCH) %>%
-    mutate(fem_5_24 = FT5_9 + FT10_14 + FT15_19 + FT20_24, mal_5_24 = MT5_9 + MT10_14 + MT15_19 + MT20_24)
-
-tbl_3_d <- read_por("ICPSR_04296_demog/DS0003/04296-0003-Data.por", col_select = c(NAME, UNFIPS, TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH)) %>%
-    rename(name = NAME, fips = UNFIPS, total = TTOT, pupils_in_school = SCH) %>%
-    mutate(fem_5_24 = FT5_9 + FT10_14 + FT15_19 + FT20_24, mal_5_24 = MT5_9 + MT10_14 + MT15_19 + MT20_24)
-
-tbl_5_d <- read_por("ICPSR_04296_demog/DS0005/04296-0005-Data.por", col_select = c(NAME, UNFIPS, TTOT, FT0_4, FT5_9, FT10_14, FT15_19, FT20_24, FT25_29, MT0_4, MT5_9, MT10_14, MT15_19, MT20_24, MT25_29, SCH69, SCH1014, SCH1517, SCH1820)) %>%
-    rename(name = NAME, fips = UNFIPS, total = TTOT) %>%
-    mutate(fem_5_24 = FT5_9 + FT10_14 + FT15_19 + FT20_24, mal_5_24 = MT5_9 + MT10_14 + MT15_19 + MT20_24,
-           # Inconsistent age ranges
-           # Not sure how to fix this
-           pupils_in_school_6_20 = SCH69 + SCH1014 + SCH1517 + SCH1820)
-
-j_1870 <- full_join(tbl_1, tbl_1_d, by = "fips") %>% 
-    mutate(perc_school = pupils_in_school/(fem_5_24 + mal_5_24)*100) %>%
-    select(name.x, fips, value_farms, perc_school) %>%
-    group_by(fips)
-
-j_1870 <- na.omit(j_1870)
-
-j_1890 <- full_join(tbl_3, tbl_3_d, by = "fips") %>%
-    mutate(perc_school = pupils_in_school/(fem_5_24 + mal_5_24)*100) %>%
-    select(name.x, fips, value_farms, perc_school) %>%
-    group_by(fips)
-
-j_1890 <- na.omit(j_1890)
-
-j_1910 <- full_join(tbl_5, tbl_5_d, by = "fips") %>%
-    mutate(perc_school = pupils_in_school_6_20/(fem_5_24 + mal_5_24)*100) %>%
-    select(name.x, fips, value_farms, perc_school) %>%
-    group_by(fips)
-
-j_1910 <- na.omit(j_1910)
-
-b_1870_1910 <- bind_rows("1870" = j_1870, "1890" = j_1890, "1910" = j_1910, .id = "year") %>%
-    group_by(year)
-
-# this should be done by state, not fips!!!
-model1 <- lm(data = b_1870_1910, perc_school ~ fips + value_farms) %>% 
-    tidy(conf.int = TRUE) %>% 
-    gt() %>%
-    fmt_number(data = ., columns = c("estimate", "std.error", "statistic", "p.value", "conf.low", "conf.high"), decimals = 2)
-
-# Input
-
-ui <- fluidPage(
-
-    # flatly is a nice theme to make the tabs look put together
-
-    # Application title
-
-    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),
-
-               # remember that tab panel is for what shows up in the tab, and title panel is 
-               # for what shows up at the top of the page within the tab
-
-               tabPanel("Tab 1",
-
-                        # Give the page a title
-                        titlePanel("Relationship Between Farm Value and Education, by County"),
-
-                        # Generate a row with a sidebar
-                        sidebarLayout(
-
-                            # Define the sidebar with one input
-                            # insteaad of region, I want year ######
-                            sidebarPanel(
-                                selectInput("year", "Year:", 
-                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
-                                            multiple = TRUE)),
-                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
-                        ),
-
-                        # Create a spot for the plot
-                        mainPanel(
-                            tabsetPanel(id = "tabsMain",
-                                        tabPanel("Plot",
-
-                                                 # this tells the UI what plot to put
-
-                                                 plotOutput("agPlot")  
-                                        )))),
-               tabPanel("Models",
-
-                        # Give the page a title
-                        titlePanel("Relationship Between Farm Value and Education, by County"),
-                        p("Make some models! (See notes)"),
-                        gt_output("model1")
-               ),
-               tabPanel("Maps",
-
-                # Give the page a title
-                titlePanel("Title Relating to Maps"),
-                p("Add a slider for different years/variables & figure out how to get two maps side by side"),
-                p("Maps may take up to 30 seconds to load."),
-               leafletOutput("map1"),
-               p()),
-
-               tabPanel("About",
-
-                        # Give the page a title
-                        titlePanel("About this project"),
-                        p("My name is Julia Englebert.
-I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights.
-Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
-                        br(),
-                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States.
-By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century.
-                        I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960.
-                          My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org.
-IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960.
-All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
-                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
-                        br(),
-                        p("To see more of my work, please visit my Github account:"),
-                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
-                        br(),
-                        titlePanel("Citations"),
-                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles.
-IPUMS National Historical Geographic Information System: Version 14.0 [Database].
-Minneapolis, MN: IPUMS.
-2019.
-http://doi.org/10.18128/D050.V14.0"),
-                        p("The College of William and Mary and the Minnesota Population Center.
-School Attendance Boundary Information System (SABINS): Version 1.0.
-Minneapolis, MN: University of Minnesota 2011.")
-               )
-    ))
-
-# use plotly to allow hover text to appear on plots, giving more info about the point
-
-server <- function(input, output) {
-
-    # Fill in the spot we created for a plot
-    output$agPlot <- renderPlot({
-
-        # Render a plot
-        b_1870_1910 %>% 
-            filter(year %in% input$year) %>% 
-            group_by(year) %>%
-            ggplot(aes(x = value_farms, y = perc_school, color = year)) + 
-            geom_point(alpha = 0.5) + 
-            ylim(0, 100) + 
-            geom_smooth(method = "lm") + 
-            labs(title = "Relationship Between Farm Value and Education, by County", 
-                 subtitle = "Great Plains Region, United States", 
-                 x = "Total Dollar Value of Farms", 
-                 y = "Percent Students Ages 5-24 Enrolled in School") +
-            scale_x_log10(labels = comma) +
-            theme_classic()
-    })
-
-    output$map1 <- renderLeaflet({
-            data1870 <- data1870j %>%
-                filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | STATE == "Indiana" | STATE == "Ohio", AJX001 > 0) %>%
-                mutate(valueAllFarmProds1870 = AJX001) %>%
-                select(GISJOIN, valueAllFarmProds1870)
-
-            # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
-            # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
-            # layer is the filename of the .shp file inside the
-            # folder dsn points to.
-
-            county1870 <- sf::st_read(dsn = "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870",
-                                      layer = "US_county_1870_conflated")
-
-            county1870 <-
-                county1870 %>%
-                merge(data1870, "GISJOIN")
-
-            # Set projection of tracts dataset to `projection` required by leaflet
-
-            county1870 <- sf::st_transform(county1870, crs="+init=epsg:4326")
-
-            # Condense size of data for faster processing
-
-            county1870 <- rmapshaper::ms_simplify(county1870)
-
-            # Set palette color
-
-            pal <- colorNumeric("viridis", NULL)
-
-            #  Plot the data
-
-            leaflet(county1870) %>%
-                addTiles() %>%
-                addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
-                            fillColor = ~pal(valueAllFarmProds1870)) %>%
-                addLegend(pal = pal, values = ~valueAllFarmProds1870, opacity = 1.0, title = "1870 Value of All Farm Productions,\nby Dollar")
-    })
-
-    output$model1 <- render_gt({
-
-        # experimental model
-
-        model1
-
-    })
-}
-
-# Run the application
-
-shinyApp(ui = ui, server = server)
-
-```
+# Paste here after
+
+```

### 2020-04-25 final/app.R

@@ -1,0 +1,222 @@
+#
+# This is a Shiny web application. You can run the application by clicking
+# the 'Run App' button above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+library(shiny)
+library(haven)
+library(tidyverse)
+library(gt)
+library(devtools)
+library(tibble)
+library(foreign)
+library(scales)
+library(shinythemes)
+library(plotly)
+library(leaflet)
+library(broom)
+
+# packages for mapping
+library(leaflet)
+library(spdplyr)
+library(rmapshaper)
+library(sf)
+
+
+data1800s <- readRDS("my_data.rds")
+
+
+# Input
+
+ui <- fluidPage(
+    
+    # flatly is a nice theme to make the tabs look put together
+    
+    # Application title
+    
+    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),
+               
+               # remember that tab panel is for what shows up in the tab, and title panel is 
+               # for what shows up at the top of the page within the tab                  
+               
+               tabPanel("Tab 1",
+                        
+                        # Give the page a title
+                        titlePanel("Relationship Between Farm Value and Education, by County"),
+                        
+                        # Generate a row with a sidebar
+                        sidebarLayout(      
+                            
+                            # Define the sidebar with one input
+                            # insteaad of region, I want year ######
+                            sidebarPanel(
+                                selectInput("year", "Year:", 
+                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
+                                            multiple = TRUE)),
+                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
+                        ),
+                        
+                        # Create a spot for the plot
+                        mainPanel(
+                            tabsetPanel(id = "tabsMain",
+                                        tabPanel("Plot",
+                                                 
+                                                 # this tells the UI what plot to put 
+                                                 
+                                                 plotOutput("agPlot")  
+                                        )))),
+               tabPanel("Models",
+                        
+                        # Give the page a title
+                        titlePanel("Relationship Between Farm Value and Education, by County"),
+                        p("Make some models! (See notes)"),
+                        gt_output("model1")
+               ),
+               tabPanel("Maps",
+                
[…]

### 2020-04-26 testing123/app.R
Functional multi-widget maps prototype!!!!

@@ -1,0 +1,257 @@
+# testing123
+
+# This is a Shiny web application. You can run the application by clicking
+# the 'Run App' button above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+options(scipen = 999)
+
+library(shiny)
+library(haven)
+library(tidyverse)
+library(gt)
+library(devtools)
+library(tibble)
+library(foreign)
+library(scales)
+library(shinythemes)
+library(plotly)
+library(leaflet)
+library(broom)
+library(readxl)
+library(gganimate)
+
+# packages for mapping
+library(leaflet)
+library(spdplyr)
+library(rmapshaper)
+library(sf)
+
+# Prep the shapefiles
+# Now save the gis data with short names for sake of the shiny app
+# hopefully this will work...
+
+dsn1870 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870"
+layer1870 <- "US_county_1870_conflated"
+dsn1880 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1880"
+layer1880 <- "US_county_1880_conflated"
+dsn1890 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1890"
+layer1890 <- "US_county_1890_conflated"
+dsn1900 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1900"
+layer1900 <- "US_county_1900_conflated"
+dsn1910 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1910"
+layer1910 <- "US_county_1910_conflated"
+dsn1920 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1920"
+layer1920 <- "US_county_1920_conflated"
+dsn1930 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1930"
+layer1930 <- "US_county_1930_conflated"
+dsn1940 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1940"
+layer1940 <- "US_county_1940_conflated"
+dsn1950 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1950"
[…]

### 2020-04-26 ms_8.Rmd
Commit with working map tab

@@
 library(sf)
-```
-
-```{r testModel}
-#
-# model1 <- lm(data = mwdata, perc_school ~ state + land_in_farms)
 ```
@@
 saveRDS(clean_data, file = "my_data.rds")
+saveRDS(clean_data, file = "testing123/my_data.rds")
 
@@
 
-```{r mapDataset}
+```{r mapDataset, cache=TRUE}
 
@@
 
+saveRDS(mwData, file = "final/mwData.rds")
+
 ```
@@
 
-```{r shapefiles}
+```{r shapefiles, cache=TRUE}
 
@@
 
-```{r gisTemplate}
+```{r gisTemplate, cache=TRUE}
 # might also want to add a filter for state??
@@
 # only change these two lines!
-year <- 1870
-myvar <- "Average Farm Value"
+year <- 1950
 
 #choices for myvar will be determined by: 
-appPlan <- allData %>%
-  filter(YEAR == "1870") %>%
+appPlan <- mwData %>%
+  filter(YEAR == year) %>%
   select_if(~ !all(is.na(.))) %>%
-  select(-STATEA:-COUNTY) %>%
-  select(-multiplier)
+  select(-STATEA:-COUNTY)
+
+# myvar options will be determined by column names of appPlan & displayed via dropdown widget
+myvar <- "Irrigated Acres"
 
@@
 
-```{r experimentalModels}
+```{r experimentalModels, eval = FALSE}
 
@@
 
-```{r farmTrends}
+```{r farmTrends, eval = FALSE}
 # this takes foever--cache it
@@
 
+tabPanel("Maps",
+
+                                 pageWithSidebar(
+
+                                     headerPanel("Midwest Map Explorer"),
+
+                                     sidebarPanel(
+                                         uiOutput("choose_dataset"),
+
+                                         uiOutput("choose_columns"),
+                                         br(),
+                                         a(href = "https://gist.github.com/4211337", "Source code")
+                                     ),
+
+                                     mainPanel(
+                                         tabsetPanel(id = "tabsMain",
+                                                     tabPanel("Map",
+                                         leafletOutput("map1")
+                                     ),
+                                     tabPanel("Interesting Observations",
+                                              # Don't write in the first p()
+                                              # Looks bad
+                                              p(""),
+                                              p("Write some stuff!")))
+                                 ))),
+
 ```

### 2020-04-26 final/app.R

@@ -27,7 +27,114 @@ library(rmapshaper)
 library(sf)
 
 
-data1800s <- readRDS("my_data.rds")
+
+# Prep the shapefiles
+# Now save the gis data with short names for sake of the shiny app
+# hopefully this will work...
+
+dsn1870 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870"
+layer1870 <- "US_county_1870_conflated"
+dsn1880 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1880"
+layer1880 <- "US_county_1880_conflated"
+dsn1890 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1890"
+layer1890 <- "US_county_1890_conflated"
+dsn1900 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1900"
+layer1900 <- "US_county_1900_conflated"
+dsn1910 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1910"
+layer1910 <- "US_county_1910_conflated"
+dsn1920 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1920"
+layer1920 <- "US_county_1920_conflated"
+dsn1930 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1930"
+layer1930 <- "US_county_1930_conflated"
+dsn1940 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1940"
+layer1940 <- "US_county_1940_conflated"
+dsn1950 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1950"
+layer1950 <- "US_county_1950_conflated"
+dsn1960 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1960"
+layer1960 <- "US_county_1960_conflated"
+
+# Make a tibble of all the shapefiles that you'll need
+shapefiles <- tibble(Year = c(1870, 1880, 1890, 1900, 1910, 1920, 1930, 1940, 1950, 1960), 
+                     Dsn = c(dsn1870, dsn1880, dsn1890, dsn1900, dsn1910, dsn1920, dsn1930, 
+                             dsn1940, dsn1950, dsn1960),
[…]

### 2020-04-26 testing123/app.R

@@ -132,9 +132,10 @@ Census_1960 <- x %>% filter(YEAR == "1960") %>%
 
 data_sets <- c("Census_1870", "Census_1880", "Census_1890", "Census_1900", "Census_1910", "Census_1920", "Census_1930", "Census_1940", "Census_1950", "Census_1960")
 
-ui <- fluidPage(pageWithSidebar(
+ui <- fluidPage(
+    pageWithSidebar(
     
-    headerPanel(""),
+    headerPanel("Midwest Map Explorer"),
     
     sidebarPanel(
         uiOutput("choose_dataset"),
@@ -154,7 +155,7 @@ server <- (function(input, output) {
     
     # Drop-down selection box for which data set
     output$choose_dataset <- renderUI({
-        selectInput("dataset", "Year", as.list(data_sets))
+        selectInput("dataset", "Choose Census Year:", as.list(data_sets))
     })
     
     # Check boxes
@@ -165,20 +166,19 @@ server <- (function(input, output) {
         
         # Get the data set with the appropriate name
         dat <- get(input$dataset)
-        # Make the first and second columns, "YEAR" and "STATE", null so that users don't pick them
-        # I can't just get rid of these columns altogether because I need them for my plots!
-        names(dat)[1] <- "NULL"
-        names(dat)[2] <- "NULL"
-        colnames <- names(dat)
         
-        # Create the checkboxes and select them all by default
-        selectInput("columns", "Choose columns", 
+        # Get rid of "YEAR" and "STATE" so that users don't pick them and make senseless maps
+
+        colnames1 <- names(dat)
+        colnames <- colnames1[!colnames1 %in% "YEAR" & !colnames1 %in% "STATE"]
+        
+        # Create the dropdown menu and select all columns by default
+        selectInput("columns", "Choose Variable:", 
                            choices  = colnames)
     })
     
  
-    
-    # My Output
+    # Map output
     output$map1 <- renderLeaflet({
         # If missing input, return to avoid error later in function
         if(is.null(input$dataset))

### 2020-04-26 final/app.R
Map comparison (layout needs work)

@@ -175,7 +175,7 @@ ui <- fluidPage(
                                                  
                                                  # this tells the UI what plot to put 
                                                  
-                                                 plotOutput("agPlot")  
+                                                 plotOutput("agPlot")
                                         )))),
                tabPanel("Models",
                         
@@ -184,33 +184,23 @@ ui <- fluidPage(
                         p("Make some models! (See notes)"),
                         gt_output("model1")
                ),
-               
+         
                
                tabPanel("Maps",
                         titlePanel("Midwest Map Explorer"),
                         p("For best results, view in fullscreen mode."),
                         p("Maps may take up to 30 seconds to load...but they're worth it!"),
-                   mainPanel(
-                         tabsetPanel(id = "tabsMain",
-                                                     tabPanel("Map",
-                                                              pageWithSidebar(
+                                sidebarLayout(
                                                                   
-                                                                  headerPanel(""),
-                                                                  
-                                                                  sidebarPanel(
-                                                                      uiOutput("choose_dataset"),
-                                                                      
-                                                                      uiOutput("choose_columns"),
+                                      sidebarPanel(
+                                            uiOutput("choose_dataset"),
+                                            uiOutput("choose_columns"),
+                                            uiOutput("choose_columns2"),
                                                                       br(),
                                                                   ),
-                                         mainPanel(leafletOutput("map1"))
+                                         mainPanel(leafletOutput("map1"),
+                                                   leafletOutput("map2"))
                                      )),
-                                     tabPanel("Interesting Observations",
-                                              # Don't write in the first p()
-                                              # Looks bad
-                                              p(""),
-                                              p("Write some stuff!")))
-                                 )),
                
                tabPanel("About",
                         
@@ -354,6 +344,100 @@ server <- (function(input, output, session) {
             addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
[…]

### 2020-04-26 final/app.R
Commit working updates (all tabs in place)

@@ -7,6 +7,8 @@
 #    http://shiny.rstudio.com/
 #
 
+# LOAD THE REQUIRED PACKAGES
+
 library(shiny)
 library(haven)
 library(tidyverse)
@@ -19,6 +21,10 @@ library(shinythemes)
 library(plotly)
 library(leaflet)
 library(broom)
+library(patchwork)
+library(magick)
+library(giphyr)
+
 
 # packages for mapping
 library(leaflet)
@@ -27,66 +33,21 @@ library(rmapshaper)
 library(sf)
 
 
+# READ IN THE DATA
 
-# Prep the shapefiles
-# Now save the gis data with short names for sake of the shiny app
-# hopefully this will work...
+# Read in the shapefiles
 
-dsn1870 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870"
-layer1870 <- "US_county_1870_conflated"
-dsn1880 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1880"
-layer1880 <- "US_county_1880_conflated"
-dsn1890 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1890"
-layer1890 <- "US_county_1890_conflated"
-dsn1900 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1900"
-layer1900 <- "US_county_1900_conflated"
-dsn1910 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1910"
-layer1910 <- "US_county_1910_conflated"
-dsn1920 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1920"
-layer1920 <- "US_county_1920_conflated"
-dsn1930 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1930"
-layer1930 <- "US_county_1930_conflated"
-dsn1940 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1940"
-layer1940 <- "US_county_1940_conflated"
-dsn1950 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1950"
-layer1950 <- "US_county_1950_conflated"
[…]

### 2020-04-26 final/app.R
Commit working app with all tabs

@@ -147,7 +147,8 @@ ui <- fluidPage(
                         # Create a spot for the plot
                         mainPanel(
                             tabsetPanel(id = "tabsMain",
-                                        tabPanel("Cows and Education",
+                                        
+                                        tabPanel("Implements",
                                                  # Generate a row with a sidebar
                                                  sidebarLayout(      
                                                      
@@ -159,13 +160,46 @@ ui <- fluidPage(
                                                                      choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
                                                                      multiple = TRUE)),
                                                      helpText("*Enter help text*")
+                                                 )
+                                        ),
+                                        
+                                        tabPanel("Cattle",
+                                                 # Generate a row with a sidebar
+                                                 sidebarLayout(      
+                                                     
+                                                     # Define the sidebar with one input
+                                                     # comprehensive options:
+                                                     # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
+                                                     sidebarPanel(
+                                                         selectInput("cowYear", "Year:", 
+                                                                     choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
+                                                                     multiple = TRUE)),
+                                                     helpText("*Enter help text*")
                                                  ),
                                                  
                                                  # this tells the UI what plot to put 
                                                  
                                                  plotOutput("agPlot"),
                                                  p("Number of cows might be seen as a stand in for farm size. Is it?")
-                                        )))),
+                                        ),
+                                        
+                                        tabPanel("Gender",
+                                                 # Generate a row with a sidebar
+                                                 sidebarLayout(      
+                                                     
+                                                     # Define the sidebar with one input
+                                                     # comprehensive options:
+                                                     # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
+                                                     sidebarPanel(
+                                                         selectInput("Year", "Year:", 
[…]

### 2020-04-26 ms_8.Rmd
Commit with maleedu gt, no plot

@@
 library(gganimate)
+library(patchwork)
+library(magick)
+library(webshot)
+webshot::install_phantomjs()
 
@@
          female = B3E002, enrolled = B2A001+B2A002+B2A003+B2A004+B2A005, urban_pop = B2J001, 
-         pct_urban = B2J001/B18001*101, illiterate = NA, acres_farms = B3K001, 
+         pct_urban = B2J001/B13001*100, illiterate = NA, acres_farms = B3K001, 
          acres_irrigated = NA, pct_irrigated = NA) %>%
@@
                        data1900j, data1910j, data1920j, data1930j, data1940j, data1950j, data1960j)) %>%
-  mutate(pct_male = male/pop*100, avg_value_farm = value_farm/n_farm, cows_per_farm = cows_per_person, 
-         avg_value_implements = value_implements/n_farm) %>%
+  mutate(pct_male = male/pop*100, avg_value_farm = value_farm/n_farm, cows_per_farm = cows_per_person, avg_value_implements = value_implements/n_farm, 
+         pct_urban = ifelse(pct_urban > 100, NA, pct_urban)) %>%
 
@@
                      layer1930, layer1940, layer1950, layer1960))
+
+saveRDS(shapefiles, file = "final/shapefiles.rds")
 ```
@@
          )
+
+saveRDS(maleedu_model, file = "final/maleedu_model.rds")
+
 # Incorporate
@@
 
-# couldn't get gtsave to save to the "final" subdirectory
-# will need to copy to "final" manually??
-gtsave(data = maleedu_gt, filename = "maleedu_gt.html")
+# save the gt table as a png file so that you can use magick on it with a plot
+
+gtsave(data = maleedu_gt, filename = "final/maleedu_gt.png")
 
@@
        title = "Percent Male Coefficients Over Time",
-       subtitle = "Measuring the effect of being male on school enrollment")
+       subtitle = "Measuring the effect of being male\non school enrollment")
 
@@
 
-# copy to "final"
-gtsave(data = cowedu_gt, filename = "cowedu_gt.html")
+gtsave(data = cowedu_gt, filename = "final/cowedu_gt.png")
 
@@
 
+```{r magick}
+
+# Combined plots using magick
+
+# Farmland acreage & number of farms plots
+
+mgif_1a <- image_read("final/plot1a.gif")
+mgif_1b <- image_read("final/plot1b.gif")
+
+acre_gif <- image_append(c(mgif_1a[1], mgif_1b[1]))
+for(i in 2:100){
+  combined <- image_append(c(mgif_1a[i], mgif_1b[i]))
+  acre_gif <- c(acre_gif, combined)
+}
+
+anim_save("final/acre_gif.gif", acre_gif)
+
+# Urbanization population plots
+
+mgif_3 <- image_read("final/plot3.gif")
+mgif_4 <- image_read("final/plot4.gif")
+
+pop_gif <- image_append(c(mgif_3[1], mgif_4[1]))
+for(i in 2:100){
+  combined <- image_append(c(mgif_3[i], mgif_4[i]))
+  pop_gif <- c(pop_gif, combined)
+}
+
+anim_save("final/pop_gif.gif", pop_gif)
+
+```
+
 ### SHINY FINAL
@@
 
-# Paste here after
+# This works
+
+library(shiny)
+library(haven)
+library(tidyverse)
+library(gt)
+library(devtools)
+library(tibble)
+library(foreign)
+library(scales)
+library(shinythemes)
+library(plotly)
+library(leaflet)
+library(broom)
+
+# packages for mapping
+library(leaflet)
+library(spdplyr)
+library(rmapshaper)
+library(sf)
+
+# Prep the shapefiles
+# Now save the gis data with short names for sake of the shiny app
+# hopefully this will work...
+
+dsn1870 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870"
+layer1870 <- "US_county_1870_conflated"
+dsn1880 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1880"
+layer1880 <- "US_county_1880_conflated"
+dsn1890 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1890"
+layer1890 <- "US_county_1890_conflated"
+dsn1900 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1900"
+layer1900 <- "US_county_1900_conflated"
+dsn1910 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1910"
+layer1910 <- "US_county_1910_conflated"
+dsn1920 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1920"
+layer1920 <- "US_county_1920_conflated"
+dsn1930 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1930"
+layer1930 <- "US_county_1930_conflated"
+dsn1940 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1940"
+layer1940 <- "US_county_1940_conflated"
+dsn1950 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1950"
+layer1950 <- "US_county_1950_conflated"
+dsn1960 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1960"
+layer1960 <- "US_county_1960_conflated"
+
+# Make a tibble of all the shapefiles that you'll need
+shapefiles <- tibble(Year = c(1870, 1880, 1890, 1900, 1910, 1920, 1930, 1940, 1950, 1960), 
+                     Dsn = c(dsn1870, dsn1880, dsn1890, dsn1900, dsn1910, dsn1920, dsn1930, 
+                             dsn1940, dsn1950, dsn1960),
+                     Layer = c(layer1870, layer1880, layer1890, layer1900, layer1910, layer1920, 
+                               layer1930, layer1940, layer1950, layer1960))
+
+x <- readRDS("my_data.rds") %>%
+    mutate("Value of Farms" = as.numeric(value_farm), "Average Farm Value" = as.numeric(avg_value_farm), 
+           "Value of Implements" = as.numeric(value_implements), 
+           "Average Implement Value" = as.numeric(avg_value_implements), 
+           "Average Farm Size" = as.numeric(avg_size), "Number of Farms" = as.numeric(n_farm), 
+           "Acres of Farmland" = as.numeric(acres_farms), "Irrigated Acres" = as.numeric(acres_irrigated), 
+           "Percentage of Farmland Irrigated" = as.numeric(pct_irrigated), "Number of Cows" = as.numeric(cows),
+           "Number of Cows per Farm" = as.numeric(cows_per_farm), "Number of Horses" = as.numeric(horses), 
+           "Total Population" = as.numeric(pop), "Male Population" = as.numeric(male), 
+           "Female Population" = as.numeric(female), "Male Percentage of Population" = as.numeric(pct_male),
+           "Total School Enrollment" = as.numeric(enrolled), 
+           "Percentage School Enrollment" = as.numeric(pct_enrolled), 
+           "Urban Population" = as.numeric(urban_pop), 
+           "Percent of Population in Urban Areas" = as.numeric(pct_urban), 
+           "Illiterate Population" = as.numeric(illiterate), 
+           "Illiteracy Rate" = as.numeric(pct_illiterate)) %>%
+    select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, 
+           "Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
+           "Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
+           "Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
+           "Number of Horses", "Total Population", "Male Population", "Female Population", 
+           "Male Percentage of Population", "Total School Enrollment", 
+           "Percentage School Enrollment", "Urban Population", 
+           "Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")
+
+# Now subset by year
+# Really should have saved these as Rds files. Will do that later
+Census_1870 <- x %>% filter(YEAR == "1870") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1880 <- x %>% filter(YEAR == "1880") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1890 <- x %>% filter(YEAR == "1890") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1900 <- x %>% filter(YEAR == "1900") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1910 <- x %>% filter(YEAR == "1910") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1920 <- x %>% filter(YEAR == "1920") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1930 <- x %>% filter(YEAR == "1930") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1940 <- x %>% filter(YEAR == "1940") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1950 <- x %>% filter(YEAR == "1950") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1960 <- x %>% filter(YEAR == "1960") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+data_sets <- c("Census_1870", "Census_1880", "Census_1890", "Census_1900", "Census_1910", "Census_1920", "Census_1930", "Census_1940", "Census_1950", "Census_1960")
+
+# Everything above this line should be condensed
+#####################
+
+allData <- readRDS("my_data.rds")
+
+# Input
+
+ui <- fluidPage(
+
+    # flatly is a nice theme to make the tabs look put together
+
+    # Application title
+
+    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),
+
+               # remember that tab panel is for what shows up in the tab, and title panel is 
+               # for what shows up at the top of the page within the tab
+
+               tabPanel("Intro to Midwestern Agriculture",
+
+                        # Give the page a title
+                        titlePanel("Relationship Between Farm Value and Education, by County"),
+
+                        # Generate a row with a sidebar
+                        sidebarLayout(
+
+                            # Define the sidebar with one input
+                            # comprehensive options:
+                            # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
+                            sidebarPanel(
+                                selectInput("Year", "Year:", 
+                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
+                                            multiple = TRUE)),
+                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
+                        ),
+
+                        # Create a spot for the plot
+                        mainPanel(
+                            tabsetPanel(id = "tabsMain",
+                                        tabPanel("Plot",
+
+                                                 # this tells the UI what plot to put
+
+                                                 plotOutput("agPlot")  
+                                        )))),
+               tabPanel("Models",
+
+                        # Give the page a title
+                        titlePanel("Relationship Between Farm Value and Education, by County"),
+                        p("Make some models! (See notes)"),
+                        gt_output("model1")
+               ),
 
                tabPanel("Maps",
-
-                                 pageWithSidebar(
-
-                                     headerPanel("Midwest Map Explorer"),
-
-                                     sidebarPanel(
-                                         uiOutput("choose_dataset"),
-
-                                         uiOutput("choose_columns"),
-                                         br(),
-                                         a(href = "https://gist.github.com/4211337", "Source code")
-                                     ),
-
+                        titlePanel("Midwest Map Explorer"),
+                        p("For best results, view in fullscreen mode."),
+                        p("Maps may take up to 30 seconds to load...but they're worth it!"),
                    mainPanel(
@@
                                                      tabPanel("Map",
-                                         leafletOutput("map1")
+                                                              pageWithSidebar(
+
+                                                                  headerPanel(""),
+
+                                                                  sidebarPanel(
+                                                                      uiOutput("choose_dataset"),
+
+                                                                      uiOutput("choose_columns"),
+                                                                      br(),
                                                                   ),
+                                         mainPanel(leafletOutput("map1"))
+                                     )),
                                      tabPanel("Interesting Observations",
@@
                                               p("Write some stuff!")))
-                                 ))),
+                                 )),
+
+               tabPanel("About",
+
+                        # Give the page a title
+                        titlePanel("About this project"),
+                        p("My name is Julia Englebert.
+I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights.
+Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
+                        br(),
+                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States.
+By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century.
+I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960.
+My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org.
+IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960.
+All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
+                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
+                        br(),
+                        p("To see more of my work, please visit my Github account:"),
+                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
+                        br(),
+                        titlePanel("Citations"),
+                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles.
+IPUMS National Historical Geographic Information System: Version 14.0 [Database].
+Minneapolis, MN: IPUMS.
+2019.
+http://doi.org/10.18128/D050.V14.0"),
+                        p("The College of William and Mary and the Minnesota Population Center.
+School Attendance Boundary Information System (SABINS): Version 1.0.
+Minneapolis, MN: University of Minnesota 2011.")
+               )
+               )
+    )
+
+# use plotly to allow hover text to appear on plots, giving more info about the point
+
+server <- (function(input, output, session) {
+
+    # Fill in the spot we created for a plot
+    output$agPlot <- renderPlot({
+
+        # Render a plot
+        allData %>%
+            filter(YEAR %in% input$Year) %>%
+            mutate(YEAR = as.factor(YEAR)) %>%
+            group_by(YEAR) %>%
+            ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
+            geom_point(alpha = 0.1) +
+            geom_smooth(method = "lm", se = FALSE) +
+            scale_color_viridis_d() +
+            ylim(0, 45) +
+            scale_x_log10(labels = comma) +
+            labs(title = "Cows and Education",
+                 subtitle = "Relationship between the number of cows per farm\nand school enrollment, by county",
+                 x = "Average Cows per Farm",
+                 y = "Percent of Population Enrolled in School") +
+            theme_classic()
+    })
+
+    output$choose_dataset <- renderUI({
+        selectInput("dataset", "Choose Census Year:", as.list(data_sets))
+    })
+
+    # Check boxes
+    output$choose_columns <- renderUI({
+        # If missing input, return to avoid error later in function
+        if(is.null(input$dataset))
+            return()
+
+        # Get the data set with the appropriate name
+        dat <- get(input$dataset)
+
+        # Get rid of "YEAR" and "STATE" so that users don't pick them and make senseless maps
+
+        colnames1 <- names(dat)
+        colnames <- colnames1[!colnames1 %in% "YEAR" & !colnames1 %in% "STATE"]
+
+        # Create the dropdown menu and select all columns by default
+        selectInput("columns", "Choose Variable:", 
+                    choices  = colnames)
+    })
+
+    # Map output
+    output$map1 <- renderLeaflet({
+        # If missing input, return to avoid error later in function
+        if(is.null(input$dataset))
+            return()
+
+        # Get the data set
+        dat <- get(input$dataset)
+
+        # Make sure columns are correct for data set (when data set changes, the
+        # columns will initially be for the previous data set)
+        if (is.null(input$columns) || !(input$columns %in% names(dat)))
+            return()
+
+        year <- dat %>% slice(1) %>% pull(YEAR)
+
+        # Keep the selected columns
+        dat <- dat[, input$columns, drop = FALSE]
+
+        myvar <- input$columns
+
+        datayear <- mwData %>%
+            # use get() to use the stored variable
+            # idk why, but it doesn't work without this
+            # the variable must be converted to a double in the original mwData dataset
+            mutate(myvariable = (get(myvar))) %>%
+            filter(YEAR == year, myvariable > 0) %>%
+            select(GISJOIN, myvariable)
+
+        # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
+        # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
+        # layer is the filename of the .shp file inside the
+        # folder dsn points to.
+
+        dsnyear <- shapefiles %>% 
+            filter(Year == year) %>% 
+            pull(Dsn)
+        layeryear <- shapefiles %>% 
+            filter(Year == year) %>% 
+            pull(Layer)
+
+        countyyear <- sf::st_read(dsn = dsnyear,
+                                  layer = layeryear)
+
+        countyyear <-
+            countyyear %>%
+            merge(datayear, "GISJOIN")
+
+        # Set projection of tracts dataset to `projection` required by leaflet
+
+        countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
+
+        # Condense size of data for faster processing
+
+        countyyear <- rmapshaper::ms_simplify(countyyear)
+
+        # Set palette color
+
+        pal <- colorNumeric("viridis", NULL)
+
+        #  Plot the data
+
+        # use shiny to add a title, which will be be equivalent to myvar
+
+        leaflet(countyyear) %>%
+            addTiles() %>%
+            addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
+                        fillColor = ~pal(myvariable)) %>%
+            addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
+
+    })
+
+    output$model1 <- render_gt({
+
+        # experimental model
+
+        model1
+
+    })
+})
+
+# Run the application
+
+shinyApp(ui = ui, server = server)
 
 ```
+
+# Ignore
+
+```{r ignore, eval=FALSE}
+data1870a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds16_1870_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1870b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds17_1870_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+# Join the two datasets by STATEA and COUNTYA, which are the state and county codes
+
+data1870j <- full_join(data1870a, data1870b, by = c("STATEA", "COUNTYA")) %>%
+
+  # Mutate to get rid of the .x suffix on variables
+
+  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
+
+  # Mutate to rename varialbes of interest to uniform, comprehensible names
+  # Some existing varialbes are combined to create the desired variables
+  # See codebook
+
+  mutate(value_farm = AJV001, n_farm = AJS001, cows =  AJ0003+AJ0005, horses = AJ0001, 
+         avg_size = (AJT001*2 + AJT002*6 + AJT003*15 + AJT004*35 + AJT005*75 + AJT006*300 + AJT007*750 + 
+                       AJT008*1200)/AKP001, value_implements = AKL001, pop = AJ3001, male = AKD001, 
+         female = AKD002, enrolled = ALG001, urban_pop = AKE001, pct_urban = AKE001/AJ3001*100, 
+         illiterate = AJ7001, acres_farms = AJU001+ AJU002+AJU003,
+
+         # Even though irrigation data doesn't exist for this year, I'm filling the columns in with NA
+         # Later years will have irrigation data
+
+         acres_irrigated = NA, pct_irrigated = NA) %>%
+
+  # Mutate again to make new variables out of the ones you just created in the last mutate
+
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+
+  # Select the desired variables
+
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm,
+         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female,
+         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
+
+# Now the dataset is much more manageable!
+
+# Start on the next year, 1880, repeating the exact same process
+# Lots of NA values due to scant data this year
+
+data1880a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds22_1880_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1880b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds23_1880_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1880j <- full_join(data1880a, data1880b, by = c("STATEA", "COUNTYA")) %>%
+  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
+  mutate(value_farm = APN001+APO001+APQ001, n_farm = AOQ001, cows =  AOG004+AOG005, horses = AOG001, value_implements = APO001, pop = AOT001, male = AP1001, female = AP1002, 
+         enrolled = NA, urban_pop = AO4001, pct_urban = AO4001/AOT001*100, illiterate = NA, 
+         acres_farms = AOR001, acres_irrigated = NA, pct_irrigated = NA) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
+         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
+
+# Start on 1890 data
+
+data1890a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds26_1890_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1890b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds27_1890_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1890j <- full_join(data1890a, data1890b, by = c("STATEA", "COUNTYA")) %>%
+  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
+  mutate(value_farm = AUK001+AUK002+AUK003, n_farm = AS9001, cows = ASY002, horses = ASY003, value_implements = AUK002, pop = AUM001, male = AV0001+AV0003+AV0005+AV0007, 
+         female = AV0002+AV0004+AV0006+AV0008, enrolled = AVK005+AVK006+AVK007+AVK008, urban_pop = AUU001, 
+         pct_urban = AUU001/AUM001*100, illiterate = NA, acres_farms = AUJ001+AUJ002, 
+         acres_irrigated = AT2001, pct_irrigated = AT2001/(AUJ001+AUJ002)*100) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
+         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
+
+# Before I mutated all of the data, running this code showed that all .x values are the same 
+# as all .y values, so it was safe to eliminate the double columns as long as "problems" returned 0
+# And it did!
+
+#problems <- bind_rows(data1870j, data1880j, data1890j) %>% 
+  #select(STATEA, COUNTYA, GISJOIN.x, GISJOIN.y, YEAR.x, YEAR.y, STATE.x, STATE.y, 
+         #COUNTY.x, COUNTY.y, AREANAME.x, AREANAME.y) %>%
+  #mutate(same = ifelse(GISJOIN.x == GISJOIN.y & YEAR.x == YEAR.y & STATE.x == STATE.y & 
+                         #COUNTY.x == COUNTY.y & AREANAME.x == AREANAME.y, TRUE, FALSE)) %>%
+  #filter(same == "FALSE") %>%
+  #nrow()
+
+# Read in the 1900s data
+
+# Only do decennial census data
+# Intermittent years only have AF12001: estimated population
+
+# 1900
+
+data1900a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds30_1900_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1900b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds31_1900_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1900j <- full_join(data1900a, data1900b, by = c("STATEA", "COUNTYA")) %>%
+  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
+  mutate(value_farm = AWV001, n_farm = AW3001, cows = AW7001, horses = AW7002, value_implements = AWW003, pop = AWS001, male = AZ0001,female = AZ0001, 
+         enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = AYS001+AYS002+AYS003+AYS004+AYS005, 
+         acres_farms = AWT001, acres_irrigated = AX8001, pct_irrigated = AX8001/AWT001*100) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
+         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
+
+# 1910
+
+data1910a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds36_1910_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1910b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds37_1910_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1910j <- full_join(data1910a, data1910b, by = c("STATEA", "COUNTYA")) %>%
+  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
+  mutate(value_farm = A0F001, n_farm = A1H001, cows = A1F001+A1F002, horses = NA, 
+         avg_size = A21001/A1H001, value_implements = A0Q005, pop = A0E001, male = NA, female = NA, 
+         enrolled = A4F001, urban_pop = A36001, pct_urban = A36001/A0E001*100, illiterate = A38001, 
+         acres_farms = A21001, acres_irrigated = A0G001, pct_irrigated = A0G001/A21001*100) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
+         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
+
+#1920
+
+data1920a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds43_1920_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME =  as.character(AREANAME))
+
+data1920b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1920_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1920c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds210_1920_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1920j <- full_join(data1920a, data1920b, by = c("STATEA", "COUNTYA"))
+data1920j <- full_join(data1920j, data1920c, by = c("STATEA", "COUNTYA"))
+data1920j <- data1920j %>%
+  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
+  mutate(value_farm = A8P001+A8P002+A8P003+A8P004, n_farm = AB3P001, cows = NA, horses = NA, 
+         avg_size = AB3T001/AB3P001, value_implements = A8P003, pop = A81001, male = A8B001, 
+         female = A8B002, enrolled = A7N001+A7N002+A7N003+A7N004, urban_pop = A80001, 
+         pct_urban = A80001/A81001*100, illiterate = A7U001, acres_farms = AB3T001, 
+         acres_irrigated = AB6L001, pct_irrigated = AB6L002/AB3T001*100) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
+         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
+
+#1930
+# 1930 has more datasets available than previous years
+
+data1930a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds212_1930_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1930b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1930_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1930c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds52_1930_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1930d <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds53_1930_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1930e <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds54_1930_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1930f <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds55_1930_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1930j <- full_join(data1930a, data1930b, by = c("STATEA", "COUNTYA"))
+data1930j <- full_join(data1930j, data1930c, by = c("STATEA", "COUNTYA"))
+data1930j <- full_join(data1930j, data1930d, by = c("STATEA", "COUNTYA"))
+data1930j <- full_join(data1930j, data1930e, by = c("STATEA", "COUNTYA"))
+data1930j <- full_join(data1930j, data1930f, by = c("STATEA", "COUNTYA"))
+
+data1930j <- data1930j %>%
+  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
+  mutate(value_farm = BFD001+BFD002+BFD004+ACEO001, n_farm = AB95001, cows = NA, horses = NA, value_implements = BFD004, pop = BCX001, male = BC3001, 
+         female = BC3002, enrolled = BDT001+BDT002+BDT003+BDT004, urban_pop = BDX001, 
+         pct_urban = BDX001/BDP001*100, illiterate = BDV001, acres_farms = AB98001, 
+         acres_irrigated = ACEX002, pct_irrigated = ACEX002/AB98001*100) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
+         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
+
+# 1940
+
+data1940a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1940_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1940b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds77_1940_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1940c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds78_1940_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1940j <- full_join(data1940a, data1940b, by = c("STATEA", "COUNTYA"))
+data1940j <- full_join(data1940j, data1940c, by = c("STATEA", "COUNTYA"))
+data1940j <- data1940j %>%
+  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
+
+  # value_farm is missing the land & livestock components!!
+  # this variable only includings buildings and implements
+
+  mutate(value_farm = BWD001+BWD002, n_farm = BY1001, cows = NA, horses = NA, value_implements = BWD002, pop = BVU001, male = BVY001, 
+         female = BVY002, enrolled = BWU001+BWU002+BWU003+BWU004+BWU005, urban_pop = NA, 
+         pct_urban = NA, illiterate = NA, acres_farms = BY3001, 
+         acres_irrigated = NA, pct_irrigated = NA) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, 
+         pct_illiterate = illiterate/pop*100) %>%
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
+         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
+
+# 1950
+
+data1950a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds83_1950_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME =  as.character(AREANAME))
+
+data1950b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds84_1950_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1950j <- full_join(data1950a, data1950b, by = c("STATEA", "COUNTYA")) %>%
+  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x, 
+         B18001 = as.numeric(B18001)) %>%
+
+  # value_farm is missing the implements & livestock components!!
+  # This variable only includings buildings and land
+
+  mutate(value_farm = B3M001, n_farm = B3Z001, cows = NA, horses = NA, value_implements = NA, pop = B13001, male = B3E001, 
+         female = B3E002, enrolled = B2A001+B2A002+B2A003+B2A004+B2A005, urban_pop = B2J001, 
+         pct_urban = B2J001/B13001*100, illiterate = NA, acres_farms = B3K001, 
+         acres_irrigated = NA, pct_irrigated = NA) %>%
+  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
+         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
+
+#1960
+# The data are SPARSE yikes
+
+data1960a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds89_1960_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME =  as.character(AREANAME))
+
+data1960b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds90_1960_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1960c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds91_1960_county.csv")) %>%
+  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
+         AREANAME = as.character(AREANAME))
+
+data1960j <- full_join(data1960a, data1960b, by = c("STATEA", "COUNTYA"))
+data1960j <- full_join(data1960j, data1960c, by = c("STATEA", "COUNTYA"))
+data1960j <- data1960j %>%
+  mutate(value_farm = NA, n_farm = NA, cows = NA, horses = NA, 
+         avg_size = NA, value_implements = NA, pop = B47001, male = B5H001, 
+         female = B5H002, enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = NA, 
+         acres_farms = NA, acres_irrigated = NA, pct_irrigated = NA) %>%
+  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
+         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
+         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
+
+```
+
+```{r}
+allData_inflated <- as_tibble(bind_rows(data1870j, data1880j, data1890j, 
+                       data1900j, data1910j, data1920j, data1930j, data1940j, data1950j, data1960j)) %>%
+  mutate(pct_male = male/pop*100, avg_value_farm = value_farm/n_farm, cows_per_farm = cows_per_person, 
+         avg_value_implements = value_implements/n_farm, pct_urban = ifelse(pct_urban > 100, NA, pct_urban)) %>%
+
+  # I mislabeled cows_per_farm as cows_per_person, so I'm doing damage control now
+
+  select(!cows_per_person) %>%
+  filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Dakota Territory" | 
+           STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | 
+           STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | 
+           STATE == "Indiana" | STATE == "Ohio",
+
+         # Get rid of the weird infinite values in this column
+
+         ! is.infinite(cows_per_farm))
+
+# Now adjust everything for inflation
+# I made a spreadsheet based on this website: 
+# https://www.officialdata.org/us/inflation/1960?amount=1
+
+inflation <- read_excel("inflation.xlsx")
+
+x <- left_join(allData_inflated, inflation, by = c("YEAR" = "year")) %>%
+  # Use the multiplier column to adjust for inflation
+  mutate(value_farm = value_farm*multiplier, value_implements = value_implements*multiplier) %>%
+  mutate(avg_value_farm = value_farm/n_farm, avg_value_implements = value_implements/n_farm) %>%
+  mutate("Value of Farms" = as.numeric(value_farm), "Average Farm Value" = as.numeric(avg_value_farm), 
+         "Value of Implements" = as.numeric(value_implements), 
+         "Average Implement Value" = as.numeric(avg_value_implements), 
+         "Average Farm Size" = as.numeric(avg_size), "Number of Farms" = as.numeric(n_farm), 
+         "Acres of Farmland" = as.numeric(acres_farms), "Irrigated Acres" = as.numeric(acres_irrigated), 
+         "Percentage of Farmland Irrigated" = as.numeric(pct_irrigated), "Number of Cows" = as.numeric(cows),
+         "Number of Cows per Farm" = as.numeric(cows_per_farm), "Number of Horses" = as.numeric(horses), 
+         "Total Population" = as.numeric(pop), "Male Population" = as.numeric(male), 
+         "Female Population" = as.numeric(female), "Male Percentage of Population" = as.numeric(pct_male),
+         "Total School Enrollment" = as.numeric(enrolled), 
+         "Percentage School Enrollment" = as.numeric(pct_enrolled), 
+         "Urban Population" = as.numeric(urban_pop), 
+         "Percent of Population in Urban Areas" = as.numeric(pct_urban), 
+         "Illiterate Population" = as.numeric(illiterate), 
+         "Illiteracy Rate" = as.numeric(pct_illiterate)) %>%
+  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, 
+         "Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
+                   "Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
+                   "Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
+                   "Number of Horses", "Total Population", "Male Population", "Female Population", 
+                   "Male Percentage of Population", "Total School Enrollment", 
+                   "Percentage School Enrollment", "Urban Population", 
+                   "Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")
+
+# saving this for later
+# will need in shiny app
+#sliderOptions <- c("Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
+                   #"Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
+                   #"Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
+                   #"Number of Horses", "Total Population", "Male Population", "Female Population", 
+                   #"Male Percentage of Population", "Total School Enrollment", 
+                   #"Percentage School Enrollment", "Urban Population", 
+                   #"Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")
+
+```
+
+```{r}
+Census_1870 <- x %>% filter(YEAR == "1870") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1880 <- x %>% filter(YEAR == "1880") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1890 <- x %>% filter(YEAR == "1890") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1900 <- x %>% filter(YEAR == "1900") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1910 <- x %>% filter(YEAR == "1910") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1920 <- x %>% filter(YEAR == "1920") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1930 <- x %>% filter(YEAR == "1930") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1940 <- x %>% filter(YEAR == "1940") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1950 <- x %>% filter(YEAR == "1950") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1960 <- x %>% filter(YEAR == "1960") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+saveRDS(Census_1870, file = "final/1870.rds")
+saveRDS(Census_1880, file = "final/1880.rds")
+saveRDS(Census_1890, file = "final/1890.rds")
+saveRDS(Census_1900, file = "final/1900.rds")
+saveRDS(Census_1910, file = "final/1910.rds")
+saveRDS(Census_1920, file = "final/1920.rds")
+saveRDS(Census_1930, file = "final/1930.rds")
+saveRDS(Census_1940, file = "final/1940.rds")
+saveRDS(Census_1950, file = "final/1950.rds")
+saveRDS(Census_1960, file = "final/1960.rds")
+
+```
+
+# Working App
+
+```{r workingApp}
+# START THE APP
+
+# Input
+
+ui <- fluidPage(
+
+    # flatly is a nice theme to make the tabs look put together
+
+    # Application title
+
+    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),
+
+               # remember that tab panel is for what shows up in the tab, and title panel is 
+               # for what shows up at the top of the page within the tab
+
+               tabPanel("Intro to Midwestern Agriculture",
+
+                        # Give the page a title
+                        titlePanel("Relationship Between Farm Value and Education, by County"),
+
+                        # Generate a row with a sidebar
+                        sidebarLayout(
+
+                            # Define the sidebar with one input
+                            # comprehensive options:
+                            # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
+                            sidebarPanel(
+                                selectInput("Year", "Year:", 
+                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
+                                            multiple = TRUE)),
+                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
+                        ),
+
+                        # Create a spot for the plot
+                        mainPanel(
+                            tabsetPanel(id = "tabsMain",
+                                        tabPanel("Plot",
+
+                                                 # this tells the UI what plot to put
+
+                                                 plotOutput("agPlot")
+                                        )))),
+               tabPanel("Models",
+
+                        # Give the page a title
+                        titlePanel("Relationship Between Farm Value and Education, by County"),
+                        p("Make some models! (See notes)"),
+                        gt_output("model1")
+               ),
+
+               tabPanel("Maps",
+                        titlePanel("Midwest Map Explorer"),
+                        p("For best results, view in fullscreen mode."),
+                        p("Maps may take up to 30 seconds to load...but they're worth it!"),
+                                sidebarLayout(
+
+                                      sidebarPanel(
+                                            uiOutput("choose_dataset"),
+                                            uiOutput("choose_columns"),
+                                            uiOutput("choose_columns2"),
+                                                                      br(),
+                                                                  ),
+                                         mainPanel(leafletOutput("map1"),
+                                                   leafletOutput("map2"))
+                                     )),
+
+               tabPanel("About",
+
+                        # Give the page a title
+                        titlePanel("About this project"),
+                        p("My name is Julia Englebert.
+I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights.
+Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
+                        br(),
+                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States.
+By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century.
+I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960.
+My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org.
+IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960.
+All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
+                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
+                        br(),
+                        p("To see more of my work, please visit my Github account:"),
+                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
+                        br(),
+                        titlePanel("Citations"),
+                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles.
+IPUMS National Historical Geographic Information System: Version 14.0 [Database].
+Minneapolis, MN: IPUMS.
+2019.
+http://doi.org/10.18128/D050.V14.0"),
+                        p("The College of William and Mary and the Minnesota Population Center.
+School Attendance Boundary Information System (SABINS): Version 1.0.
+Minneapolis, MN: University of Minnesota 2011.")
+               )
+               )
+    )
+
+# use plotly to allow hover text to appear on plots, giving more info about the point
+
+server <- (function(input, output, session) {
+
+    # Fill in the spot we created for a plot
+    output$agPlot <- renderPlot({
+
+        # Render a plot
+        allData %>%
+            filter(YEAR %in% input$Year) %>%
+            mutate(YEAR = as.factor(YEAR)) %>%
+            group_by(YEAR) %>%
+            ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
+            geom_point(alpha = 0.1) +
+            geom_smooth(method = "lm", se = FALSE) +
+            scale_color_viridis_d() +
+            ylim(0, 45) +
+            scale_x_log10(labels = comma) +
+            labs(title = "Cows and Education",
+                 subtitle = "Relationship between the number of cows per farm\nand school enrollment, by county",
+                 x = "Average Cows per Farm",
+                 y = "Percent of Population Enrolled in School") +
+            theme_classic()
+    })
+
+    output$choose_dataset <- renderUI({
+        selectInput("dataset", "Choose Census Year:", as.list(data_sets))
+    })
+
+    # Check boxes
+    output$choose_columns <- renderUI({
+        # If missing input, return to avoid error later in function
+        if(is.null(input$dataset))
+            return()
+
+        # Get the data set with the appropriate name
+        dat <- get(input$dataset)
+
+        # Get rid of "YEAR" and "STATE" so that users don't pick them and make senseless maps
+
+        colnames1 <- names(dat)
+        colnames <- colnames1[!colnames1 %in% "YEAR" & !colnames1 %in% "STATE"]
+
+        # Create the dropdown menu and select all columns by default
+        selectizeInput("columns", "Choose Variable:", 
+                    choices  = colnames)
+    })
+
+    # Map output
+    output$map1 <- renderLeaflet({
+        # If missing input, return to avoid error later in function
+        if(is.null(input$dataset))
+            return()
+
+        # Get the data set
+        dat <- get(input$dataset)
+
+        # Make sure columns are correct for data set (when data set changes, the
+        # columns will initially be for the previous data set)
+        if (is.null(input$columns) || !(input$columns %in% names(dat)))
+            return()
+
+        year <- dat %>% slice(1) %>% pull(YEAR)
+
+        # Keep the selected columns
+        dat <- dat[, input$columns, drop = FALSE]
+
+        myvar <- input$columns
+
+        datayear <- mwData %>%
+            # use get() to use the stored variable
+            # idk why, but it doesn't work without this
+            # the variable must be converted to a double in the original mwData dataset
+            mutate(myvariable = (get(myvar))) %>%
+            filter(YEAR == year, myvariable > 0) %>%
+            select(GISJOIN, myvariable)
+
+        # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
+        # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
+        # layer is the filename of the .shp file inside the
+        # folder dsn points to.
+
+        dsnyear <- shapefiles %>% 
+            filter(Year == year) %>% 
+            pull(Dsn)
+        layeryear <- shapefiles %>% 
+            filter(Year == year) %>% 
+            pull(Layer)
+
+        countyyear <- sf::st_read(dsn = dsnyear,
+                                  layer = layeryear)
+
+        countyyear <-
+            countyyear %>%
+            merge(datayear, "GISJOIN")
+
+        # Set projection of tracts dataset to `projection` required by leaflet
+
+        countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
+
+        # Condense size of data for faster processing
+
+        countyyear <- rmapshaper::ms_simplify(countyyear)
+
+        # Set palette color
+
+        pal <- colorNumeric("viridis", NULL)
+
+        #  Plot the data
+
+        # use shiny to add a title, which will be be equivalent to myvar
+
+        leaflet(countyyear) %>%
+            addTiles() %>%
+            addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
+                        fillColor = ~pal(myvariable)) %>%
+            addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
+
+    })
+
+    # map2 dropdown menu
+    output$choose_columns2 <- renderUI({
+        # If missing input, return to avoid error later in function
+        if(is.null(input$dataset))
+            return()
+
+        # Get the data set with the appropriate name
+        dat <- get(input$dataset)
+
+        # Get rid of "YEAR" and "STATE" so that users don't pick them and make senseless maps
+
+        colnames1 <- names(dat)
+        colnames <- colnames1[!colnames1 %in% "YEAR" & !colnames1 %in% "STATE"]
+
+        # Create the dropdown menu and select all columns by default
+        selectizeInput("columns2", "Choose Variable to Compare:", 
+                    choices  = colnames,
+                    options = list(
+                        placeholder = 'Variables...',
+                        onInitialize = I('function() { this.setValue(""); }')))
+    })
+
+    output$map2 <- renderLeaflet({
+        # If missing input, return to avoid error later in function
+        if(is.null(input$dataset))
+            return()
+
+        # Get the data set
+        dat <- get(input$dataset)
+
+        # Make sure columns are correct for data set (when data set changes, the
+        # columns will initially be for the previous data set)
+        if (is.null(input$columns2) || !(input$columns2 %in% names(dat)))
+            return()
+
+        year <- dat %>% slice(1) %>% pull(YEAR)
+
+        # Keep the selected columns
+        dat <- dat[, input$columns2, drop = FALSE]
+
+        myvar <- input$columns2
+
+        datayear <- mwData %>%
+            # use get() to use the stored variable
+            # idk why, but it doesn't work without this
+            # the variable must be converted to a double in the original mwData dataset
+            mutate(myvariable = (get(myvar))) %>%
+            filter(YEAR == year, myvariable > 0) %>%
+            select(GISJOIN, myvariable)
+
+        # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
+        # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
+        # layer is the filename of the .shp file inside the
+        # folder dsn points to.
+
+        dsnyear <- shapefiles %>% 
+            filter(Year == year) %>% 
+            pull(Dsn)
+        layeryear <- shapefiles %>% 
+            filter(Year == year) %>% 
+            pull(Layer)
+
+        countyyear <- sf::st_read(dsn = dsnyear,
+                                  layer = layeryear)
+
+        countyyear <-
+            countyyear %>%
+            merge(datayear, "GISJOIN")
+
+        # Set projection of tracts dataset to `projection` required by leaflet
+
+        countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
+
+        # Condense size of data for faster processing
+
+        countyyear <- rmapshaper::ms_simplify(countyyear)
+
+        # Set palette color
+
+        pal <- colorNumeric("viridis", NULL)
+
+        #  Plot the data
+
+        # use shiny to add a title, which will be be equivalent to myvar
+
+        leaflet(countyyear) %>%
+            addTiles() %>%
+            addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
+                        fillColor = ~pal(myvariable)) %>%
+            addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
+
+    })
+
+    output$model1 <- render_gt({
+
+        # experimental model
+
+        model1
+
+    })
+})
+
+# Run the application
+
+shinyApp(ui = ui, server = server)
+
+```

### 2020-04-26 final/app.R

@@ -24,6 +24,7 @@ library(broom)
 library(patchwork)
 library(magick)
 library(giphyr)
+library(gridExtra)
 
 
 # packages for mapping
@@ -34,6 +35,10 @@ library(sf)
 
 
 # READ IN THE DATA
+
+# Read in the model for gt and plot
+
+maleedu_model <- readRDS("maleedu_model.rds") 
 
 # Read in the shapefiles
 
@@ -135,7 +140,7 @@ ui <- fluidPage(
                                                  mainPanel(
                                                      imageOutput("plot2")
                                                  ))
-                                        ))),
+                            ))),
                
                
                tabPanel("Models",
@@ -143,12 +148,12 @@ ui <- fluidPage(
                         # Give the page a title
                         titlePanel("Agriculture and Education, by County"),
                         
-                         
+                        
                         # Create a spot for the plot
-                        mainPanel(
+                        mainPanel("",
                             tabsetPanel(id = "tabsMain",
                                         
-                                        tabPanel("Implements",
+                                        tabPanel("Implements and Education",
                                                  # Generate a row with a sidebar
                                                  sidebarLayout(      
                                                      
@@ -163,7 +168,7 @@ ui <- fluidPage(
                                                  )
                                         ),
                                         
-                                        tabPanel("Cattle",
+                                        tabPanel("Cattle and Education",
                                                  # Generate a row with a sidebar
                                                  sidebarLayout(      
                                                      
@@ -183,62 +188,55 @@ ui <- fluidPage(
                                                  p("Number of cows might be seen as a stand in for farm size. Is it?")
                                         ),
                                         
-                                        tabPanel("Gender",
-                                                 # Generate a row with a sidebar
-                                                 sidebarLayout(      
-                                                     
-                                                     # Define the sidebar with one input
-                                                     # comprehensive options:
-                                                     # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
-                                                     sidebarPanel(
-                                                         selectInput("Year", "Year:", 
[…]

### 2020-04-27 ms_8.Rmd
Commit backup app code

@@
 ```{r workingApp}
+#
+# This is a Shiny web application. You can run the application by clicking
+# the 'Run App' button above.
+#
+# Find out more about building applications with Shiny here:
+#
+#    http://shiny.rstudio.com/
+#
+
+# LOAD THE REQUIRED PACKAGES
+
+library(shiny)
+library(haven)
+library(tidyverse)
+library(gt)
+library(devtools)
+library(tibble)
+library(foreign)
+library(scales)
+library(shinythemes)
+library(plotly)
+library(leaflet)
+library(broom)
+library(patchwork)
+library(magick)
+library(giphyr)
+library(gridExtra)
+
+# packages for mapping
+library(leaflet)
+library(spdplyr)
+library(rmapshaper)
+library(sf)
+
+# READ IN THE DATA
+
+# Read in the model for gt and plot
+
+maleedu_model <- readRDS("maleedu_model.rds")
+
+# Read in the shapefiles
+
+shapefiles <- readRDS(file = "shapefiles.rds")
+
+# Read in the dataset made specifically for mapping
+
+x <- readRDS("mwData.rds")
+
+# Now subset by year
+
+# I tried saving these as Rds files as well, but I kept getting a weird error
+# Might be a bug, according to SO
+Census_1870 <- x %>% filter(YEAR == "1870") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1880 <- x %>% filter(YEAR == "1880") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1890 <- x %>% filter(YEAR == "1890") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1900 <- x %>% filter(YEAR == "1900") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1910 <- x %>% filter(YEAR == "1910") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1920 <- x %>% filter(YEAR == "1920") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1930 <- x %>% filter(YEAR == "1930") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1940 <- x %>% filter(YEAR == "1940") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1950 <- x %>% filter(YEAR == "1950") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+Census_1960 <- x %>% filter(YEAR == "1960") %>%
+    select_if(~ !all(is.na(.))) %>%
+    select(-STATEA:-GISJOIN & -COUNTY)
+
+# If I put "Census_1870" = "1870", would only the year show up in dropdown menu?
+data_sets <- c("1870" = "Census_1870", "1880" = "Census_1880", "1890" = "Census_1890", "1900" = "Census_1900", "1910" = "Census_1910", "1920" = "Census_1920",  "1930" ="Census_1930",  "1940" ="Census_1940",  "1950" ="Census_1950",  "1960" ="Census_1960")
+
+# Read in the dataset for maps and models
+
+allData <- readRDS("my_data.rds")
+
 # START THE APP
@@
 
-               tabPanel("Intro to Midwestern Agriculture",
+               tabPanel("Introduction",
+                        titlePanel("A Brief Overview of Midwest Agriculture"),
+
+                        # Create a spot for the plot
+                        # Why is the text on top of the image?? :(
+                        mainPanel(
+                            tabsetPanel(id = "tabsMain",
+                                        tabPanel("Farms & Acreage",
+
+                                                 # Create a fluid row in order to comment next to the gifs
+                                                 # column specifications might need to be changed
+
+                                                 fluidRow(column(width = 3, textOutput("acre_com")),
+                                                          column(width = 9, align = "center",
+                                                                 imageOutput("acre_gif")))
+
+                                        ),
+                                        tabPanel("Urbanization",
+                                                 # Generate a row with a sidebar
+                                                 fluidRow(column(width = 3, textOutput("pop_com")),
+                                                              column(width = 9, align = "center",
+                                                                     imageOutput("pop_gif")))
+                                                 ),
+                                        tabPanel("Mechanization",
+                                                fluidRow(column(width = 3, textOutput("plot2_com")),
+                                                              column(width = 9, align = "center",
+                                                                     imageOutput("plot2")))
+                            )))),
+
+               tabPanel("Models",
 
                         # Give the page a title
-                        titlePanel("Relationship Between Farm Value and Education, by County"),
-
+                        titlePanel("Agriculture and Education, by County"),
+
+                        # Create a spot for the plot
+                        mainPanel("",
+                                  tabsetPanel(id = "tabsMain",
+
+                                              tabPanel("Implements and Education",
                                                        # Generate a row with a sidebar
@@
                                                            sidebarPanel(
-                                selectInput("Year", "Year:", 
+                                                               selectInput("impYear", "Year:", 
+                                                                           choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940"), 
+                                                                           multiple = TRUE)),
+                                                           helpText("*Enter help text*")
+                                                       ),
+
+                                                       # this tells the UI what plot to put
+
+                                                       plotOutput("impedu_plot"),
+                                                       p("say some stuff")
+                                              ),
+
+                                              tabPanel("Cattle and Education",
+                                                       # Generate a row with a sidebar
+                                                       sidebarLayout(
+
+                                                           # Define the sidebar with one input
+                                                           # comprehensive options:
+                                                           # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
+                                                           sidebarPanel(
+                                                               selectInput("cowYear", "Year:", 
                                                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
                                                                            multiple = TRUE)),
-                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
+                                                           helpText("*Enter help text*")
                                                        ),
 
-                        # Create a spot for the plot
-                        mainPanel(
-                            tabsetPanel(id = "tabsMain",
-                                        tabPanel("Plot",
-
                                                        # this tells the UI what plot to put
 
-                                                 plotOutput("agPlot")
-                                        )))),
-               tabPanel("Models",
-
-                        # Give the page a title
-                        titlePanel("Relationship Between Farm Value and Education, by County"),
-                        p("Make some models! (See notes)"),
-                        gt_output("model1")
+                                                       plotOutput("agPlot"),
+                                                       p("Number of cows might be seen as a stand in for farm size. Is it?")
                                               ),
+
+                                              tabPanel("Gender and Education",
+
+                                                       #mainPanel("",
+                                                                 #gt_output("maleedu_gt"))
+
+                                                       fluidRow(column(width = 6, plotOutput("maleedu_plot")),
+                                                                column(width = 6, #align = "center",
+                                                                       gt_output("maleedu_gt")))
+
+                                              )
+                                  )
+                        )),
 
@@
 Minneapolis, MN: University of Minnesota 2011.")
+               ),
+
+               tabPanel("Codebook",
+                        titlePanel("How were variables measured for this project?"),
+                        p("Put some info here. An interactive gt table would be nice if you have time."))
     )
 )
-    )
 
@@
 server <- (function(input, output, session) {
+
+    output$acre_com <- renderText({
+
+        "This is the commentary about farm acreage"
+    })
+
+    output$acre_gif <- renderImage({
+        list(src = "acre_gif.gif")
+
+    }, deleteFile = FALSE)
+
+    output$pop_com <- renderText({
+
+        "This is the commentary about population"
+    })
+
+    output$pop_gif <- renderImage({
+        list(src = "pop_gif.gif")
+
+    }, deleteFile = FALSE)
+
+    output$plot2_com <- renderText({
+
+        "This is the commentary about mechanization"
+    })
+
+    output$plot2 <- renderImage({
+        list(src = "plot2.gif")
+
+    }, deleteFile = FALSE)
+
+    # Implements plot
+
+    output$impedu_plot <- renderPlot({
+
+        allData %>%
+            filter(YEAR %in% input$impYear) %>%
+            mutate(YEAR = as.factor(YEAR)) %>%
+            mutate(YEAR = fct_explicit_na(YEAR, na_level = NA)) %>%
+            group_by(YEAR) %>%
+            select(YEAR, avg_value_implements, pct_enrolled) %>%
+            na.omit() %>%
+            ggplot(aes(x = avg_value_implements, y = pct_enrolled, color = YEAR)) + 
+            geom_point(alpha = 0.1) +
+            geom_smooth(method = "lm", se = FALSE) +
+            scale_fill_viridis_d(aesthetics = "color") +
+            ylim(0, 45) + 
+            labs(title = "Relationship Between Farm Implements and Education, by County",
+                 subtitle = "Midwestern United States", 
+                 x = "Average Value of Implements", 
+                 y = "Percent Population Enrolled in School") +
+            scale_x_log10(labels = comma) +
+            theme_classic()
+    })
 
@@
         allData %>%
-            filter(YEAR %in% input$Year) %>%
+            filter(YEAR %in% input$cowYear) %>%
             mutate(YEAR = as.factor(YEAR)) %>%
+            mutate(YEAR = fct_explicit_na(YEAR, na_level = NA)) %>%
             group_by(YEAR) %>%
+            select(YEAR, cows_per_farm, pct_enrolled) %>%
+            na.omit() %>%
             ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
@@
 
+    # maleedu plot and gt
+
+    output$maleedu_plot <- renderPlot({
+        maleedu_model %>%
+            ungroup() %>%
+            select(YEAR, pct_male_coef, pct_male_upper, pct_male_lower) %>%
+            mutate(YEAR = as.factor(YEAR)) %>%
+            group_by(YEAR) %>%
+            ggplot(aes(x = YEAR, y = pct_male_coef)) + 
+            geom_errorbar(ymin = maleedu_model$pct_male_lower, ymax = maleedu_model$pct_male_upper, color = "#453781FF") + 
+            geom_point(color = "#440154FF") +
+            theme_classic() +
+            ylim(-1.5, 0.5) + 
+            labs(x = "Year",
+                 y = "Coefficient",
+                 title = "Percent Male Coefficients Over Time",
+                 subtitle = "Measuring the effect of being male\non school enrollment")
+    })
+
+    output$maleedu_gt <- render_gt({
+        maleedu_model%>%
+            ungroup() %>%
+
+            # Select the columns needed, as per the instructions
+
+            select(-data:-reg_results) %>%
+
+            gt() %>%
+
+            # Add the desired titles, subtitles, and column labels
+
+            tab_header(title = "Effect of Gender and Number of Farms on Enrollment",
+                       subtitle = "With 95% Confidence Interval") %>%
+            tab_spanner(label = "Farm Coefficients", columns = vars(n_farm_lower, n_farm_coef, n_farm_upper)) %>%
+            tab_spanner(label = "Percent Male Coefficients", columns = vars(pct_male_lower, pct_male_coef,pct_male_upper)) %>%
+            tab_spanner(label = "Interaction Coefficients", columns = vars(interaction_lower, interaction_coef, interaction_upper)) %>%
+            cols_label(YEAR = "Year", n_farm_coef	= "Estimate", n_farm_upper = "Upper", n_farm_lower = "Lower", pct_male_coef = "Estimate", pct_male_upper = "Upper", pct_male_lower = "Lower",	interaction_coef = "Estimate", interaction_upper = "Upper", interaction_lower = "Lower") %>%
+            # Round all values to two decimal places
+
+            fmt_number(columns = vars(n_farm_lower, n_farm_coef, n_farm_upper, pct_male_lower, pct_male_coef,pct_male_upper, interaction_lower, interaction_coef, interaction_upper), decimals = 4)
+
+    })
+
+    # Map output
+
     output$choose_dataset <- renderUI({

### 2020-04-27 final/app.R
Commit working fluid rows

@@ -123,24 +123,30 @@ ui <- fluidPage(
                         
                         
                         # Create a spot for the plot
+                        # Why is the text on top of the image?? :(
                         mainPanel(
                             tabsetPanel(id = "tabsMain",
                                         tabPanel("Farms & Acreage",
-                                                 # Generate a row with a sidebar
-                                                 mainPanel(
-                                                     imageOutput("acre_gif")
-                                                 )),
+                                                 
+                                                 # Create a fluid row in order to comment next to the gifs
+                                                 # column specifications might need to be changed
+                                                 
+                                                 fluidRow(column(width = 3, textOutput("acre_com")),
+                                                          column(width = 9, align = "center",
+                                                                 imageOutput("acre_gif")))
+                                          
+                                        ),
                                         tabPanel("Urbanization",
                                                  # Generate a row with a sidebar
-                                                 mainPanel(
-                                                     imageOutput("pop_gif")
-                                                 )),
+                                                 fluidRow(column(width = 3, textOutput("pop_com")),
+                                                              column(width = 9, align = "center",
+                                                                     imageOutput("pop_gif")))
+                                                 ),
                                         tabPanel("Mechanization",
-                                                 # Generate a row with a sidebar
-                                                 mainPanel(
-                                                     imageOutput("plot2")
-                                                 ))
-                            ))),
+                                                fluidRow(column(width = 3, textOutput("plot2_com")),
+                                                              column(width = 9, align = "center",
+                                                                     imageOutput("plot2")))
+                            )))),
                
                
                tabPanel("Models",
@@ -151,92 +157,99 @@ ui <- fluidPage(
                         
                         # Create a spot for the plot
                         mainPanel("",
-                            tabsetPanel(id = "tabsMain",
-                                        
-                                        tabPanel("Implements and Education",
-                                                 # Generate a row with a sidebar
-                                                 sidebarLayout(      
-                                                     
-                                                     # Define the sidebar with one input
-                                                     # comprehensive options:
-                                                     # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
-                                                     sidebarPanel(
-                                                         selectInput("Year", "Year:", 
-                                                                     choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
-                                                                     multiple = TRUE)),
-                                                     helpText("*Enter help text*")
[…]

### 2020-04-27 final/app.R
Commit with mtcars dataset in place of codebook (add later)

@@ -159,7 +159,7 @@ ui <- fluidPage(
                         mainPanel("",
                                   tabsetPanel(id = "tabsMain",
                                               
-                                              tabPanel("Implements and Education",
+                                              tabPanel("Implements",
                                                        # Generate a row with a sidebar
                                                        sidebarLayout(      
                                                            
@@ -179,8 +179,10 @@ ui <- fluidPage(
                                                        p("say some stuff")
                                               ),
                                               
-                                              tabPanel("Cattle and Education",
+                                              tabPanel("Cattle",
                                                        # Generate a row with a sidebar
+                                                       
+                                                       
                                                        sidebarLayout(      
                                                            
                                                            # Define the sidebar with one input
@@ -199,14 +201,17 @@ ui <- fluidPage(
                                                        p("Number of cows might be seen as a stand in for farm size. Is it?")
                                               ),
                                               
-                                              tabPanel("Gender and Education",
+                                              tabPanel("Gender",
                                                        
                                                        #mainPanel("",
                                                                  #gt_output("maleedu_gt"))
                                                        
                                                        fluidRow(column(width = 6, plotOutput("maleedu_plot")),
                                                                 column(width = 6, #align = "center",
-                                                                       gt_output("maleedu_gt")))
+                                                                       gt_output("maleedu_gt"))), 
+                                                       fluidRow(column(width = 6, textOutput("maletext1")),
+                                                                column(width = 6, #align = "center",
+                                                                       textOutput("maletext2")))
                                                        
                                               )
                                   )
@@ -248,7 +253,11 @@ ui <- fluidPage(
                
                tabPanel("Codebook",
                         titlePanel("How were variables measured for this project?"),
-                        p("Put some info here. An interactive gt table would be nice if you have time."))
+                        p("Choose a variable to find out!"), 
+                        varSelectInput("variables", "Variable:", mtcars, multiple = FALSE),
+                        helpText("*Enter help text*"),
+                        tableOutput("codebook")
+                        )
     )
 )
 
@@ -378,6 +387,17 @@ server <- (function(input, output, session) {
             
[…]

### 2020-04-27 final/app.R
Commit all plots & models (need commentary and codebook)

@@ -40,6 +40,10 @@ library(sf)
 
 maleedu_model <- readRDS("maleedu_model.rds") 
 
+cowedu_model <- readRDS("cowedu_model.rds")
+
+cows_size_model <- readRDS("cows_size_model.rds")
+
 # Read in the shapefiles
 
 shapefiles <- readRDS(file = "shapefiles.rds")
@@ -176,7 +180,10 @@ ui <- fluidPage(
                                                        # this tells the UI what plot to put 
                                                        
                                                        plotOutput("impedu_plot"),
-                                                       p("say some stuff")
+                                                       p("say some stuff"),
+                                                       gt_output("cowedu_gt"),
+                                                       p("comments")
+                                                       
                                               ),
                                               
                                               tabPanel("Cattle",
@@ -198,7 +205,9 @@ ui <- fluidPage(
                                                        # this tells the UI what plot to put 
                                                        
                                                        plotOutput("agPlot"),
-                                                       p("Number of cows might be seen as a stand in for farm size. Is it?")
+                                                       p("Number of cows might be seen as a stand in for farm size. Is it?"),
+                                                       gt_output("cows_size_gt"),
+                                                       p("comments")
                                               ),
                                               
                                               tabPanel("Gender",
@@ -322,6 +331,17 @@ server <- (function(input, output, session) {
             theme_classic()
     })
     
+    output$cowedu_gt <- render_gt({
+        cowedu_model %>%
+            select(term, conf.low, estimate, conf.high) %>%
+            gt() %>%
+            tab_header(title = "Effect Cows and Implements on Enrollment",
+                       subtitle = "With 95% Confidence Interval") %>%
+            cols_label(term = "Coefficient Name", conf.low = "Lower Bound", estimate = "Estimated Value", conf.high = "Upper Bound") %>%
+            fmt_number(columns = vars(conf.low, estimate, conf.high), decimals = 4)
+        
+    })
+    
     # Fill in the spot we created for a plot
     output$agPlot <- renderPlot({
         
@@ -344,6 +364,18 @@ server <- (function(input, output, session) {
                  x = "Average Cows per Farm",
                  y = "Percent of Population Enrolled in School") +
             theme_classic()
+    })
+    
+    
+    output$cows_size_gt <- render_gt({
[…]

### 2020-04-27 final/app.R
Commit welcome text and codebook

@@ -25,6 +25,7 @@ library(patchwork)
 library(magick)
 library(giphyr)
 library(gridExtra)
+library(readxl)
 
 
 # packages for mapping
@@ -43,6 +44,13 @@ maleedu_model <- readRDS("maleedu_model.rds")
 cowedu_model <- readRDS("cowedu_model.rds")
 
 cows_size_model <- readRDS("cows_size_model.rds")
+
+# Read in the codebook
+# This is an excel sheet that I made in order to condense the original codebooks provided by NHGIS
+
+codebook <- read_excel("mw_codebook.xlsx") %>% 
+    remove_rownames %>% 
+    column_to_rownames(var="Year")
 
 # Read in the shapefiles
 
@@ -98,7 +106,8 @@ Census_1960 <- x %>% filter(YEAR == "1960") %>%
     select(-STATEA:-GISJOIN & -COUNTY)
 
 
-# If I put "Census_1870" = "1870", would only the year show up in dropdown menu?
+
+# Combine the datasets into a vector, renaming them by only the year 
 data_sets <- c("1870" = "Census_1870", "1880" = "Census_1880", "1890" = "Census_1890", "1900" = "Census_1900", "1910" = "Census_1910", "1920" = "Census_1920",  "1930" ="Census_1930",  "1940" ="Census_1940",  "1950" ="Census_1950",  "1960" ="Census_1960")
 
 
@@ -122,33 +131,33 @@ ui <- fluidPage(
                # remember that tab panel is for what shows up in the tab, and title panel is 
                # for what shows up at the top of the page within the tab                  
                
-               tabPanel("Introduction",
-                        titlePanel("A Brief Overview of Midwest Agriculture"),
-                        
+               tabPanel("Welcome",
+                        titlePanel("Midwest Agriculture in Brief"),
                         
                         # Create a spot for the plot
                         # Why is the text on top of the image?? :(
                         mainPanel(
+                            helpText("To understand the relationship between agriculture and education, it’s helpful to elucidate trends in agriculture over time. Click the tabs below to visualize major changes."),
                             tabsetPanel(id = "tabsMain",
                                         tabPanel("Farms & Acreage",
                                                  
                                                  # Create a fluid row in order to comment next to the gifs
                                                  # column specifications might need to be changed
                                                  
-                                                 fluidRow(column(width = 3, textOutput("acre_com")),
-                                                          column(width = 9, align = "center",
+                                                 fluidRow(column(width = 5, textOutput("acre_com")),
[…]

### 2020-04-27 ms_8.Rmd
Final commit of ms_8 draft (sans original datafiles)

@@
 # See implement plot (plot2) saved as gif
+# also use cowedu model
 
@@
 
+cows_size_model <- allData %>%
+  select(YEAR, pct_enrolled, cows_per_farm, avg_size) %>%
+  na.omit() %>%
+  lm(pct_enrolled ~ cows_per_farm*avg_size, data = .) %>%
+  tidy(conf.int = TRUE)
+
+saveRDS(cows_size_model, "final/cows_size_model.rds")
+
 # INCORPORATE
@@
   tidy(conf.int = TRUE)
+
+saveRDS(cowedu_model, "final/cowedu_model.rds")
 
@@
 
-saveRDS(Census_1870, file = "final/1870.rds")
-saveRDS(Census_1880, file = "final/1880.rds")
-saveRDS(Census_1890, file = "final/1890.rds")
-saveRDS(Census_1900, file = "final/1900.rds")
-saveRDS(Census_1910, file = "final/1910.rds")
-saveRDS(Census_1920, file = "final/1920.rds")
-saveRDS(Census_1930, file = "final/1930.rds")
-saveRDS(Census_1940, file = "final/1940.rds")
-saveRDS(Census_1950, file = "final/1950.rds")
-saveRDS(Census_1960, file = "final/1960.rds")
-
 ```

### 2020-04-27 final/app.R

@@ -183,15 +183,15 @@ ui <- fluidPage(
                                                                selectInput("impYear", "Year:", 
                                                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940"), 
                                                                            multiple = TRUE)),
-                                                           helpText("*Enter help text*")
+                                                           helpText("Select a year from the dropdown menu in order to explore the correlation between mechanization and school enrollment over time.")
                                                        ),
                                                        
                                                        # this tells the UI what plot to put 
                                                        
                                                        plotOutput("impedu_plot"),
-                                                       p("say some stuff"),
+                                                       p("The plot shows a positive correlation between school enrollment and implement value in 1870, but that correlation disappears in later years."),
                                                        gt_output("cowedu_gt"),
-                                                       p("comments")
+                                                       p("This linear regression model shows that an increase in average implement value has essentially no effect on school enrollment. Curiously, the number of cows per farm is strongly associated with higher school enrollment, as we will explore in the next tab.")
                                                        
                                               ),
                                               
@@ -208,15 +208,15 @@ ui <- fluidPage(
                                                                selectInput("cowYear", "Year:", 
                                                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
                                                                            multiple = TRUE)),
-                                                           helpText("*Enter help text*")
+                                                           helpText("Select a year from the dropdown menu in order to explore the correlation between cattle ownership and school enrollment over time.")
                                                        ),
                                                        
                                                        # this tells the UI what plot to put 
                                                        
                                                        plotOutput("agPlot"),
-                                                       p("Number of cows might be seen as a stand in for farm size. Is it?"),
+                                                       p("The plot shows a positive correlation between school enrollment and the average number of cows per farm for the years where data was available. The number of cows might simply be a confounding variable that acts as a stand-in for farm size, but as the table below will show, this does not seem to be the case."),
                                                        gt_output("cows_size_gt"),
-                                                       p("comments")
[…]

### 2020-04-28 ms_8.Rmd
Commit with map error fixed

--- 2020-04-26 ms_8.Rmd
@@
 library(tibble)
-library(foreign)
+#library(foreign)
 library(scales)
@@
 
-saveRDS(clean_data, file = "final/my_data.rds")
+# saveRDS(clean_data, file = "final/my_data.rds")
 saveRDS(clean_data, file = "my_data.rds")
@@
 
-saveRDS(mwData, file = "final/mwData.rds")
+#saveRDS(mwData, file = "final/mwData.rds")
 
@@
 
-saveRDS(shapefiles, file = "final/shapefiles.rds")
+#saveRDS(shapefiles, file = "final/shapefiles.rds")
 ```
@@
 # myvar options will be determined by column names of appPlan & displayed via dropdown widget
-myvar <- "Irrigated Acres"
+myvar <- "Value of Farms"
 
@@
 
-```{r experimentalModels, eval = FALSE}
+```{r experimentalModels, cache = TRUE}
 
@@
 
-ggsave("final/impedu_plot.png", plot = impedu_plot)
+#ggsave("final/impedu_plot.png", plot = impedu_plot)
 
 # See implement plot (plot2) saved as gif
+# also use cowedu model
 
@@
 
-saveRDS(maleedu_model, file = "final/maleedu_model.rds")
+#saveRDS(maleedu_model, file = "final/maleedu_model.rds")
 
@@
 
-gtsave(data = maleedu_gt, filename = "final/maleedu_gt.png")
+#gtsave(data = maleedu_gt, filename = "final/maleedu_gt.png")
 
@@
 
-ggsave("final/maleedu_plot.png", plot = maleedu_plot)
+#ggsave("final/maleedu_plot.png", plot = maleedu_plot)
+
+cows_size_model <- allData %>%
+  select(YEAR, pct_enrolled, cows_per_farm, avg_size) %>%
+  na.omit() %>%
+  lm(pct_enrolled ~ cows_per_farm*avg_size, data = .) %>%
+  tidy(conf.int = TRUE)
+
+#saveRDS(cows_size_model, "final/cows_size_model.rds")
 
@@
 
+#saveRDS(cowedu_model, "final/cowedu_model.rds")
+
 cowedu_gt <- cowedu_model %>%
@@
 
-gtsave(data = cowedu_gt, filename = "final/cowedu_gt.png")
+#gtsave(data = cowedu_gt, filename = "final/cowedu_gt.png")
 
@@
 
-ggsave("final/cowedu_plot.png", plot = cowedu_plot)
+#ggsave("final/cowedu_plot.png", plot = cowedu_plot)
 
@@
 
-```{r farmTrends, eval = FALSE}
+```{r farmTrends, cache=TRUE}
 # this takes foever--cache it
@@
 
-anim_save("final/plot1a.gif", plot1a)
+#anim_save("final/plot1a.gif", plot1a)
 
@@
 
-anim_save("final/plot1b.gif", plot1b)
+#anim_save("final/plot1b.gif", plot1b)
 
@@
 
-anim_save("final/plot2.gif", plot2)
+#anim_save("final/plot2.gif", plot2)
 
@@
 
-anim_save("final/plot3.gif", plot3)
+#anim_save("final/plot3.gif", plot3)
 
@@
 
-plot4_prepd <- full_join(plot4_prepb, plot4_prepc, by = c("YEAR")) %>%
-  select(YEAR, n_pop, n_urban_pop) %>%
-  mutate(YEAR = as.factor(YEAR)) %>%
+# I was having trouble with the foreign package, so I commented out all code that requires it
+
+#plot4_prepd <- full_join(plot4_prepb, plot4_prepc, by = c("YEAR")) %>%
+  #select(YEAR, n_pop, n_urban_pop) %>%
+  #mutate(YEAR = as.factor(YEAR)) %>%
   # convert to long format
-  melt(id="YEAR")
+  #melt(id="YEAR")
 
@@
 
-plot4 <- as_tibble(plot4_prepd) %>%
+#plot4 <- as_tibble(plot4_prepd) %>%
   # as.numeric is changing years to numbers 1:7
-  mutate(YEAR = as.numeric(YEAR), variable = as.character(variable), 
-         value = value/1000) %>%
-  ggplot(aes(YEAR, value, color = variable)) +
-  geom_line() +
-  scale_x_discrete(limits = limits4, labels = labels4) +
-  scale_fill_viridis_d(aesthetics = "color", breaks=c("n_urban_pop", "n_pop"), labels = c("Urban", "Rural")) +
-  labs(x = "Year",
-       y = "Population (in thousands)",
-       title = "Midwest Urbanization",
-       subtitle = "Rural and urban populations over Time",
-       color = "") +
-  theme_classic() +
-  geom_point(aes(group = seq_along(YEAR))) +
-  transition_reveal(YEAR)
-
-anim_save("final/plot4.gif", plot4)
+  #mutate(YEAR = as.numeric(YEAR), variable = as.character(variable), 
+         #value = value/1000) %>%
+  #ggplot(aes(YEAR, value, color = variable)) +
+  #geom_line() +
+  #scale_x_discrete(limits = limits4, labels = labels4) +
+  #scale_fill_viridis_d(aesthetics = "color", breaks=c("n_urban_pop", "n_pop"), labels = c("Urban", "Rural")) +
+  #labs(x = "Year",
+       #y = "Population (in thousands)",
+       #title = "Midwest Urbanization",
+       #subtitle = "Rural and urban populations over Time",
+       #color = "") +
+  #theme_classic() +
+  #geom_point(aes(group = seq_along(YEAR))) +
+  #transition_reveal(YEAR)
+
+#anim_save("final/plot4.gif", plot4)
 ```
 
+# magick code
+
 ```{r magick}
 
+```
 # Combined plots using magick
@@
 anim_save("final/pop_gif.gif", pop_gif)
-
-```
-
-### SHINY FINAL
-```{r shiny, eval=FALSE}
-
-# This works
-
-library(shiny)
-library(haven)
-library(tidyverse)
-library(gt)
-library(devtools)
-library(tibble)
-library(foreign)
-library(scales)
-library(shinythemes)
-library(plotly)
-library(leaflet)
-library(broom)
-
-# packages for mapping
-library(leaflet)
-library(spdplyr)
-library(rmapshaper)
-library(sf)
-
-# Prep the shapefiles
-# Now save the gis data with short names for sake of the shiny app
-# hopefully this will work...
-
-dsn1870 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870"
-layer1870 <- "US_county_1870_conflated"
-dsn1880 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1880"
-layer1880 <- "US_county_1880_conflated"
-dsn1890 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1890"
-layer1890 <- "US_county_1890_conflated"
-dsn1900 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1900"
-layer1900 <- "US_county_1900_conflated"
-dsn1910 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1910"
-layer1910 <- "US_county_1910_conflated"
-dsn1920 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1920"
-layer1920 <- "US_county_1920_conflated"
-dsn1930 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1930"
-layer1930 <- "US_county_1930_conflated"
-dsn1940 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1940"
-layer1940 <- "US_county_1940_conflated"
-dsn1950 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1950"
-layer1950 <- "US_county_1950_conflated"
-dsn1960 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1960"
-layer1960 <- "US_county_1960_conflated"
-
-# Make a tibble of all the shapefiles that you'll need
-shapefiles <- tibble(Year = c(1870, 1880, 1890, 1900, 1910, 1920, 1930, 1940, 1950, 1960), 
-                     Dsn = c(dsn1870, dsn1880, dsn1890, dsn1900, dsn1910, dsn1920, dsn1930, 
-                             dsn1940, dsn1950, dsn1960),
-                     Layer = c(layer1870, layer1880, layer1890, layer1900, layer1910, layer1920, 
-                               layer1930, layer1940, layer1950, layer1960))
-
-x <- readRDS("my_data.rds") %>%
-    mutate("Value of Farms" = as.numeric(value_farm), "Average Farm Value" = as.numeric(avg_value_farm), 
-           "Value of Implements" = as.numeric(value_implements), 
-           "Average Implement Value" = as.numeric(avg_value_implements), 
-           "Average Farm Size" = as.numeric(avg_size), "Number of Farms" = as.numeric(n_farm), 
-           "Acres of Farmland" = as.numeric(acres_farms), "Irrigated Acres" = as.numeric(acres_irrigated), 
-           "Percentage of Farmland Irrigated" = as.numeric(pct_irrigated), "Number of Cows" = as.numeric(cows),
-           "Number of Cows per Farm" = as.numeric(cows_per_farm), "Number of Horses" = as.numeric(horses), 
-           "Total Population" = as.numeric(pop), "Male Population" = as.numeric(male), 
-           "Female Population" = as.numeric(female), "Male Percentage of Population" = as.numeric(pct_male),
-           "Total School Enrollment" = as.numeric(enrolled), 
-           "Percentage School Enrollment" = as.numeric(pct_enrolled), 
-           "Urban Population" = as.numeric(urban_pop), 
-           "Percent of Population in Urban Areas" = as.numeric(pct_urban), 
-           "Illiterate Population" = as.numeric(illiterate), 
-           "Illiteracy Rate" = as.numeric(pct_illiterate)) %>%
-    select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, 
-           "Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
-           "Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
-           "Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
-           "Number of Horses", "Total Population", "Male Population", "Female Population", 
-           "Male Percentage of Population", "Total School Enrollment", 
-           "Percentage School Enrollment", "Urban Population", 
-           "Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")
-
-# Now subset by year
-# Really should have saved these as Rds files. Will do that later
-Census_1870 <- x %>% filter(YEAR == "1870") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1880 <- x %>% filter(YEAR == "1880") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1890 <- x %>% filter(YEAR == "1890") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1900 <- x %>% filter(YEAR == "1900") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1910 <- x %>% filter(YEAR == "1910") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1920 <- x %>% filter(YEAR == "1920") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1930 <- x %>% filter(YEAR == "1930") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1940 <- x %>% filter(YEAR == "1940") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1950 <- x %>% filter(YEAR == "1950") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1960 <- x %>% filter(YEAR == "1960") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-data_sets <- c("Census_1870", "Census_1880", "Census_1890", "Census_1900", "Census_1910", "Census_1920", "Census_1930", "Census_1940", "Census_1950", "Census_1960")
-
-# Everything above this line should be condensed
-#####################
-
-allData <- readRDS("my_data.rds")
-
-# Input
-
-ui <- fluidPage(
-
-    # flatly is a nice theme to make the tabs look put together
-
-    # Application title
-
-    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),
-
-               # remember that tab panel is for what shows up in the tab, and title panel is 
-               # for what shows up at the top of the page within the tab
-
-               tabPanel("Intro to Midwestern Agriculture",
-
-                        # Give the page a title
-                        titlePanel("Relationship Between Farm Value and Education, by County"),
-
-                        # Generate a row with a sidebar
-                        sidebarLayout(
-
-                            # Define the sidebar with one input
-                            # comprehensive options:
-                            # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
-                            sidebarPanel(
-                                selectInput("Year", "Year:", 
-                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
-                                            multiple = TRUE)),
-                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
-                        ),
-
-                        # Create a spot for the plot
-                        mainPanel(
-                            tabsetPanel(id = "tabsMain",
-                                        tabPanel("Plot",
-
-                                                 # this tells the UI what plot to put
-
-                                                 plotOutput("agPlot")  
-                                        )))),
-               tabPanel("Models",
-
-                        # Give the page a title
-                        titlePanel("Relationship Between Farm Value and Education, by County"),
-                        p("Make some models! (See notes)"),
-                        gt_output("model1")
-               ),
-
-               tabPanel("Maps",
-                        titlePanel("Midwest Map Explorer"),
-                        p("For best results, view in fullscreen mode."),
-                        p("Maps may take up to 30 seconds to load...but they're worth it!"),
-                   mainPanel(
-                         tabsetPanel(id = "tabsMain",
-                                                     tabPanel("Map",
-                                                              pageWithSidebar(
-
-                                                                  headerPanel(""),
-
-                                                                  sidebarPanel(
-                                                                      uiOutput("choose_dataset"),
-
-                                                                      uiOutput("choose_columns"),
-                                                                      br(),
-                                                                  ),
-                                         mainPanel(leafletOutput("map1"))
-                                     )),
-                                     tabPanel("Interesting Observations",
-                                              # Don't write in the first p()
-                                              # Looks bad
-                                              p(""),
-                                              p("Write some stuff!")))
-                                 )),
-
-               tabPanel("About",
-
-                        # Give the page a title
-                        titlePanel("About this project"),
-                        p("My name is Julia Englebert.
-I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights.
-Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
-                        br(),
-                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States.
-By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century.
-I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960.
-My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org.
-IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960.
-All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
-                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
-                        br(),
-                        p("To see more of my work, please visit my Github account:"),
-                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
-                        br(),
-                        titlePanel("Citations"),
-                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles.
-IPUMS National Historical Geographic Information System: Version 14.0 [Database].
-Minneapolis, MN: IPUMS.
-2019.
-http://doi.org/10.18128/D050.V14.0"),
-                        p("The College of William and Mary and the Minnesota Population Center.
-School Attendance Boundary Information System (SABINS): Version 1.0.
-Minneapolis, MN: University of Minnesota 2011.")
-               )
-               )
-    )
-
-# use plotly to allow hover text to appear on plots, giving more info about the point
-
-server <- (function(input, output, session) {
-
-    # Fill in the spot we created for a plot
-    output$agPlot <- renderPlot({
-
-        # Render a plot
-        allData %>%
-            filter(YEAR %in% input$Year) %>%
-            mutate(YEAR = as.factor(YEAR)) %>%
-            group_by(YEAR) %>%
-            ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
-            geom_point(alpha = 0.1) +
-            geom_smooth(method = "lm", se = FALSE) +
-            scale_color_viridis_d() +
-            ylim(0, 45) +
-            scale_x_log10(labels = comma) +
-            labs(title = "Cows and Education",
-                 subtitle = "Relationship between the number of cows per farm\nand school enrollment, by county",
-                 x = "Average Cows per Farm",
-                 y = "Percent of Population Enrolled in School") +
-            theme_classic()
-    })
-
-    output$choose_dataset <- renderUI({
-        selectInput("dataset", "Choose Census Year:", as.list(data_sets))
-    })
-
-    # Check boxes
-    output$choose_columns <- renderUI({
-        # If missing input, return to avoid error later in function
-        if(is.null(input$dataset))
-            return()
-
-        # Get the data set with the appropriate name
-        dat <- get(input$dataset)
-
-        # Get rid of "YEAR" and "STATE" so that users don't pick them and make senseless maps
-
-        colnames1 <- names(dat)
-        colnames <- colnames1[!colnames1 %in% "YEAR" & !colnames1 %in% "STATE"]
-
-        # Create the dropdown menu and select all columns by default
-        selectInput("columns", "Choose Variable:", 
-                    choices  = colnames)
-    })
-
-    # Map output
-    output$map1 <- renderLeaflet({
-        # If missing input, return to avoid error later in function
-        if(is.null(input$dataset))
-            return()
-
-        # Get the data set
-        dat <- get(input$dataset)
-
-        # Make sure columns are correct for data set (when data set changes, the
-        # columns will initially be for the previous data set)
-        if (is.null(input$columns) || !(input$columns %in% names(dat)))
-            return()
-
-        year <- dat %>% slice(1) %>% pull(YEAR)
-
-        # Keep the selected columns
-        dat <- dat[, input$columns, drop = FALSE]
-
-        myvar <- input$columns
-
-        datayear <- mwData %>%
-            # use get() to use the stored variable
-            # idk why, but it doesn't work without this
-            # the variable must be converted to a double in the original mwData dataset
-            mutate(myvariable = (get(myvar))) %>%
-            filter(YEAR == year, myvariable > 0) %>%
-            select(GISJOIN, myvariable)
-
-        # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
-        # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
-        # layer is the filename of the .shp file inside the
-        # folder dsn points to.
-
-        dsnyear <- shapefiles %>% 
-            filter(Year == year) %>% 
-            pull(Dsn)
-        layeryear <- shapefiles %>% 
-            filter(Year == year) %>% 
-            pull(Layer)
-
-        countyyear <- sf::st_read(dsn = dsnyear,
-                                  layer = layeryear)
-
-        countyyear <-
-            countyyear %>%
-            merge(datayear, "GISJOIN")
-
-        # Set projection of tracts dataset to `projection` required by leaflet
-
-        countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
-
-        # Condense size of data for faster processing
-
-        countyyear <- rmapshaper::ms_simplify(countyyear)
-
-        # Set palette color
-
-        pal <- colorNumeric("viridis", NULL)
-
-        #  Plot the data
-
-        # use shiny to add a title, which will be be equivalent to myvar
-
-        leaflet(countyyear) %>%
-            addTiles() %>%
-            addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
-                        fillColor = ~pal(myvariable)) %>%
-            addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
-
-    })
-
-    output$model1 <- render_gt({
-
-        # experimental model
-
-        model1
-
-    })
-})
-
-# Run the application
-
-shinyApp(ui = ui, server = server)
-
-```
-
-# Ignore
-
-```{r ignore, eval=FALSE}
-data1870a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds16_1870_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1870b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds17_1870_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-# Join the two datasets by STATEA and COUNTYA, which are the state and county codes
-
-data1870j <- full_join(data1870a, data1870b, by = c("STATEA", "COUNTYA")) %>%
-
-  # Mutate to get rid of the .x suffix on variables
-
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-
-  # Mutate to rename varialbes of interest to uniform, comprehensible names
-  # Some existing varialbes are combined to create the desired variables
-  # See codebook
-
-  mutate(value_farm = AJV001, n_farm = AJS001, cows =  AJ0003+AJ0005, horses = AJ0001, 
-         avg_size = (AJT001*2 + AJT002*6 + AJT003*15 + AJT004*35 + AJT005*75 + AJT006*300 + AJT007*750 + 
-                       AJT008*1200)/AKP001, value_implements = AKL001, pop = AJ3001, male = AKD001, 
-         female = AKD002, enrolled = ALG001, urban_pop = AKE001, pct_urban = AKE001/AJ3001*100, 
-         illiterate = AJ7001, acres_farms = AJU001+ AJU002+AJU003,
-
-         # Even though irrigation data doesn't exist for this year, I'm filling the columns in with NA
-         # Later years will have irrigation data
-
-         acres_irrigated = NA, pct_irrigated = NA) %>%
-
-  # Mutate again to make new variables out of the ones you just created in the last mutate
-
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
-
-  # Select the desired variables
-
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm,
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female,
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# Now the dataset is much more manageable!
-
-# Start on the next year, 1880, repeating the exact same process
-# Lots of NA values due to scant data this year
-
-data1880a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds22_1880_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1880b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds23_1880_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1880j <- full_join(data1880a, data1880b, by = c("STATEA", "COUNTYA")) %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = APN001+APO001+APQ001, n_farm = AOQ001, cows =  AOG004+AOG005, horses = AOG001, value_implements = APO001, pop = AOT001, male = AP1001, female = AP1002, 
-         enrolled = NA, urban_pop = AO4001, pct_urban = AO4001/AOT001*100, illiterate = NA, 
-         acres_farms = AOR001, acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# Start on 1890 data
-
-data1890a <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds26_1890_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1890b <- as_tibble(read.csv("raw-data/nhgis0002_csv/nhgis0002_ds27_1890_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1890j <- full_join(data1890a, data1890b, by = c("STATEA", "COUNTYA")) %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = AUK001+AUK002+AUK003, n_farm = AS9001, cows = ASY002, horses = ASY003, value_implements = AUK002, pop = AUM001, male = AV0001+AV0003+AV0005+AV0007, 
-         female = AV0002+AV0004+AV0006+AV0008, enrolled = AVK005+AVK006+AVK007+AVK008, urban_pop = AUU001, 
-         pct_urban = AUU001/AUM001*100, illiterate = NA, acres_farms = AUJ001+AUJ002, 
-         acres_irrigated = AT2001, pct_irrigated = AT2001/(AUJ001+AUJ002)*100) %>%
-  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# Before I mutated all of the data, running this code showed that all .x values are the same 
-# as all .y values, so it was safe to eliminate the double columns as long as "problems" returned 0
-# And it did!
-
-#problems <- bind_rows(data1870j, data1880j, data1890j) %>% 
-  #select(STATEA, COUNTYA, GISJOIN.x, GISJOIN.y, YEAR.x, YEAR.y, STATE.x, STATE.y, 
-         #COUNTY.x, COUNTY.y, AREANAME.x, AREANAME.y) %>%
-  #mutate(same = ifelse(GISJOIN.x == GISJOIN.y & YEAR.x == YEAR.y & STATE.x == STATE.y & 
-                         #COUNTY.x == COUNTY.y & AREANAME.x == AREANAME.y, TRUE, FALSE)) %>%
-  #filter(same == "FALSE") %>%
-  #nrow()
-
-# Read in the 1900s data
-
-# Only do decennial census data
-# Intermittent years only have AF12001: estimated population
-
-# 1900
-
-data1900a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds30_1900_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1900b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds31_1900_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1900j <- full_join(data1900a, data1900b, by = c("STATEA", "COUNTYA")) %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = AWV001, n_farm = AW3001, cows = AW7001, horses = AW7002, value_implements = AWW003, pop = AWS001, male = AZ0001,female = AZ0001, 
-         enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = AYS001+AYS002+AYS003+AYS004+AYS005, 
-         acres_farms = AWT001, acres_irrigated = AX8001, pct_irrigated = AX8001/AWT001*100) %>%
-  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# 1910
-
-data1910a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds36_1910_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1910b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds37_1910_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1910j <- full_join(data1910a, data1910b, by = c("STATEA", "COUNTYA")) %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = A0F001, n_farm = A1H001, cows = A1F001+A1F002, horses = NA, 
-         avg_size = A21001/A1H001, value_implements = A0Q005, pop = A0E001, male = NA, female = NA, 
-         enrolled = A4F001, urban_pop = A36001, pct_urban = A36001/A0E001*100, illiterate = A38001, 
-         acres_farms = A21001, acres_irrigated = A0G001, pct_irrigated = A0G001/A21001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-#1920
-
-data1920a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds43_1920_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME =  as.character(AREANAME))
-
-data1920b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1920_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1920c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds210_1920_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1920j <- full_join(data1920a, data1920b, by = c("STATEA", "COUNTYA"))
-data1920j <- full_join(data1920j, data1920c, by = c("STATEA", "COUNTYA"))
-data1920j <- data1920j %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = A8P001+A8P002+A8P003+A8P004, n_farm = AB3P001, cows = NA, horses = NA, 
-         avg_size = AB3T001/AB3P001, value_implements = A8P003, pop = A81001, male = A8B001, 
-         female = A8B002, enrolled = A7N001+A7N002+A7N003+A7N004, urban_pop = A80001, 
-         pct_urban = A80001/A81001*100, illiterate = A7U001, acres_farms = AB3T001, 
-         acres_irrigated = AB6L001, pct_irrigated = AB6L002/AB3T001*100) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-#1930
-# 1930 has more datasets available than previous years
-
-data1930a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds212_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds52_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930d <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds53_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930e <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds54_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930f <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds55_1930_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1930j <- full_join(data1930a, data1930b, by = c("STATEA", "COUNTYA"))
-data1930j <- full_join(data1930j, data1930c, by = c("STATEA", "COUNTYA"))
-data1930j <- full_join(data1930j, data1930d, by = c("STATEA", "COUNTYA"))
-data1930j <- full_join(data1930j, data1930e, by = c("STATEA", "COUNTYA"))
-data1930j <- full_join(data1930j, data1930f, by = c("STATEA", "COUNTYA"))
-
-data1930j <- data1930j %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-  mutate(value_farm = BFD001+BFD002+BFD004+ACEO001, n_farm = AB95001, cows = NA, horses = NA, value_implements = BFD004, pop = BCX001, male = BC3001, 
-         female = BC3002, enrolled = BDT001+BDT002+BDT003+BDT004, urban_pop = BDX001, 
-         pct_urban = BDX001/BDP001*100, illiterate = BDV001, acres_farms = AB98001, 
-         acres_irrigated = ACEX002, pct_irrigated = ACEX002/AB98001*100) %>%
-  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# 1940
-
-data1940a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds224_1940_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1940b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds77_1940_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1940c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds78_1940_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1940j <- full_join(data1940a, data1940b, by = c("STATEA", "COUNTYA"))
-data1940j <- full_join(data1940j, data1940c, by = c("STATEA", "COUNTYA"))
-data1940j <- data1940j %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x) %>%
-
-  # value_farm is missing the land & livestock components!!
-  # this variable only includings buildings and implements
-
-  mutate(value_farm = BWD001+BWD002, n_farm = BY1001, cows = NA, horses = NA, value_implements = BWD002, pop = BVU001, male = BVY001, 
-         female = BVY002, enrolled = BWU001+BWU002+BWU003+BWU004+BWU005, urban_pop = NA, 
-         pct_urban = NA, illiterate = NA, acres_farms = BY3001, 
-         acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, 
-         pct_illiterate = illiterate/pop*100) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-# 1950
-
-data1950a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds83_1950_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME =  as.character(AREANAME))
-
-data1950b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds84_1950_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1950j <- full_join(data1950a, data1950b, by = c("STATEA", "COUNTYA")) %>%
-  mutate(GISJOIN = GISJOIN.x, YEAR = YEAR.x, STATE = STATE.x, COUNTY = COUNTY.x, 
-         B18001 = as.numeric(B18001)) %>%
-
-  # value_farm is missing the implements & livestock components!!
-  # This variable only includings buildings and land
-
-  mutate(value_farm = B3M001, n_farm = B3Z001, cows = NA, horses = NA, value_implements = NA, pop = B13001, male = B3E001, 
-         female = B3E002, enrolled = B2A001+B2A002+B2A003+B2A004+B2A005, urban_pop = B2J001, 
-         pct_urban = B2J001/B13001*100, illiterate = NA, acres_farms = B3K001, 
-         acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(avg_size = acres_farms/n_farm, cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-#1960
-# The data are SPARSE yikes
-
-data1960a <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds89_1960_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME =  as.character(AREANAME))
-
-data1960b <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds90_1960_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1960c <- as_tibble(read.csv("raw-data/nhgis0003_csv/nhgis0003_ds91_1960_county.csv")) %>%
-  mutate(GISJOIN = as.character(GISJOIN), STATE = as.character(STATE), COUNTY = as.character(COUNTY), 
-         AREANAME = as.character(AREANAME))
-
-data1960j <- full_join(data1960a, data1960b, by = c("STATEA", "COUNTYA"))
-data1960j <- full_join(data1960j, data1960c, by = c("STATEA", "COUNTYA"))
-data1960j <- data1960j %>%
-  mutate(value_farm = NA, n_farm = NA, cows = NA, horses = NA, 
-         avg_size = NA, value_implements = NA, pop = B47001, male = B5H001, 
-         female = B5H002, enrolled = NA, urban_pop = NA, pct_urban = NA, illiterate = NA, 
-         acres_farms = NA, acres_irrigated = NA, pct_irrigated = NA) %>%
-  mutate(cows_per_person = cows/n_farm, pct_enrolled = enrolled/pop*100, pct_illiterate = illiterate/pop*100) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, value_farm, value_implements, avg_size, n_farm, 
-         acres_farms, acres_irrigated, pct_irrigated, cows, cows_per_person, horses, pop, male, female, 
-         enrolled, pct_enrolled, urban_pop, pct_urban, illiterate, pct_illiterate)
-
-```
-
-```{r}
-allData_inflated <- as_tibble(bind_rows(data1870j, data1880j, data1890j, 
-                       data1900j, data1910j, data1920j, data1930j, data1940j, data1950j, data1960j)) %>%
-  mutate(pct_male = male/pop*100, avg_value_farm = value_farm/n_farm, cows_per_farm = cows_per_person, 
-         avg_value_implements = value_implements/n_farm, pct_urban = ifelse(pct_urban > 100, NA, pct_urban)) %>%
-
-  # I mislabeled cows_per_farm as cows_per_person, so I'm doing damage control now
-
-  select(!cows_per_person) %>%
-  filter(STATE == "North Dakota" | STATE == "South Dakota" | STATE == "Dakota Territory" | 
-           STATE == "Nebraska" | STATE == "Minnesota" | STATE == "Iowa" | STATE == "Missouri" | 
-           STATE == "Wisconsin" | STATE == "Illinois" | STATE == "Kansas" | STATE == "Michigan" | 
-           STATE == "Indiana" | STATE == "Ohio",
-
-         # Get rid of the weird infinite values in this column
-
-         ! is.infinite(cows_per_farm))
-
-# Now adjust everything for inflation
-# I made a spreadsheet based on this website: 
-# https://www.officialdata.org/us/inflation/1960?amount=1
-
-inflation <- read_excel("inflation.xlsx")
-
-x <- left_join(allData_inflated, inflation, by = c("YEAR" = "year")) %>%
-  # Use the multiplier column to adjust for inflation
-  mutate(value_farm = value_farm*multiplier, value_implements = value_implements*multiplier) %>%
-  mutate(avg_value_farm = value_farm/n_farm, avg_value_implements = value_implements/n_farm) %>%
-  mutate("Value of Farms" = as.numeric(value_farm), "Average Farm Value" = as.numeric(avg_value_farm), 
-         "Value of Implements" = as.numeric(value_implements), 
-         "Average Implement Value" = as.numeric(avg_value_implements), 
-         "Average Farm Size" = as.numeric(avg_size), "Number of Farms" = as.numeric(n_farm), 
-         "Acres of Farmland" = as.numeric(acres_farms), "Irrigated Acres" = as.numeric(acres_irrigated), 
-         "Percentage of Farmland Irrigated" = as.numeric(pct_irrigated), "Number of Cows" = as.numeric(cows),
-         "Number of Cows per Farm" = as.numeric(cows_per_farm), "Number of Horses" = as.numeric(horses), 
-         "Total Population" = as.numeric(pop), "Male Population" = as.numeric(male), 
-         "Female Population" = as.numeric(female), "Male Percentage of Population" = as.numeric(pct_male),
-         "Total School Enrollment" = as.numeric(enrolled), 
-         "Percentage School Enrollment" = as.numeric(pct_enrolled), 
-         "Urban Population" = as.numeric(urban_pop), 
-         "Percent of Population in Urban Areas" = as.numeric(pct_urban), 
-         "Illiterate Population" = as.numeric(illiterate), 
-         "Illiteracy Rate" = as.numeric(pct_illiterate)) %>%
-  select(STATEA, COUNTYA, GISJOIN, YEAR, STATE, COUNTY, 
-         "Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
-                   "Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
-                   "Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
-                   "Number of Horses", "Total Population", "Male Population", "Female Population", 
-                   "Male Percentage of Population", "Total School Enrollment", 
-                   "Percentage School Enrollment", "Urban Population", 
-                   "Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")
-
-# saving this for later
-# will need in shiny app
-#sliderOptions <- c("Value of Farms", "Average Farm Value", "Value of Implements", "Average Implement Value",
-                   #"Average Farm Size", "Number of Farms", "Acres of Farmland", "Irrigated Acres", 
-                   #"Percentage of Farmland Irrigated", "Number of Cows", "Number of Cows per Farm", 
-                   #"Number of Horses", "Total Population", "Male Population", "Female Population", 
-                   #"Male Percentage of Population", "Total School Enrollment", 
-                   #"Percentage School Enrollment", "Urban Population", 
-                   #"Percent of Population in Urban Areas", "Illiterate Population", "Illiteracy Rate")
-
-```
-
-```{r}
-Census_1870 <- x %>% filter(YEAR == "1870") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1880 <- x %>% filter(YEAR == "1880") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1890 <- x %>% filter(YEAR == "1890") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1900 <- x %>% filter(YEAR == "1900") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1910 <- x %>% filter(YEAR == "1910") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1920 <- x %>% filter(YEAR == "1920") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1930 <- x %>% filter(YEAR == "1930") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1940 <- x %>% filter(YEAR == "1940") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1950 <- x %>% filter(YEAR == "1950") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1960 <- x %>% filter(YEAR == "1960") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-saveRDS(Census_1870, file = "final/1870.rds")
-saveRDS(Census_1880, file = "final/1880.rds")
-saveRDS(Census_1890, file = "final/1890.rds")
-saveRDS(Census_1900, file = "final/1900.rds")
-saveRDS(Census_1910, file = "final/1910.rds")
-saveRDS(Census_1920, file = "final/1920.rds")
-saveRDS(Census_1930, file = "final/1930.rds")
-saveRDS(Census_1940, file = "final/1940.rds")
-saveRDS(Census_1950, file = "final/1950.rds")
-saveRDS(Census_1960, file = "final/1960.rds")
-
-```
-
-# Working App
-
-```{r workingApp}
-# START THE APP
-
-# Input
-
-ui <- fluidPage(
-
-    # flatly is a nice theme to make the tabs look put together
-
-    # Application title
-
-    navbarPage("Agriculture and Education in the U.S.", theme = shinytheme("flatly"),
-
-               # remember that tab panel is for what shows up in the tab, and title panel is 
-               # for what shows up at the top of the page within the tab
-
-               tabPanel("Intro to Midwestern Agriculture",
-
-                        # Give the page a title
-                        titlePanel("Relationship Between Farm Value and Education, by County"),
-
-                        # Generate a row with a sidebar
-                        sidebarLayout(
-
-                            # Define the sidebar with one input
-                            # comprehensive options:
-                            # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
-                            sidebarPanel(
-                                selectInput("Year", "Year:", 
-                                            choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
-                                            multiple = TRUE)),
-                            helpText("From Great Plains Population and Environment Survey Agricultural and Social & Demographic Data.")
-                        ),
-
-                        # Create a spot for the plot
-                        mainPanel(
-                            tabsetPanel(id = "tabsMain",
-                                        tabPanel("Plot",
-
-                                                 # this tells the UI what plot to put
-
-                                                 plotOutput("agPlot")
-                                        )))),
-               tabPanel("Models",
-
-                        # Give the page a title
-                        titlePanel("Relationship Between Farm Value and Education, by County"),
-                        p("Make some models! (See notes)"),
-                        gt_output("model1")
-               ),
-
-               tabPanel("Maps",
-                        titlePanel("Midwest Map Explorer"),
-                        p("For best results, view in fullscreen mode."),
-                        p("Maps may take up to 30 seconds to load...but they're worth it!"),
-                                sidebarLayout(
-
-                                      sidebarPanel(
-                                            uiOutput("choose_dataset"),
-                                            uiOutput("choose_columns"),
-                                            uiOutput("choose_columns2"),
-                                                                      br(),
-                                                                  ),
-                                         mainPanel(leafletOutput("map1"),
-                                                   leafletOutput("map2"))
-                                     )),
-
-               tabPanel("About",
-
-                        # Give the page a title
-                        titlePanel("About this project"),
-                        p("My name is Julia Englebert.
-I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights.
-Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
-                        br(),
-                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States.
-By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century.
-I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960.
-My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org.
-IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960.
-All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
-                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
-                        br(),
-                        p("To see more of my work, please visit my Github account:"),
-                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
-                        br(),
-                        titlePanel("Citations"),
-                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles.
-IPUMS National Historical Geographic Information System: Version 14.0 [Database].
-Minneapolis, MN: IPUMS.
-2019.
-http://doi.org/10.18128/D050.V14.0"),
-                        p("The College of William and Mary and the Minnesota Population Center.
-School Attendance Boundary Information System (SABINS): Version 1.0.
-Minneapolis, MN: University of Minnesota 2011.")
-               )
-               )
-    )
-
-# use plotly to allow hover text to appear on plots, giving more info about the point
-
-server <- (function(input, output, session) {
-
-    # Fill in the spot we created for a plot
-    output$agPlot <- renderPlot({
-
-        # Render a plot
-        allData %>%
-            filter(YEAR %in% input$Year) %>%
-            mutate(YEAR = as.factor(YEAR)) %>%
-            group_by(YEAR) %>%
-            ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
-            geom_point(alpha = 0.1) +
-            geom_smooth(method = "lm", se = FALSE) +
-            scale_color_viridis_d() +
-            ylim(0, 45) +
-            scale_x_log10(labels = comma) +
-            labs(title = "Cows and Education",
-                 subtitle = "Relationship between the number of cows per farm\nand school enrollment, by county",
-                 x = "Average Cows per Farm",
-                 y = "Percent of Population Enrolled in School") +
-            theme_classic()
-    })
-
-    output$choose_dataset <- renderUI({
-        selectInput("dataset", "Choose Census Year:", as.list(data_sets))
-    })
-
-    # Check boxes
-    output$choose_columns <- renderUI({
-        # If missing input, return to avoid error later in function
-        if(is.null(input$dataset))
-            return()
-
-        # Get the data set with the appropriate name
-        dat <- get(input$dataset)
-
-        # Get rid of "YEAR" and "STATE" so that users don't pick them and make senseless maps
-
-        colnames1 <- names(dat)
-        colnames <- colnames1[!colnames1 %in% "YEAR" & !colnames1 %in% "STATE"]
-
-        # Create the dropdown menu and select all columns by default
-        selectizeInput("columns", "Choose Variable:", 
-                    choices  = colnames)
-    })
-
-    # Map output
-    output$map1 <- renderLeaflet({
-        # If missing input, return to avoid error later in function
-        if(is.null(input$dataset))
-            return()
-
-        # Get the data set
-        dat <- get(input$dataset)
-
-        # Make sure columns are correct for data set (when data set changes, the
-        # columns will initially be for the previous data set)
-        if (is.null(input$columns) || !(input$columns %in% names(dat)))
-            return()
-
-        year <- dat %>% slice(1) %>% pull(YEAR)
-
-        # Keep the selected columns
-        dat <- dat[, input$columns, drop = FALSE]
-
-        myvar <- input$columns
-
-        datayear <- mwData %>%
-            # use get() to use the stored variable
-            # idk why, but it doesn't work without this
-            # the variable must be converted to a double in the original mwData dataset
-            mutate(myvariable = (get(myvar))) %>%
-            filter(YEAR == year, myvariable > 0) %>%
-            select(GISJOIN, myvariable)
-
-        # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
-        # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
-        # layer is the filename of the .shp file inside the
-        # folder dsn points to.
-
-        dsnyear <- shapefiles %>% 
-            filter(Year == year) %>% 
-            pull(Dsn)
-        layeryear <- shapefiles %>% 
-            filter(Year == year) %>% 
-            pull(Layer)
-
-        countyyear <- sf::st_read(dsn = dsnyear,
-                                  layer = layeryear)
-
-        countyyear <-
-            countyyear %>%
-            merge(datayear, "GISJOIN")
-
-        # Set projection of tracts dataset to `projection` required by leaflet
-
-        countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
-
-        # Condense size of data for faster processing
-
-        countyyear <- rmapshaper::ms_simplify(countyyear)
-
-        # Set palette color
-
-        pal <- colorNumeric("viridis", NULL)
-
-        #  Plot the data
-
-        # use shiny to add a title, which will be be equivalent to myvar
-
-        leaflet(countyyear) %>%
-            addTiles() %>%
-            addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
-                        fillColor = ~pal(myvariable)) %>%
-            addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
-
-    })
-
-    # map2 dropdown menu
-    output$choose_columns2 <- renderUI({
-        # If missing input, return to avoid error later in function
-        if(is.null(input$dataset))
-            return()
-
-        # Get the data set with the appropriate name
-        dat <- get(input$dataset)
-
-        # Get rid of "YEAR" and "STATE" so that users don't pick them and make senseless maps
-
-        colnames1 <- names(dat)
-        colnames <- colnames1[!colnames1 %in% "YEAR" & !colnames1 %in% "STATE"]
-
-        # Create the dropdown menu and select all columns by default
-        selectizeInput("columns2", "Choose Variable to Compare:", 
-                    choices  = colnames,
-                    options = list(
-                        placeholder = 'Variables...',
-                        onInitialize = I('function() { this.setValue(""); }')))
-    })
-
-    output$map2 <- renderLeaflet({
-        # If missing input, return to avoid error later in function
-        if(is.null(input$dataset))
-            return()
-
-        # Get the data set
-        dat <- get(input$dataset)
-
-        # Make sure columns are correct for data set (when data set changes, the
-        # columns will initially be for the previous data set)
-        if (is.null(input$columns2) || !(input$columns2 %in% names(dat)))
-            return()
-
-        year <- dat %>% slice(1) %>% pull(YEAR)
-
-        # Keep the selected columns
-        dat <- dat[, input$columns2, drop = FALSE]
-
-        myvar <- input$columns2
-
-        datayear <- mwData %>%
-            # use get() to use the stored variable
-            # idk why, but it doesn't work without this
-            # the variable must be converted to a double in the original mwData dataset
-            mutate(myvariable = (get(myvar))) %>%
-            filter(YEAR == year, myvariable > 0) %>%
-            select(GISJOIN, myvariable)
-
-        # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
-        # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
-        # layer is the filename of the .shp file inside the
-        # folder dsn points to.
-
-        dsnyear <- shapefiles %>% 
-            filter(Year == year) %>% 
-            pull(Dsn)
-        layeryear <- shapefiles %>% 
-            filter(Year == year) %>% 
-            pull(Layer)
-
-        countyyear <- sf::st_read(dsn = dsnyear,
-                                  layer = layeryear)
-
-        countyyear <-
-            countyyear %>%
-            merge(datayear, "GISJOIN")
-
-        # Set projection of tracts dataset to `projection` required by leaflet
-
-        countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
-
-        # Condense size of data for faster processing
-
-        countyyear <- rmapshaper::ms_simplify(countyyear)
-
-        # Set palette color
-
-        pal <- colorNumeric("viridis", NULL)
-
-        #  Plot the data
-
-        # use shiny to add a title, which will be be equivalent to myvar
-
-        leaflet(countyyear) %>%
-            addTiles() %>%
-            addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
-                        fillColor = ~pal(myvariable)) %>%
-            addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
-
-    })
-
-    output$model1 <- render_gt({
-
-        # experimental model
-
-        model1
-
-    })
-})
-
-# Run the application
-
-shinyApp(ui = ui, server = server)
-
-```

### 2020-04-28 final/app.R

@@ -15,7 +15,8 @@ library(tidyverse)
 library(gt)
 library(devtools)
 library(tibble)
-library(foreign)
+# don't think I need this
+#library(foreign)
 library(scales)
 library(shinythemes)
 library(plotly)
@@ -54,7 +55,8 @@ codebook <- read_excel("mw_codebook.xlsx") %>%
 
 # Read in the shapefiles
 
-shapefiles <- readRDS(file = "shapefiles.rds")
+shapefiles <- readRDS(file = "shapefiles.rds") %>% 
+    head(n = 3L)
 
 
 # Read in the dataset made specifically for mapping
@@ -64,7 +66,7 @@ x <- readRDS("mwData.rds")
 # Now subset by year
 
 # I tried saving these as Rds files as well, but I kept getting a weird error
-# Might be a bug, according to SO
+
 Census_1870 <- x %>% filter(YEAR == "1870") %>%
     select_if(~ !all(is.na(.))) %>%
     select(-STATEA:-GISJOIN & -COUNTY)
@@ -77,39 +79,44 @@ Census_1890 <- x %>% filter(YEAR == "1890") %>%
     select_if(~ !all(is.na(.))) %>%
     select(-STATEA:-GISJOIN & -COUNTY)
 
-Census_1900 <- x %>% filter(YEAR == "1900") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
+# Code for the rest of the files
+# If only I could upload them all 
 
-Census_1910 <- x %>% filter(YEAR == "1910") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
+#Census_1900 <- x %>% filter(YEAR == "1900") %>%
+    #select_if(~ !all(is.na(.))) %>%
+    #select(-STATEA:-GISJOIN & -COUNTY)
 
-Census_1920 <- x %>% filter(YEAR == "1920") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
+#Census_1910 <- x %>% filter(YEAR == "1910") %>%
+    #select_if(~ !all(is.na(.))) %>%
+    #select(-STATEA:-GISJOIN & -COUNTY)
 
-Census_1930 <- x %>% filter(YEAR == "1930") %>%
[…]

### 2020-05-01 ms_8.Rmd
Giant sf object under construction

@@
 
+# I had to shorten the file paths in order to deploy my app
+# The maps can no longer render in this directory as a result
 dsn1870 <- "raw-data/nhgis0002_shape/nhgis0002_shapefile_tl2008_us_county_1870"
@@
 layer1890 <- "US_county_1890_conflated"
+
 dsn1900 <- "raw-data/nhgis0003_shape/nhgis0003_shapefile_tl2008_us_county_1900"
@@
 # only change these two lines!
-year <- 1950
+year <- 1870
 
@@
 anim_save("final/pop_gif.gif", pop_gif)
+
+# New Dataset for Mapping
+
+```{r mwData-shapefile-merger}
+# NEW tibble
+# columns: COUNTYA?, YEAR, variable, value
+
+mapData1 <- mwData %>% pivot_longer(cols = "Value of Farms":"Illiteracy Rate", names_to = "Variable")
+
+mapData <- left_join(mapData1, shapefiles, by = c("YEAR" = "Year"))
+
+```
+
+```{r}
+year <- 1870
+
+#choices for myvar will be determined by: 
+appPlan <- mwData %>%
+  filter(YEAR == year) %>%
+  select_if(~ !all(is.na(.))) %>%
+  select(-STATEA:-COUNTY)
+
+# myvar options will be determined by column names of appPlan & displayed via dropdown widget
+myvar <- "Value of Farms"
+
+datayear <- mapData %>%
+  # use get() to use the stored variable
+  # idk why, but it doesn't work without this
+  # the variable must be converted to a double in the original mwData dataset
+  filter(YEAR == year, Variable == myvar) %>%
+  select(GISJOIN, YEAR, Variable, value, Dsn, Layer)
+
+# Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
+# dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
+# layer is the filename of the .shp file inside the
+# folder dsn points to.
+
+countyyear <- sf::st_read(dsn = dsnyear,
+                         layer = layeryear)
+# IN PROGRESS, START HERE
+# make a huge sf object that has all of the info
+# save as RDS
+# use only one filter (for year and variable) with the code calling the palette and below
+
+countyyear <-
+  countyyear %>%
+  merge(datayear, "GISJOIN")
+
+# wow you can save sf objects as RDS files, apparently!
+# do this once you have everything done
+# but then again, if I can save this as an RDS, do I actually need to pivot longer? does that make sense?
+#saveRDS(countyyear, file = "testSfObj.rds")
+
+# Set projection of tracts dataset to `projection` required by leaflet
+
+countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
+
+# Condense size of data for faster processing
+
+countyyear <- rmapshaper::ms_simplify(countyyear)
+
+# Set palette color
+
+pal <- colorNumeric("viridis", NULL)
+
+#  Plot the data
+
+# use shiny to add a title, which will be be equivalent to myvar
+
+leaflet(countyyear) %>%
+  addTiles() %>%
+  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
+              fillColor = ~pal(value)) %>%
+  addLegend(pal = pal, values = ~value, opacity = 1.0, title = myvar)
+```

### 2020-05-01 final/app.R

@@ -14,17 +14,21 @@ library(haven)
 library(tidyverse)
 library(gt)
 library(devtools)
-library(tibble)
-# don't think I need this
+#library(tibble)
+
+# I'm getting a warning about the foreign package when I try to deploy
+# Luckily, it's not necessary
 #library(foreign)
+
 library(scales)
 library(shinythemes)
-library(plotly)
+#library(plotly)
 library(leaflet)
 library(broom)
-library(patchwork)
-library(magick)
-library(giphyr)
+# probably don't need these, but double check
+#library(patchwork)
+#library(magick)
+#library(giphyr)
 library(gridExtra)
 library(readxl)
 
@@ -32,7 +36,7 @@ library(readxl)
 # packages for mapping
 library(leaflet)
 library(spdplyr)
-library(rmapshaper)
+#library(rmapshaper)
 library(sf)
 
 
@@ -54,6 +58,8 @@ codebook <- read_excel("mw_codebook.xlsx") %>%
     column_to_rownames(var="Year")
 
 # Read in the shapefiles
+# These filepaths were initially too long (101 bytes, I kid you not)
+# So I changed the file names
 
 shapefiles <- readRDS(file = "shapefiles.rds") %>% 
     head(n = 3L)
@@ -245,18 +251,22 @@ ui <- fluidPage(
                
                tabPanel("Maps",
                         titlePanel("Midwest Map Explorer"),
-                        p("Choose a census year and up to two variables to generate and compare maps of the midwest."),
+                        p("Choose a census year and variable to generate and compare maps of the midwest."),
                         helpText("Please note that maps may take up to 30 seconds to render."),
                         sidebarLayout(
+                            
+                            # App keeps disconnecting from the server
+                            # Try doing only one map
                             
                             sidebarPanel(
                                 uiOutput("choose_dataset"),
                                 uiOutput("choose_columns"),
-                                uiOutput("choose_columns2"),
+                                #uiOutput("choose_columns2"),
                                 br(),
                             ),
-                            mainPanel(leafletOutput("map1"),
-                                      leafletOutput("map2"))
+                            mainPanel(leafletOutput("map1")#,
+                                      #leafletOutput("map2")
+                                      )
                         )),
                
                tabPanel("About",
@@ -505,10 +515,11 @@ server <- (function(input, output, session) {
[…]

### 2020-05-05 ms_8.Rmd
Commit plan z, png maps

@@
 library(webshot)
-webshot::install_phantomjs()
+library(mapview)
+# Probably not necessary
+#webshot::install_phantomjs()
 
@@
 # myvar options will be determined by column names of appPlan & displayed via dropdown widget
-myvar <- "Value of Farms"
+myvar <- "Average Implement Value"
 
@@
 
-countyyear <- rmapshaper::ms_simplify(countyyear)
+#countyyear <- rmapshaper::ms_simplify(countyyear)
 
@@
               fillColor = ~pal(myvariable)) %>%
+  #setView(-87.4290, 43, zoom = 6) %>%
   addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
 
+```
+
+```{r}
+# Saving maps as pngs
+# Since I couldn't get my interactive maps to run on shiny, this is the supremely disappointing 
+# work-around
+
+## 'leaflet' objects (image above)
+
+# mapshot(enroll_1870, file = "enroll_1870.png")
+
+# This initially worked, but then R started crashing every time I tried to run it
+# Now I'm resorting to sCreEnsHotS
+
+```
+
+# SCREENSHOT Section
+# Delete later
+
+```{r}
+year <- 1880
+
+#choices for myvar will be determined by: 
+appPlan <- mwData %>%
+  filter(YEAR == year) %>%
+  select_if(~ !all(is.na(.))) %>%
+  select(-STATEA:-COUNTY)
+
+# myvar options will be determined by column names of appPlan & displayed via dropdown widget
+myvar <- "Average Implement Value"
+
+datayear <- mwData %>%
+  # use get() to use the stored variable
+  # idk why, but it doesn't work without this
+  # the variable must be converted to a double in the original mwData dataset
+  mutate(myvariable = (get(myvar))) %>%
+  filter(YEAR == year, myvariable > 0) %>%
+  select(GISJOIN, myvariable)
+
+# Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
+# dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
+# layer is the filename of the .shp file inside the
+# folder dsn points to.
+
+dsnyear <- shapefiles %>% 
+  filter(Year == year) %>% 
+  pull(Dsn)
+layeryear <- shapefiles %>% 
+  filter(Year == year) %>% 
+  pull(Layer)
+
+countyyear <- sf::st_read(dsn = dsnyear,
+                         layer = layeryear)
+
+countyyear <-
+  countyyear %>%
+  merge(datayear, "GISJOIN")
+
+# Set projection of tracts dataset to `projection` required by leaflet
+
+countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
+
+# Condense size of data for faster processing
+
+#countyyear <- rmapshaper::ms_simplify(countyyear)
+
+# Set palette color
+
+pal <- colorNumeric("viridis", NULL)
+
+#  Plot the data
+
+# use shiny to add a title, which will be be equivalent to myvar
+
+leaflet(countyyear) %>%
+  addTiles() %>%
+  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
+              fillColor = ~pal(myvariable)) %>%
+  #setView(-87.4290, 43, zoom = 6) %>%
+  addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
+```
+
+```{r}
+year <- 1890
+
+# myvar options will be determined by column names of appPlan & displayed via dropdown widget
+myvar <- "Average Implement Value"
+
+datayear <- mwData %>%
+  # use get() to use the stored variable
+  # idk why, but it doesn't work without this
+  # the variable must be converted to a double in the original mwData dataset
+  mutate(myvariable = (get(myvar))) %>%
+  filter(YEAR == year, myvariable > 0) %>%
+  select(GISJOIN, myvariable)
+
+# Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
+# dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
+# layer is the filename of the .shp file inside the
+# folder dsn points to.
+
+dsnyear <- shapefiles %>% 
+  filter(Year == year) %>% 
+  pull(Dsn)
+layeryear <- shapefiles %>% 
+  filter(Year == year) %>% 
+  pull(Layer)
+
+countyyear <- sf::st_read(dsn = dsnyear,
+                         layer = layeryear)
+
+countyyear <-
+  countyyear %>%
+  merge(datayear, "GISJOIN")
+
+# Set projection of tracts dataset to `projection` required by leaflet
+
+countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
+
+# Condense size of data for faster processing
+
+#countyyear <- rmapshaper::ms_simplify(countyyear)
+
+# Set palette color
+
+pal <- colorNumeric("viridis", NULL)
+
+#  Plot the data
+
+# use shiny to add a title, which will be be equivalent to myvar
+
+leaflet(countyyear) %>%
+  addTiles() %>%
+  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
+              fillColor = ~pal(myvariable)) %>%
+  #setView(-87.4290, 43, zoom = 6) %>%
+  addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)
 ```
@@
 ```{r mwData-shapefile-merger}
+
+# Ultimately, this didn't work
+# I'm giving up and saving individual leaflets as
+
 # NEW tibble
@@
 
+# Plan: make the huge sf object for each variable from 1870-1890
+# bind all the rows
+# it will be huge
+# save as rds
+# read in at the beginning, just filter by year and variable, which now each have their own column
+# figure out how to get widgets to work
+
 mapData1 <- mwData %>% pivot_longer(cols = "Value of Farms":"Illiteracy Rate", names_to = "Variable")
 
-mapData <- left_join(mapData1, shapefiles, by = c("YEAR" = "Year"))
+mapData <- left_join(mapData1, shapefiles, by = c("YEAR" = "Year")) %>%
+
+  # Only do the first three decades
+  # Shiny doesn't give you enough space to upload the rest anyway
+
+  filter(YEAR == "1870" | YEAR == "1880" | YEAR == "1890")
 
@@
 
+dsnyear <- datayear %>% 
+  slice(1) %>% 
+  pull(Dsn)
+
+layeryear <- datayear %>% 
+  slice(1) %>% 
+  pull(Layer)
+
 countyyear <- sf::st_read(dsn = dsnyear,
@@
 
-countyyear <- rmapshaper::ms_simplify(countyyear)
+countyyear1 <- rmapshaper::ms_simplify(countyyear)
 
@@
 ```
+
+```{r}
+year <- 1870
+
+# myvar options will be determined by column names of appPlan & displayed via dropdown widget
+myvar <- "Average Farm Value"
+
+datayear <- mapData %>%
+  # use get() to use the stored variable
+  # idk why, but it doesn't work without this
+  # the variable must be converted to a double in the original mwData dataset
+  filter(YEAR == year) %>%
+  select(GISJOIN, YEAR, Variable, value, Dsn, Layer)
+
+# Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
+# dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
+# layer is the filename of the .shp file inside the
+# folder dsn points to.
+
+dsnyear <- datayear %>% 
+  slice(1) %>% 
+  pull(Dsn)
+
+layeryear <- datayear %>% 
+  slice(1) %>% 
+  pull(Layer)
+
+countyyear <- sf::st_read(dsn = dsnyear,
+                         layer = layeryear)
+
+# make a huge sf object that has all of the info
+# save as RDS
+# use only one filter (for year and variable) with the code calling the palette and below
+
+countyyear <-
+  countyyear %>%
+  merge(datayear, "GISJOIN")
+
+# wow you can save sf objects as RDS files, apparently!
+# do this once you have everything done
+# but then again, if I can save this as an RDS, do I actually need to pivot longer? does that make sense?
+#saveRDS(countyyear, file = "testSfObj.rds")
+
+# Set projection of tracts dataset to `projection` required by leaflet
+
+countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")
+
+# Condense size of data for faster processing
+# This is taking too long to run; session encounters a fatal error
+# What if I simplify after filtering?
+# countyyear <- rmapshaper::ms_simplify(countyyear)
+
+countyear <- countyyear %>%
+  filter(Variable == myvar)
+
+countyyear <- rmapshaper::ms_simplify(countyyear)
+
+pal <- colorNumeric("viridis", NULL)
+
+#  Plot the data
+
+# use shiny to add a title, which will be be equivalent to myvar
+
+leaflet(countyyear) %>%
+  addTiles() %>%
+  addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
+              fillColor = ~pal(value)) %>%
+  #setView(-87.4290, 43, zoom = 6) %>%
+  addLegend(pal = pal, values = ~value, opacity = 1.0, title = myvar)
+
+```

### 2020-05-05 shiny-aspirations.Rmd
Commit ms_8 shiny draft copy

---
title: "shiny-aspirations"
author: "Julia Englebert"
date: "5/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is what I would have done, if I could have gotten the maps to run on shiny

```{r}
#
# This is a Shiny web application. You can run the application by clicking
# the 'Run App' button above.
#
# Find out more about building applications with Shiny here:
#
#    http://shiny.rstudio.com/
#

# LOAD THE REQUIRED PACKAGES

library(shiny)
library(haven)
library(tidyverse)
library(gt)
library(devtools)
library(tibble)

# I'm getting a warning about the foreign package when I try to deploy
# Luckily, it's not necessary
#library(foreign)

library(scales)
library(shinythemes)
#library(plotly)
library(leaflet)
library(broom)
# probably don't need these, but double check
#library(patchwork)
#library(magick)
#library(giphyr)
library(gridExtra)
library(readxl)

# packages for mapping
library(leaflet)
library(spdplyr)
#library(rmapshaper)
library(sf)

# READ IN THE DATA

# Read in the model for gt and plot

maleedu_model <- readRDS("maleedu_model.rds")

cowedu_model <- readRDS("cowedu_model.rds")

cows_size_model <- readRDS("cows_size_model.rds")

# Read in the codebook
# This is an excel sheet that I made in order to condense the original codebooks provided by NHGIS

codebook <- read_excel("mw_codebook.xlsx") %>% 
    remove_rownames %>% 
    column_to_rownames(var="Year")

# Read in the shapefiles
# These filepaths were initially too long (101 bytes, I kid you not)
# So I changed the file names

shapefiles <- readRDS(file = "shapefiles.rds") %>% 
    head(n = 3L)

# Read in the dataset made specifically for mapping

x <- readRDS("mwData.rds")

# Now subset by year

# I tried saving these as Rds files as well, but I kept getting a weird error

Census_1870 <- x %>% filter(YEAR == "1870") %>%
    select_if(~ !all(is.na(.))) %>%
    select(-STATEA:-GISJOIN & -COUNTY)

Census_1880 <- x %>% filter(YEAR == "1880") %>%
    select_if(~ !all(is.na(.))) %>%
    select(-STATEA:-GISJOIN & -COUNTY)

Census_1890 <- x %>% filter(YEAR == "1890") %>%
    select_if(~ !all(is.na(.))) %>%
    select(-STATEA:-GISJOIN & -COUNTY)

# Code for the rest of the files
# If only I could upload them all

#Census_1900 <- x %>% filter(YEAR == "1900") %>%
    #select_if(~ !all(is.na(.))) %>%
    #select(-STATEA:-GISJOIN & -COUNTY)

#Census_1910 <- x %>% filter(YEAR == "1910") %>%
    #select_if(~ !all(is.na(.))) %>%
    #select(-STATEA:-GISJOIN & -COUNTY)

#Census_1920 <- x %>% filter(YEAR == "1920") %>%
    #select_if(~ !all(is.na(.))) %>%
    #select(-STATEA:-GISJOIN & -COUNTY)

#Census_1930 <- x %>% filter(YEAR == "1930") %>%
    #select_if(~ !all(is.na(.))) %>%
    #select(-STATEA:-GISJOIN & -COUNTY)

#Census_1940 <- x %>% filter(YEAR == "1940") %>%
    #select_if(~ !all(is.na(.))) %>%
    #select(-STATEA:-GISJOIN & -COUNTY)

#Census_1950 <- x %>% filter(YEAR == "1950") %>%
    #select_if(~ !all(is.na(.))) %>%
    #select(-STATEA:-GISJOIN & -COUNTY)

#Census_1960 <- x %>% filter(YEAR == "1960") %>%
    #select_if(~ !all(is.na(.))) %>%
    #select(-STATEA:-GISJOIN & -COUNTY)

# Combine the datasets into a vector, renaming them by only the year 
# This is the code for include all GIS files, but I had to delete most of them because they were HUGE
# data_sets <- c("1870" = "Census_1870", "1880" = "Census_1880", "1890" = "Census_1890", "1900" = "Census_1900", "1910" = "Census_1910", "1920" = "Census_1920",  "1930" ="Census_1930",  "1940" ="Census_1940",  "1950" ="Census_1950",  "1960" ="Census_1960")
# This is the code for the remaining files
data_sets <- c("1870" = "Census_1870", "1880" = "Census_1880", "1890" = "Census_1890")

# Read in the dataset for maps and models

allData <- readRDS("my_data.rds")

# START THE APP

# Input

ui <- fluidPage(

    # flatly is a nice theme to make the tabs look put together

    # Application title

    navbarPage("Agriculture and Education in the Midwest", theme = shinytheme("flatly"),

               # remember that tab panel is for what shows up in the tab, and title panel is 
               # for what shows up at the top of the page within the tab

               tabPanel("Welcome",
                        titlePanel("Midwest Agriculture in Brief"),

                        # Create a spot for the plot
                        # Why is the text on top of the image?? :(
                        mainPanel(
                            helpText("To understand the relationship between agriculture and education, let's look at some trends in agriculture over time. Click the tabs below to visualize major changes."),
                            tabsetPanel(id = "tabsMain",
                                        tabPanel("Farms & Acreage",

                                                 # Create a fluid row in order to comment next to the gifs
                                                 # column specifications might need to be changed

                                                 fluidRow(column(width = 5, textOutput("acre_com")),
                                                          column(width = 7, align = "center",
                                                                 imageOutput("acre_gif")))

                                        ),
                                        tabPanel("Urbanization",
                                                 # Generate a row with a sidebar
                                                 fluidRow(column(width = 5, textOutput("pop_com")),
                                                              column(width = 7, align = "center",
                                                                     imageOutput("pop_gif")))
                                                 ),
                                        tabPanel("Mechanization",
                                                fluidRow(column(width = 5, textOutput("plot2_com")),
                                                              column(width = 7, align = "center",
                                                                     imageOutput("plot2")))
                            )))),

               tabPanel("Models",

                        # Give the page a title
                        titlePanel("Agriculture and Education, by County"),

                        # Create a spot for the plot
                        mainPanel("",
                                  tabsetPanel(id = "tabsMain",

                                              tabPanel("Implements",
                                                       # Generate a row with a sidebar
                                                       sidebarLayout(

                                                           # Define the sidebar with one input
                                                           # comprehensive options:
                                                           # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
                                                           sidebarPanel(
                                                               selectInput("impYear", "Year:", 
                                                                           choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940"), 
                                                                           multiple = TRUE)),
                                                           helpText("Select a year from the dropdown menu in order to explore the correlation between mechanization and school enrollment over time.")
                                                       ),

                                                       # this tells the UI what plot to put

                                                       plotOutput("impedu_plot"),
                                                       p("The plot shows a positive correlation between school enrollment and implement value in 1870, but that correlation disappears in later years."),
                                                       gt_output("cowedu_gt"),
                                                       p("This linear regression model shows that an increase in average implement value has essentially no effect on school enrollment. Curiously, the number of cows per farm is strongly associated with higher school enrollment, as we will explore in the next tab.")

                                              ),

                                              tabPanel("Cattle",
                                                       # Generate a row with a sidebar

                                                       sidebarLayout(

                                                           # Define the sidebar with one input
                                                           # comprehensive options:
                                                           # c("1870" = "1870", "1880" = "1880", "1890" = "1890", "1900" = "1900","1910" = "1910", "1920" = "1920", "1930" = "1930", "1940" = "1940", "1950" = "1950", "1960" = "1960")
                                                           sidebarPanel(
                                                               selectInput("cowYear", "Year:", 
                                                                           choices=c("1870" = "1870", "1890" = "1890", "1910" = "1910"), 
                                                                           multiple = TRUE)),
                                                           helpText("Select a year from the dropdown menu in order to explore the correlation between cattle ownership and school enrollment over time.")
                                                       ),

                                                       # this tells the UI what plot to put

                                                       plotOutput("agPlot"),
                                                       p("The plot shows a positive correlation between school enrollment and the average number of cows per farm for the years where data was available. The number of cows might simply be a confounding variable that acts as a stand-in for farm size, but as the table below will show, this does not seem to be the case."),
                                                       gt_output("cows_size_gt"),
                                                       p("This linear regression model shows that a one-cow increase in the number of cows per farm in a given county results in a 0.23% increase in school enrollment. This is significant given that school enrollment hovers around just 25% in most counties for the years studied. Curiously, farm size is not associated with a similarly significant increase in enrollment, and the interaction between cows per farm and farm size actually leads to a decrease in enrollment.")
                                              ),

                                              tabPanel("Gender",

                                                       #mainPanel("",
                                                                 #gt_output("maleedu_gt"))

                                                       fluidRow(column(width = 6, plotOutput("maleedu_plot")),
                                                                column(width = 6, #align = "center",
                                                                       gt_output("maleedu_gt"))), 
                                                       fluidRow(column(width = 6, textOutput("maletext1")),
                                                                column(width = 6, #align = "center",
                                                                       textOutput("maletext2")))

                                              )
                                  )
                        )),

               tabPanel("Maps",
                        titlePanel("Midwest Map Explorer"),
                        p("Choose a census year and variable to generate and compare maps of the midwest."),
                        helpText("Please note that maps may take up to 30 seconds to render."),
                        sidebarLayout(

                            # App keeps disconnecting from the server
                            # Try doing only one map

                            sidebarPanel(
                                uiOutput("choose_dataset"),
                                uiOutput("choose_columns"),
                                #uiOutput("choose_columns2"),
                                br(),
                            ),
                            mainPanel(leafletOutput("map1")#,
                                      #leafletOutput("map2")
                                      )
                        )),

               tabPanel("About",

                        # Give the page a title
                        titlePanel("About this project"),
                        p("My name is Julia Englebert. I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights. Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
                        br(),
                        p("This project uses historical census and American Community Survey (ACS) data to explore correlations between agriculture and education in the Midwestern United States. By joining the two datasets, the former containing agricultural data and the later containing demographic data, I was able to look at the impact of farming, land value, crop value, and more on education in Midwestern communities over the course of a century. I chose this particular time range in part because of the data that was available to me, and in part due to the economic, industrial, and education policy changes that ocurred from 1870 to 1960. My data is sourced from the IPUMS National Historical Geographic Information System (NHGIS) at nhgis.org. IPUMS provides census data from around the world, but my project focuses specifically on the United States from 1870 to 1960. All datasets and gis files used for this project can be downloaded from the NHGIS Data Finder:"),
                        p(a(href="https://data2.nhgis.org/main", "NHGIS Data Finder")),
                        br(),
                        p("To see more of my work, please visit my Github account:"),
                        p(a(href="https://github.com/julia-englebert", "Julia Englebert")),
                        br(),
                        titlePanel("Citations"),
                        p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles. IPUMS National Historical Geographic Information System: Version 14.0 [Database]. Minneapolis, MN: IPUMS. 2019. http://doi.org/10.18128/D050.V14.0"),
                        p("The College of William and Mary and the Minnesota Population Center. School Attendance Boundary Information System (SABINS): Version 1.0. Minneapolis, MN: University of Minnesota 2011."),
                        p("Ian Webster. U.S. Inflation Calculator: 1635-2020, Department of Labor Data. Official Data Foundation, https://www.officialdata.org.")
                        ),

               tabPanel("Codebook",
                        titlePanel("How were variables measured for this project?"),
                        p("Choose a variable to find out! Take special note of the changes in school enrollment and farm value metrics over time."), 
                        varSelectInput("variables", "Variable:", codebook, multiple = FALSE),
                        helpText("These variable descriptions represent a condensed version of the codebooks provided by IPUMS NHGIS."),
                        tableOutput("codebook")
                        )
    )
)

# use plotly to allow hover text to appear on plots, giving more info about the point

server <- (function(input, output, session) {

    output$acre_com <- renderText({

        "The bar plot on the left shows that a general upward trend in the number of acres allotted to farming each year. What happened in 1920? It’s difficult to say, but perhaps the Economic Boom of the 1920s followed by the Great Crash of 1929 caused farmers to purchase and then lose vast amounts of land. The line plot on the right shows that the number of farms increased until the early 20th century before beginning to decline in most states. Taken together, these plots show that farms in the Midwest have been steadily increasing in size, while the number of people owning farms has been steadily decreasing."
    })

    output$acre_gif <- renderImage({
        list(src = "acre_gif.gif")

    }, deleteFile = FALSE)

    output$pop_com <- renderText({

        "Along with most regions of the world, the Midwest was experiencing rapid population growth during the years studied. As the plot on the left shows, population in all states except for North and South Dakota were trending upward. At the same time, the plot on the right illustrates that people were increasingly dwelling in urban areas. Although the urban population was almost nonexistent in 1870, it far surpasses the rural population by 1950."
    })

    output$pop_gif <- renderImage({
        list(src = "pop_gif.gif")

    }, deleteFile = FALSE)

    output$plot2_com <- renderText({

        "Farming methods were undergoing dramatic change during the period studied.This plot shows the value of farm implements and machinery each year, demonstrating the rapid mechanization of Midwest agriculture."
    })

    output$plot2 <- renderImage({
        list(src = "plot2.gif")

    }, deleteFile = FALSE)

    # Implements plot

    output$impedu_plot <- renderPlot({

        allData %>%
            filter(YEAR %in% input$impYear) %>%
            mutate(YEAR = as.factor(YEAR)) %>%
            mutate(YEAR = fct_explicit_na(YEAR, na_level = NA)) %>%
            group_by(YEAR) %>%
            select(YEAR, avg_value_implements, pct_enrolled) %>%
            na.omit() %>%
            ggplot(aes(x = avg_value_implements, y = pct_enrolled, color = YEAR)) + 
            geom_point(alpha = 0.1) +
            geom_smooth(method = "lm", se = FALSE) +
            scale_fill_viridis_d(aesthetics = "color") +
            ylim(0, 45) + 
            labs(title = "Relationship Between Farm Implements and Education, by County",
                 subtitle = "Midwestern United States", 
                 x = "Average Value of Implements", 
                 y = "Percent Population Enrolled in School") +
            scale_x_log10(labels = comma) +
            theme_classic()
    })

    output$cowedu_gt <- render_gt({
        cowedu_model %>%
            select(term, conf.low, estimate, conf.high) %>%
            gt() %>%
            tab_header(title = "Effect Cows and Implements on Enrollment",
                       subtitle = "With 95% Confidence Interval") %>%
            cols_label(term = "Coefficient Name", conf.low = "Lower Bound", estimate = "Estimated Value", conf.high = "Upper Bound") %>%
            fmt_number(columns = vars(conf.low, estimate, conf.high), decimals = 4)

    })

    # Fill in the spot we created for a plot
    output$agPlot <- renderPlot({

        # Render a plot
        allData %>%
            filter(YEAR %in% input$cowYear) %>%
            mutate(YEAR = as.factor(YEAR)) %>%
            mutate(YEAR = fct_explicit_na(YEAR, na_level = NA)) %>%
            group_by(YEAR) %>%
            select(YEAR, cows_per_farm, pct_enrolled) %>%
            na.omit() %>%
            ggplot(aes(x = cows_per_farm, y = pct_enrolled, color = YEAR)) +
            geom_point(alpha = 0.1) +
            geom_smooth(method = "lm", se = FALSE) +
            scale_color_viridis_d() +
            ylim(0, 45) +
            scale_x_log10(labels = comma) +
            labs(title = "Cows and Education",
                 subtitle = "Relationship between the number of cows per farm\nand school enrollment, by county",
                 x = "Average Cows per Farm",
                 y = "Percent of Population Enrolled in School") +
            theme_classic()
    })

    output$cows_size_gt <- render_gt({
        cows_size_model %>%
            select(term, conf.low, estimate, conf.high) %>%
            gt() %>%
            tab_header(title = "Effect Cows and Farm Size on Enrollment",
                       subtitle = "With 95% Confidence Interval") %>%
            cols_label(term = "Coefficient Name", conf.low = "Lower Bound", estimate = "Estimated Value", conf.high = "Upper Bound") %>%
            fmt_number(columns = vars(conf.low, estimate, conf.high), decimals = 4)

    })

    # maleedu plot and gt

    output$maleedu_plot <- renderPlot({
        maleedu_model %>%
            ungroup() %>%
            select(YEAR, pct_male_coef, pct_male_upper, pct_male_lower) %>%
            mutate(YEAR = as.factor(YEAR)) %>%
            group_by(YEAR) %>%
            ggplot(aes(x = YEAR, y = pct_male_coef)) + 
            geom_errorbar(ymin = maleedu_model$pct_male_lower, ymax = maleedu_model$pct_male_upper, color = "#453781FF") + 
            geom_point(color = "#440154FF") +
            theme_classic() +
            ylim(-1.5, 0.5) + 
            labs(x = "Year",
                 y = "Coefficient",
                 title = "Percent Male Coefficients Over Time",
                 subtitle = "Measuring the effect of being male\non school enrollment")
    })

    output$maleedu_gt <- render_gt({
        maleedu_model%>%
            ungroup() %>%

            # Select the columns needed, as per the instructions

            select(-data:-reg_results) %>%

            gt() %>%

            # Add the desired titles, subtitles, and column labels

            tab_header(title = "Effect of Gender and Number of Farms on Enrollment",
                       subtitle = "With 95% Confidence Interval") %>%
            tab_spanner(label = "Farm Coefficients", columns = vars(n_farm_lower, n_farm_coef, n_farm_upper)) %>%
            tab_spanner(label = "Percent Male Coefficients", columns = vars(pct_male_lower, pct_male_coef,pct_male_upper)) %>%
            tab_spanner(label = "Interaction Coefficients", columns = vars(interaction_lower, interaction_coef, interaction_upper)) %>%
            cols_label(YEAR = "Year", n_farm_coef	= "Estimate", n_farm_upper = "Upper", n_farm_lower = "Lower", pct_male_coef = "Estimate", pct_male_upper = "Upper", pct_male_lower = "Lower",	interaction_coef = "Estimate", interaction_upper = "Upper", interaction_lower = "Lower") %>%
            # Round all values to two decimal places

            fmt_number(columns = vars(n_farm_lower, n_farm_coef, n_farm_upper, pct_male_lower, pct_male_coef,pct_male_upper, interaction_lower, interaction_coef, interaction_upper), decimals = 4)

    })

    # Gender text
    output$maletext1 <- renderText({

        "This plot shows shows how the effect of being male on school enrollment reversed over time. From 1870 to 1920, having a higher percentage of males in the population associated with a decrease in school enrollment. From 1930 onward, however, that males and enrollment were positively correlated."
    })

    output$maletext2 <- renderText({

        "This table shows a linear regression model of the number of farms in a given county, the percent of the population that was male, and the interaction of those two as compared to the percent of the population enrolled in school. In contrast to gender, having more farms in a given county was not associated with a significant change in enrollment in any year."
    })

    # Map output

    output$choose_dataset <- renderUI({
        selectInput("dataset", "Choose Census Year:", as.list(data_sets))
    })

    # Check boxes
    output$choose_columns <- renderUI({
        # If missing input, return to avoid error later in function
        if(is.null(input$dataset))
            return()

        # Get the data set with the appropriate name
        dat <- get(input$dataset)

        # Get rid of "YEAR" and "STATE" so that users don't pick them and make senseless maps

        colnames1 <- names(dat)
        colnames <- colnames1[!colnames1 %in% "YEAR" & !colnames1 %in% "STATE"]

        # Create the dropdown menu and select all columns by default
        selectizeInput("columns", "Choose Variable:", 
                       choices  = colnames)
    })

    # Map output
    output$map1 <- renderLeaflet({
        # If missing input, return to avoid error later in function
        if(is.null(input$dataset))
            return()

        # Get the data set
        dat <- get(input$dataset)

        # Make sure columns are correct for data set (when data set changes, the
        # columns will initially be for the previous data set)
        if (is.null(input$columns) || !(input$columns %in% names(dat)))
            return()

        year <- dat %>% slice(1) %>% pull(YEAR)

        # Keep the selected columns
        dat <- dat[, input$columns, drop = FALSE]

        myvar <- input$columns

        datayear <- x %>%
            # use get() to use the stored variable
            # idk why, but it doesn't work without this
            # the variable must be converted to a double in the original mwData dataset
            mutate(myvariable = (get(myvar))) %>%
            filter(YEAR == year, myvariable > 0) %>%
            select(GISJOIN, myvariable)

        # Import shapefiles into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
        # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
        # layer is the filename of the .shp file inside the
        # folder dsn points to. 
        # all of these filepaths are stored in the "shapefiles" dataframe

        dsnyear <- shapefiles %>% 
            filter(Year == year) %>% 
            pull(Dsn)
        layeryear <- shapefiles %>% 
            filter(Year == year) %>% 
            pull(Layer)

        countyyear <- sf::st_read(dsn = dsnyear,
                                  layer = layeryear)

        countyyear <-
            countyyear %>%
            merge(datayear, "GISJOIN")

        # Set projection of tracts dataset to `projection` required by leaflet

        countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")

        # Condense size of data for faster processing

        #countyyear <- rmapshaper::ms_simplify(countyyear)

        # Set palette color

        pal <- colorNumeric("viridis", NULL)

        #  Plot the data

        # use shiny to add a title, which will be be equivalent to myvar

        leaflet(countyyear) %>%
            addTiles() %>%
            addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
                        fillColor = ~pal(myvariable)) %>%
            addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)

    })

    # map2 dropdown menu
    #output$choose_columns2 <- renderUI({
        # If missing input, return to avoid error later in function
        #if(is.null(input$dataset))
            #return()

        # Get the data set with the appropriate name
        #dat <- get(input$dataset)

        # Get rid of "YEAR" and "STATE" so that users don't pick them and make senseless maps

        #colnames1 <- names(dat)
        #colnames <- colnames1[!colnames1 %in% "YEAR" & !colnames1 %in% "STATE"]

        # Create the dropdown menu and select all columns by default
        #selectizeInput("columns2", "Choose Variable to Compare:", 
                       #choices  = colnames,
                       #options = list(
                           #placeholder = 'Variables...',
                           #onInitialize = I('function() { this.setValue(""); }')))
    #})

    #output$map2 <- renderLeaflet({
        # If missing input, return to avoid error later in function
        #if(is.null(input$dataset))
            #return()

        # Get the data set
        #dat <- get(input$dataset)

        # Make sure columns are correct for data set (when data set changes, the
        # columns will initially be for the previous data set)
        #if (is.null(input$columns2) || !(input$columns2 %in% names(dat)))
            #return()

        #year <- dat %>% slice(1) %>% pull(YEAR)

        # Keep the selected columns
        #dat <- dat[, input$columns2, drop = FALSE]

        #myvar <- input$columns2

        #datayear <- x %>%
            # use get() to use the stored variable
            # idk why, but it doesn't work without this
            # the variable must be converted to a double in the original mwData dataset
            #mutate(myvariable = (get(myvar))) %>%
            #filter(YEAR == year, myvariable > 0) %>%
            #select(GISJOIN, myvariable)

        # Import Census Tract Shapefile into R as SpatialPolygonsDataFrameFormat (SP Dataframe)
        # dsn is location of folder which contains shapefiles, (.proj, .shp etc.)
        # layer is the filename of the .shp file inside the
        # folder dsn points to.

        #dsnyear <- shapefiles %>% 
            #filter(Year == year) %>% 
            #pull(Dsn)
        #layeryear <- shapefiles %>% 
            #filter(Year == year) %>% 
            #pull(Layer)

        #countyyear <- sf::st_read(dsn = dsnyear,
                                  #layer = layeryear)

        #countyyear <-
            #countyyear %>%
            #merge(datayear, "GISJOIN")

        # Set projection of tracts dataset to `projection` required by leaflet

        #countyyear <- sf::st_transform(countyyear, crs="+init=epsg:4326")

        # Condense size of data for faster processing

        #countyyear <- rmapshaper::ms_simplify(countyyear)

        # Set palette color

        #pal <- colorNumeric("viridis", NULL)

        #  Plot the data

        # use shiny to add a title, which will be be equivalent to myvar

        #leaflet(countyyear) %>%
            #addTiles() %>%
            #addPolygons(stroke = FALSE, smoothFactor = 0.3, fillOpacity = 1,
                        #fillColor = ~pal(myvariable)) %>%
            #addLegend(pal = pal, values = ~myvariable, opacity = 1.0, title = myvar)

    #})

    output$model1 <- render_gt({

        # experimental model

        model1

    })

    # sometimes R flags this code chunk and says there's a bug, but it still works
    # (just press continue in the R console)

    output$codebook <- renderTable({
        if (length(input$variables) == 0) return(codebook)
        codebook %>% dplyr::select(!!!input$variables)
    }, rownames = TRUE)

})

# Run the application

shinyApp(ui = ui, server = server)

```

### 2020-05-05 final/app.R
Commit sad app compromise

@@ -14,7 +14,7 @@ library(haven)
 library(tidyverse)
 library(gt)
 library(devtools)
-#library(tibble)
+library(tibble)
 
 # I'm getting a warning about the foreign package when I try to deploy
 # Luckily, it's not necessary
@@ -23,7 +23,7 @@ library(devtools)
 library(scales)
 library(shinythemes)
 #library(plotly)
-library(leaflet)
+#library(leaflet)
 library(broom)
 # probably don't need these, but double check
 #library(patchwork)
@@ -69,60 +69,11 @@ shapefiles <- readRDS(file = "shapefiles.rds") %>%
 
 x <- readRDS("mwData.rds")
 
-# Now subset by year
+# Build a tibble for the map pngs
 
-# I tried saving these as Rds files as well, but I kept getting a weird error
-
-Census_1870 <- x %>% filter(YEAR == "1870") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1880 <- x %>% filter(YEAR == "1880") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-Census_1890 <- x %>% filter(YEAR == "1890") %>%
-    select_if(~ !all(is.na(.))) %>%
-    select(-STATEA:-GISJOIN & -COUNTY)
-
-# Code for the rest of the files
-# If only I could upload them all 
-
-#Census_1900 <- x %>% filter(YEAR == "1900") %>%
-    #select_if(~ !all(is.na(.))) %>%
-    #select(-STATEA:-GISJOIN & -COUNTY)
-
-#Census_1910 <- x %>% filter(YEAR == "1910") %>%
-    #select_if(~ !all(is.na(.))) %>%
-    #select(-STATEA:-GISJOIN & -COUNTY)
-
-#Census_1920 <- x %>% filter(YEAR == "1920") %>%
-    #select_if(~ !all(is.na(.))) %>%
-    #select(-STATEA:-GISJOIN & -COUNTY)
-
-#Census_1930 <- x %>% filter(YEAR == "1930") %>%
-    #select_if(~ !all(is.na(.))) %>%
-    #select(-STATEA:-GISJOIN & -COUNTY)
-
-#Census_1940 <- x %>% filter(YEAR == "1940") %>%
-    #select_if(~ !all(is.na(.))) %>%
[…]

### 2020-05-05 final/app.R
Commit app before restarting R

@@ -202,7 +202,7 @@ ui <- fluidPage(
                
                tabPanel("Maps",
                         titlePanel("Midwest Map Explorer"),
-                        p("Choose a census year and variable to view and compare maps of the midwest."),
+                        p("Choose a census year and variable to view maps of the midwest."),
                         helpText("Please note that no enrollment data was available for the year 1880."),
                         sidebarLayout(
                             
@@ -235,12 +235,14 @@ ui <- fluidPage(
                         titlePanel("Citations"),
                         p("Steven Manson, Jonathan Schroeder, David Van Riper, and Steven Ruggles. IPUMS National Historical Geographic Information System: Version 14.0 [Database]. Minneapolis, MN: IPUMS. 2019. http://doi.org/10.18128/D050.V14.0"),
                         p("The College of William and Mary and the Minnesota Population Center. School Attendance Boundary Information System (SABINS): Version 1.0. Minneapolis, MN: University of Minnesota 2011."),
-                        p("Ian Webster. U.S. Inflation Calculator: 1635-2020, Department of Labor Data. Official Data Foundation, https://www.officialdata.org.")
+                        p("Ian Webster. U.S. Inflation Calculator: 1635-2020, Department of Labor Data. Official Data Foundation, https://www.officialdata.org."),
+                        p("Todd Kosmerick. “World War I and Agriculture.” NC State University Libraries, 18 Aug. 2017, https://www.lib.ncsu.edu/news/special-collections/world-war-i-and-agriculture.")
                         ),
                
                tabPanel("Codebook",
                         titlePanel("How were variables measured for this project?"),
-                        p("Choose a variable to find out! Take special note of the changes in school enrollment and farm value metrics over time."), 
+                        p("Choose a variable to find out! Take special note of the changes in school enrollment and farm value metrics over time."),
+                        p("Please note that all dollar values are adjusted for inflation, using 2020 as the base year."),
                         varSelectInput("variables", "Variable:", codebook, multiple = FALSE),
                         helpText("These variable descriptions represent a condensed version of the codebooks provided by IPUMS NHGIS."),
                         tableOutput("codebook")

### 2020-05-05 final/app.R
Commit app with updated acreage explanation

@@ -111,7 +111,6 @@ ui <- fluidPage(
                                                  fluidRow(column(width = 5, textOutput("acre_com")),
                                                           column(width = 7, align = "center",
                                                                  imageOutput("acre_gif")))
-                                          
                                         ),
                                         tabPanel("Urbanization",
                                                  # Generate a row with a sidebar
@@ -259,13 +258,15 @@ server <- (function(input, output, session) {
     
     output$acre_com <- renderText({
         
-        "The bar plot on the left shows that a general upward trend in the number of acres allotted to farming each year. What happened in 1920? It’s difficult to say, but perhaps the Economic Boom of the 1920s followed by the Great Crash of 1929 caused farmers to purchase and then lose vast amounts of land. The line plot on the right shows that the number of farms increased until the early 20th century before beginning to decline in most states. Taken together, these plots show that farms in the Midwest have been steadily increasing in size, while the number of people owning farms has been steadily decreasing."
+        "The bar plot on the left show a general upward trend in the number of acres allotted to farming each year. What explains the jump in 1920? When the United States entered World War I in 1917, Europe was facing serious food shortages due to war casualties and farmers leaving their homes to fight. The US played an essential role in sustaining the Allied war effort through increased production— the government called for increases of up to 10% on many commodities.
+        The war spurred on the Economic Boom of the 1920s followed by the Great Crash of 1929, causing farmers to purchase and then sell off vast amounts of land. The line plot on the right shows that the number of farms increased until the early 20th century before beginning to decline in most states. Taken together, these plots show that farms in the Midwest have been steadily increasing in size, while the number of people owning farms has been steadily decreasing."
     })
     
     output$acre_gif <- renderImage({
         list(src = "acre_gif.gif")
         
     }, deleteFile = FALSE)
+    
     
     output$pop_com <- renderText({
         

### 2020-05-06 ms_8.Rmd
Final commit

@@
 # myvar options will be determined by column names of appPlan & displayed via dropdown widget
-myvar <- "Average Implement Value"
+myvar <- "Number of Cows per Farm"
 
@@
 
+```{r horses}
+# Do horses have the same effect as cows??
+
+# Prep data
+horseData <- allData %>% mutate(horses_per_farm = horses/n_farm)
+
+horseedu_model <- horseData %>%
+  select(YEAR, pct_enrolled, horses_per_farm, avg_value_farm) %>%
+  na.omit() %>%
+  lm(pct_enrolled ~ horses_per_farm*avg_value_farm, data = .) %>%
+  tidy(conf.int = TRUE)
+
+#saveRDS(cowedu_model, "final/cowedu_model.rds")
+
+horseedu_gt <- horseedu_model %>%
+  select(term, conf.low, estimate, conf.high) %>%
+  gt() %>%
+  tab_header(title = "Effect Horses and Farm Value Enrollment",
+             subtitle = "With 95% Confidence Interval") %>%
+  cols_label(term = "Coefficient Name", conf.low = "Lower Bound", estimate = "Estimated Value", conf.high = "Upper Bound") %>%
+  fmt_number(columns = vars(conf.low, estimate, conf.high), decimals = 4)
+
+horseedu_gt
+
+#gtsave(data = horseedu_gt, filename = "final/horseedu_gt.png")
+
+# horse plot
+
+horseedu_plot <- horseData %>%
+  select(YEAR, pct_enrolled, horses_per_farm) %>%
+  na.omit() %>%
+  mutate(YEAR = as.factor(YEAR)) %>%
+  group_by(YEAR) %>%
+  ggplot(aes(x = horses_per_farm, y = pct_enrolled, color = YEAR)) +
+  geom_point(alpha = 0.1) +
+  geom_smooth(method = "lm", se = FALSE) +
+  scale_color_viridis_d() +
+  ylim(0, 45) +
+  scale_x_log10(labels = comma) +
+  labs(title = "Horses and Education",
+       subtitle = "Relationship between the number of horses per farm\nand school enrollment, by county",
+       x = "Average Horses per Farm",
+       y = "Percent of Population Enrolled in School") +
+  theme_classic()
+
+horseedu_plot
+```
+
 ```{r farmTrends, cache=TRUE}

### 2020-05-06 final/app.R

@@ -179,7 +179,8 @@ ui <- fluidPage(
                                                        plotOutput("agPlot"),
                                                        p("The plot shows a positive correlation between school enrollment and the average number of cows per farm for the years where data was available. The number of cows might simply be a confounding variable that acts as a stand-in for farm size, but as the table below will show, this does not seem to be the case."),
                                                        gt_output("cows_size_gt"),
-                                                       p("This linear regression model shows that a one-cow increase in the number of cows per farm in a given county results in a 0.23% increase in school enrollment. This is significant given that school enrollment hovers around just 25% in most counties for the years studied. Curiously, farm size is not associated with a similarly significant increase in enrollment, and the interaction between cows per farm and farm size actually leads to a decrease in enrollment.")
+                                                       p("This linear regression model shows that a one-cow increase in the number of cows per farm in a given county results in a 0.23% increase in school enrollment. This is significant given that school enrollment hovers around just 25% in most counties for the years studied. Curiously, farm size is not associated with a similarly significant increase in enrollment, and the interaction between cows per farm and farm size actually leads to a decrease in enrollment."),
+                                                       p("An exploration of the effects of other livestock showed that horses are similarly associated with higher enrollment percentages, even when controlling for farm value. Horse data was only available for 1870 and 1890.")
                                               ),
                                               
                                               tabPanel("Gender",
@@ -225,7 +226,7 @@ ui <- fluidPage(
                         titlePanel("About this project"),
                         p("My name is Julia Englebert. I am a senior at Harvard College studying Near Eastern Languages and Civilizations with a secondary field in Ethnicity, Migration, and Rights. Some of my academic interests include the history and future of Midwestern agriculture, and the role of data science in shaping social policy."),
                         br(),
[…]
